### 2013-06-04 Papier/article.tex
Added Papers

\documentclass[reqno,11pt]{amsart}
\usepackage{amsmath, amssymb, amsthm}
\usepackage{url}
\usepackage{pst-tree}
\usepackage{algorithm}
\usepackage{algorithmicx}
\usepackage{algpseudocode}
\usepackage{pgfplots}
\theoremstyle{plain}
\newtheorem{prop}{Proposition}[section]
\theoremstyle{definition}
\newtheorem{defi}[prop]{Definition}
\newtheorem{nota}[prop]{Notation}
\newtheorem{exam}[prop]{Example}
\newtheorem{lem}[prop]{Lemma}
\newtheorem{fact}[prop]{Fact}

\title{Exploring the tree of numerical semigroups}
\author{XXXX}

\newcommand{\gr}[1]{{\color{gray} #1}}
%\newcommand{\card}{\#}
\newcommand{\ie}{\emph{i.e.}}
\newcommand{\kkp}{{k+1}}
\newcommand{\gen}[1]{\left<#1\right>}
\renewcommand{\leq}{\leqslant}
\renewcommand{\geq}{\geqslant}
\newcommand{\NN}{\mathbb{N}}
\renewcommand{\tt}[1]{\texttt{#1}}
\newcommand{\sgnode}[1]{\TR{\bf \left<#1\right>}}
\renewcommand{\ie}{\emph{i.e}}
\DeclareMathOperator{\Irr}{Irr}
\DeclareMathOperator{\App}{App}
\DeclareMathOperator{\card}{card}

\begin{document}

\begin{abstract}
In this paper we describe an algorithm visiting all the numerical semigroups up to a given genus using a new representation of numerical semigroups. 
\end{abstract}

\maketitle

\section{Introduction}

A \emph{numerical semigroup} $S$ is a subset of $\NN$ containing $0$, close under addition and of finite complement in $\NN$.  
For example the set 
\begin{equation}
\label{E:NSG}
S_E=\{0,3,6,7,9,10\}\cup[12,+\infty[
\end{equation}
is a numerical semigroup.
The \emph{genus} of a numerical semigroup $S$, denoted by~$g(S)$, is the cardinality of $\NN\setminus S$, \ie, $g(S)=\card(\NN\setminus S)$. 
 For example the genus of $S_E$ defined in \eqref{E:NSG} is $6$,  the cardinality of $\{1,2,4,5,8,11\}$

For a given positive integer $g$, the number of numerical semigroups  is finite and is denoted by $n_g$. 
In  J.A. Sloane's \emph{on-line encyclopedia of integer}~\cite{OEIS} we find the values of $n_g$ for $g\leq 52$. 
These values have been obtain by M. Bras-Amor\'os (view \cite{BrasAmoros2008} for more details for $g\leq 50)$. 
On his home page \cite{Delgado}, M. Delgado  gives the value of $n_{55}$ without specifying the values of $n_{53}$ and $n_{54}$. 
M.~-Bras Amor\'os used a depth first search exploration of the tree of numerical semigroups up to a given genus.
This tree was introduced by J.C. Rosales and al. in \cite{Rosales} and it is the subject of the Section~\ref{S:Tree}.

Starting with all the numerical semigroups of genus $49$ she obtained the number of numerical semigroups of genus $50$ in $18$ days on a pentium D runing at $3$GHz. 
In the package~\tt{NumericalSgs} \cite{NumericalSgps} of \tt{GAP} \cite{GAP}, M. Delgado together with P.A.~Garcia-Sanchez and J. Morais used the same method of exploration.

The paper is divided as follows.
In section~\ref{S:Tree} we describe the tree of numerical semigroups and give bounds for some parameters attached to a numerical semigroup.
In Section~\ref{S:DecNumber} we describe a new representation of numerical semigroups that is well suited to the construction of the tree. 
In Section~\ref{S:Algo} we describe an algorithm based on the representation given in Section~\ref{S:DecNumber} and give its complexity. 
Section~\ref{S:Opti} is more technical, and is devoted to the optimisation of the algorithm introduced in Section~\ref{S:Algo}.

\section{The tree of numerical semigroups}
\label{S:Tree}
We first start by some notations.

\begin{defi}
Let $S$ be a numerical semigroup. We define

$i)$ $m(S)=\min(S\setminus\{0\})$, the \emph{multiplicity} of $S$;

$ii)$ $g(S)=\card(\NN\setminus S)$, the \emph{genus} of $S$;

$iii)$ $c(S)=1+\max(\NN\setminus S)$, the \emph{conductor} of $S$ for $S$ different from $\NN$. By convention the conductor of $\NN$ is $0$. 
\end{defi}

By definition a numerical semigroup is an infinite object. 
We need a finite description of such a semigroup. 
That is the role of generating sets.

\begin{defi}
A subset $X$ of a semigroup is a \emph{generating set} of $S$ if every element of $S$ can be express as a sum of elements in $X$. 
\end{defi}

\begin{nota}
A numerical semigroup admitting $X=\{x_1<x_2<...<x_n\}$ as generating set is denoted  by $S=\left<x_1,...,x_\ell\right>$.
\end{nota}

We now introduce a specific generating set.

\begin{defi}
A non-zero element $x$ of a numerical semigroup $S$ is said to be \emph{irreducible} if it cannot be expressed as a sum of two non-zeros elements of $S$.  
We denote by $\Irr(S)$ the set of all irreducible elements of $S$.
\end{defi}

\begin{prop}
Let $S$ be a numerical semigroup.
Then $\Irr(S)$ is the minimal generating set of $S$.
\end{prop}

\begin{proof}
Assume for a contradiction, that there exists an integer  $x$ in $S$ than cannot be decomposed as a sum of irreducible elements. 
We may assume that $x$ is  minimal with this property. 
As  $x$ cannot be irreducible, there exist $y$ and $z$ in~$S\setminus\{0\}$ satisfying $x=y+z$. 
Since we have $y<x$ and $z<x$, the integers $y$ and $z$ can be expressed as a sum of irreducible elements of $S$ and so $x=y+z$ is a sum of irreducible elements, in contradiction to hypothesis.

As irreducible elements of $S$ cannot be decomposed as the sum of two non-zeros integers in $S$, they must occur in each generating set of $S$.
\end{proof}

We recall that Ap\'ery elements of a numerical semigroup $S$ associated to $m(S)$ are the integers $x$ in $S$ such that $x-m(S)$ is no longer in $S$.
We denote by~$\App(S)$ the set of these elements. 
It is well known that the cardinality of $\App(S)$ is exactly $m(S)$ (see~\cite{Book} for example). 
Note that the set $\Irr(S)$ is in included in $\App(S)$.
In particular, the set $\Irr(S)$ is finite and its cardinality is at most $m(S)$.

If we reconsider the numerical semigroup of \eqref{E:NSG}, we obtain
\begin{equation}
\label{E:GNSG}
S_E=\{0,3,6,7,9,10\}\cup[12,+\infty[=\left<3,7\right>
\end{equation}

Let  $S$ be a numerical semigroup. The set $T=S\cup\{c(S)-1\}$ is also a numerical  semigroup and its genus is $g(S)-1$. 
As  $[c(S)-1,+\infty[$ is included in $T$ we have $c(T)\leq c(S)-1$. 
Therefore every semigroup $S$ of genus $g$ can be obtained from a semigroup $T$ of genus $g-1$ by removing an element of $T$ greater than or equal to $c(T)$.

\begin{prop}
\label{P:Sx}
Let $S$ be a numerical semigroup and $x$ an element of $S$. The set~$S^x=S\setminus\{x\}$ is a numerical semigroup if and only if $x$ is irreducible in $S$.
\end{prop}

\begin{proof}
If $x$ is not irreducible in  $S$, then there exist $a$ and $b$ in $S\setminus{0}$ such that $x=a+b$. 
Since $a\not=0$ and $b\not=0$ hold, the integers  $a$ and $b$ belong to $S\setminus\{x\}$. 
Since $x=a+b$ and $x\not\in S^x$, it follows that  $S^x$ is not stable under addition.

Conversely, assume that $x$ is irreducible in  $S$. 
As $0$ is never irreducible, the set~$S^x$ contains~$0$. 
Let $a$ and $b$ be two integers belonging to $S^x$. 
The set $S$ is stable under addition, hence $a+b$ lies in $S$. 
As $S$ is equal to $S^x\cup\{x\}$, the integer $a+b$ also lies in $S^x$ except if it is equal to $x$. 
The latter is impossible since $a$ and $b$ are different from $x$ and $x$ is irreducible.
\end{proof}

Proposition~\ref{P:Sx} implies that every semigroup $T$ of genus $g$ can be obtained from a semigroup $S$ by removing a generator $x$ of $S$ that is greater than $c(S)$.
Hence relations~$T=S^x=S\setminus\{x\}$ hold.

We construct the tree of numerical semigroups, denoted by $\mathcal{T}$ as follows. 
The root of the tree is the unique semigroup of genus $0$, \ie, $\left<1\right>$ that is equal to $\NN$. 
If~$S$ is a semigroup in the tree,  the sons of $S$ are exactly the semigroup $S^x$ where $x$ belongs to $\Irr(S)\cap[c(s),+\infty[$. 
By convention, when depicting the tree, the ,numerical semigroup $S^x$ is in the left of $S^y$ if $x$ is smaller than $y$.

The above remarks imply that a semigroup $S$ has depth $g$ in $≠mathcal{T}$ if and only if its genus is~$g$, see Figure~\ref{F:Tree}.
We denote by $\mathcal{T}^{\leq g}$ the subtree of $\mathcal{T}$ resticted to all semigroup of genus $\leq g$.

 \begin{figure}[htf!]
 \[
 \pstree[,nodesep=3pt]{\sgnode{1}}{
 	\pstree[nodesep=3pt]{\sgnode{2,3} \tlput1}{
 		\pstree[treesep=4pt,nodesep=2pt]{\sgnode{3,4,5} \tlput2}{
 			\pstree[treesep=3pt,nodesep=3pt]{\sgnode{4,5,6,7} \tlput3}{
 				\pstree{\sgnode{5,6,7,8,9} \tlput4}{}
 				\pstree{\sgnode{\gr 4,6,7,9} \tlput5}{}
 				\pstree{\sgnode{\gr 4,\gr 5,7} \trput6}{}
 				\pstree{\sgnode{\gr 4,\gr 5,\gr 6} \trput7}{}					
 			}
 			\pstree[treesep=3pt,nodesep=3pt]{\sgnode{\gr 3,5,7} \tlput4}{
 				\pstree{\sgnode{\gr 3,7,8} \tlput 5}{}
 				\pstree{\sgnode{\gr 3,\gr 5} \trput 7}{}
 			}
 			\pstree[treesep=10pt]{\sgnode{\gr 3,\gr 4} \trput5}{\Tn \Tn \Tn}
 		}
 		\pstree[nodesep=3pt]{~\hspace{-5em}\sgnode{\gr2,5} \trput3}{
 			\pstree[nodesep=3pt]{\sgnode{\gr2,7} \trput5}{
 				\pstree[nodesep=3pt]{\sgnode{\gr2,9} \trput7}{}
 			}
 		}
 	}
 }
 \vspace{-5em}
 \]
\caption{The first five layers of the tree $\mathcal{T}$ of numerical semigroups, corresponding to $\mathcal{T}^{\leq 5}$. A generator of a semigroup is in gray if it is not in $[c(S),+\infty[$. An edge between a semigroup $S$ and its son $S'$ is labelled by  $x$ is $S'$ is obtained from $S$ by removing~$x$, that is $S'=S^x$.}
\label{F:Tree}
\end{figure}

As the reader can check, the main difficulty to characterize the sond of a semigroup is to determine its irreducible elements.
 In \cite{NumericalSgps}, the semigroup are given by their Ap\'ery set $\App(S)$ and then the main difficulty is to describe $\App(S_x)$ from $\App(S)$. 
 This approach is elegant but not sufficiently \emph{basic} for our optimisations.

We conclude this section with some basic results on numerical semigroups of a given genus.
Let $S$ be a numerical semigroup. 
We first prove :
\begin{equation}
\label{E:Irr}
x\in \Irr(S)\quad\textrm{implies}\quad x\leq c(S)+m(S).
\end{equation}
If $y$ is a positive integer with $y \geq c(S)+m(S)$ then $y$ lies in $S$ (as $y\geq c(S)$ holds). Moreover, we always have $y-m(S)\geq c(s)$ and so $y$ is not irreducible.

As $\NN\setminus{S}$ contains $\{1,..,m(S)-1\}$ we have 
\begin{equation}
\label{E:m}
m(S)\leq g(S)+1.
\end{equation}
Let $x$ be an element of $S\cap[0,c(S)-1]$. 
The integer $y=c(S)-1-x$ lies in $[0,c(S)-1]$. 
Moreover $y$ is not in $S$, for otherwise  we write $c(S)-1=y+x$ where $x,y$ are elements of $S$ implying $c(S)-1\in S$. 
In contradiction with the definition of conductor. 
Thus we define an involution 
\[
\begin{array}{rcl}
\psi:[0,c(S)-1]&\to&[0,c(S)-1]\\
x&\mapsto&c(S)-1-x
\end{array}
\]
that send $S\cap[0,c(S)-1]$ into $[0,c(S)-1]\setminus S$.
The cardinality of $[0,c(S)[\setminus S$ is exactly $g(S)$. Let us denote by $k$ the cardinality  of $S\cap[0,c(S)-1]$. 
Since $\psi$ is injective we must have $k\leq g(S)$. 
>From the relation 
\[
[0,c(S)[=S\cap[0,c(S)[\sqcup[0,c(S)[\setminus S
\]
 we obtain $c(S)=k+g(S)$ and so 
\begin{equation}
\label{E:c}
c(S)\leq 2 g(S).
\end{equation}

\section{Decomposition number}
\label{S:DecNumber}

The aim of this section is to describe a new representation of numerical semigroups.
which is well suited to a efficient exploration of the tree~$\mathcal{T}$ of numerical semigroups.

\begin{defi}
Let $S$ be a numerical semigroup.
For every $x$ of $\NN$  we set
\[
D_S(x)=\{y \in S\ |\ x-y\in S\ \text{and}\ 2y\leq x\}
\]
and $d_S(x)=\card D_S(x)$.
We called $d_S(x)$ the $S$-\emph{decomposition number} of $x$.
The application $d_S:\NN\to\NN$ is the $S$-\emph{decomposition numbers function}.
\end{defi}

Assume that $y$ is an element of $D_S(x)$.
By very definition of $D_S(x)$, the integer~$z=x-y$ also belongs to $S$. 
Then $x$ can be decomposed as $x=y+z$ with $y$ and $z$ in~$S$. 
Moreover the condition $2y\leq x$ implies $y\leq z$. 
In other words if we define~$D'_S(x)$ to be the set of all $(y,z)\in S\times S$ with $x=y+z$ and $y\leq z$ then~$D_S(x)$ is the image of $D'_S(x)$ under the projection on the first coordinate.
 Hence~$D_S(x)$ describes how~$x$ can be decomposed as sums of two elements of $S$,-.
 This is why $d_S(x)$ is called the $S$-decomposition number of $x$.

\begin{exam}
Reconsider the semigroup $S_E$ given at \eqref{E:NSG}.
The integer $14$ admits three decompositions as sums of two elements of $S$, namely $14=0+14$, $14=3+11$ and $14=7+7$. 
Thus the set $D_{S_E}(14)$ is equal to $\{0,3,7\}$ and $d_{S_E}$ equals $3$.
\end{exam}

\begin{lem}
\label{L:RepN}
For every numerical semigroup $S$ and every integer $x\in \NN$,  we have $d_S(x)\leq 1+\left\lfloor \dfrac{x}2\right\rfloor$ and the equality holds for $S=\NN$.
\end{lem}

\begin{proof}
As the set $D_S(x)$ is included in $\{0,...,\left\lfloor\frac x2\right\rfloor\}$, the relation $d_S(x)\leq 1+\left\lfloor \frac{x}2\right\rfloor$ holds.
For $S=\NN$ we have the equality for $D_S(x)$ and so for $d_S(x)$.
\end{proof}

\begin{prop}
\label{P:Dandd}
Let $S$ be a numerical semigroup and $x\in\NN\setminus\{0\}$.  We have:

 $i)$ $x$ lies in $S$ if and only if $d_S(x)>0$.

 $ii)$ $x$ is in $\Irr(S)$ if and only if $d_S(x)=1$.
\end{prop}

\begin{proof}
We start with $i)$. If $x$ is an element of $S$ then $x$ equals $0+x$. 
The relation~$2\times 0\leq x$ and $0\in S$ imply that $D_S(x)$ contains $0$, and so $d_S(x)>0$ holds.
Conversely, the relation  $d_S(x)>0$  implies that $D_S(x)$ is non-empty. 
Let $y$ be an element of~$D_S(x)$. 
As $y$ and $x-y$ belong to  $S$, by definition, the integer $x=(x-y)+y$ is in $S$  (since $S$ is stable by addition).

Let us show $ii)$. 
Assume $x$ is irreducible in $S$.
There cannot exist $y$ and $z$ in~$(S\setminus\{0\})^2$ such that $x=y+z$. 
The only possible decomposition of $x$ as a sum of two elements of $S$ is $x=0+x$. 
Hence, the set $D_S(x)$ is equal to~$\{0\}$ and we have $d_S(x)=1$. 
Conversely, let $x$ such that $d_S(x)=1$. 
By  $i)$ the  integer $x$ must be in $S$. 
As $x=0+x$ is always a decomposition of $x$ as a sum of two elements in $S$, we obtain $D_S(x)=\{0\}$.
If there exist $y$ and $z$ in $S$ such that $y\leq z$ and $x=y+z$ hold then $y$ lies in $D_S(x)$. 
This implies $y=0$ and $z=x$. 
Hence $x$ is irreducible in~$S$.
\end{proof}

We note  that $0$ is never irreducible despite the fact $d_S(0)$ is $1$ for all numerical semigroup $S$.

We now explain how to compute the $S$-decomposition numbers function of a numerical semigroup  from these of its father.

\begin{prop}
Let $S$ be a numerical semigroup and $x$ be an irreducible element of $S$. 
Then for all $y\in \NN\setminus\{0\}$ we have
\[
d_{S^x}(y)=\begin{cases}
d_S(y)-1&\text{if $y\geq x$ and $d_S(y-x)>0$,}\\
d_S(y) &\text{otherwise.}
\end{cases}
\] 
\end{prop}

\begin{proof}
Let $y$ be in $\NN\setminus\{0\}$. We have 
\[
D_{S^x}(y)=\{z \in S^x\ |\ y-z\in S^x\ \text{and}\  2z\leq y\}
\]
and $D_{S^x}(y)$ is a subset of $D_S(y)$.
We have $D_S(y)\setminus D_{S^x}(y)=E\sqcup F$ where
\begin{align*}
E&=\begin{cases}
\{x\} &\text{if $y-x\in S^x$ and $2x\leq y$,}\\
\emptyset & \text{otherwise.}
\end{cases}\\
F&=\{z\in S\ |\ y-z=x\ \text{and}\ 2z\leq y\}
\end{align*}
Since  the relation $y-x\in S^x$ can be rewritten as the conjunction of $y-x\in S$ and $y\not=2x$, we obtain
\[
E=\begin{cases}
\{x\} &\text{if $y-x\in S$ and $2x< y$,}\\
\emptyset & \text{otherwise.}
\end{cases}
\]
By Proposition~\ref{P:Dandd}, the relation $y-x\in S$ is equivalent to $y\geq x$ and $d_S(y-x)>0$, we have
\[
E=\begin{cases}
\{x\} &\text{if $y\geq x$ and $d_S(y-x)>0$ and $2x< y$,}\\
\emptyset & \text{otherwise.}
\end{cases}
\]
On the other hand, we have 
\begin{align*}
F&=\{z\in S\ |\ y-z=x\ \text{and}\ 2z\leq y\}\\
&=\{z\in S\ |\ z=y-x\ \text{and}\ y\leq 2x\}\\
&=\begin{cases}
\{y-x\} & \text{if $y-x\in S$ and $y\leq 2x$,}\\
\emptyset & \text{otherwise.}
\end{cases}
\end{align*}
As in the case of $E$, we get 
\[
F=\begin{cases}
\{y-x\} & \text{if $y\geq x$ and $d_S(y-x)>0$ and $y\leq 2x$,}\\
\emptyset & \text{otherwise.}
\end{cases}
\]
In $E$ we have the constraint $2x<y$ while in $F$ we have $y\leq 2x$, hence only one of the sets $E$ or $F$ can be non-empty. 
This implies
\[
D_S(y)\setminus D_{S^x}(y)=
\begin{cases}
\{x\} & \text{if $y\geq x$ and $d_S(y-x)>0$ and $y\leq 2x$,}\\
\{y-x\} & \text{if $y\geq x$ and $d_S(y-x)>0$ and $y>2x$,}\\
\emptyset & \text{otherwise.}
\end{cases}
\]
Therefore we obtain
\[
d_{S^x}(y)=\begin{cases}
d_S(y)-1&\text{if $y\geq x$ and $d_S(y-x)>0$,}\\
d_S(y)& \text{otherwise.}
\end{cases}
\]
\end{proof}

\section{A new algorithm}
\label{S:Algo}
We can easily explore the tree of numerical semigroups up to a genus $G$ using a depth search first algorithm using a stack (see~\ref{A:Exp}). 
This approach does not seem to have been used before. 
In particular,  M. Bras-Amor\'os and M. Delgado use instead a breadth search first exploration. 
The main advantage in our approach is the small memory needs. 
The cost to pay is that, if we want to explore the tree deepero, we must restart from the root.

\begin{algorithm}
\caption{Depth search first exploration of the tree of numerical semigroups}\label{A:Exp}
\begin{algorithmic}[1]
\Procedure{Explore}{\tt{G}}
\State Stack \tt{stack} \Comment{the empty stack}
\State $\tt{S} \gets\left<1\right>$
\While{stack is not empty} 
	\State $\tt{S} \gets \tt{stack.top()}$
	\State \tt{stack.pop()}
	\If{$g(\tt{S})< \tt{G}$}
		\For{\tt{x} from $c(\tt{S})$ to $c(\tt{S})+m(\tt{S})$}
			\If{$\tt{x}\in \Irr(\tt{S})$}
				\State $\tt{S.push}(\tt{S}^\tt{x})$
			\EndIf
		\EndFor
	\EndIf
\EndWhile
\EndProcedure
\end{algorithmic}
\end{algorithm}

In Algorithm~\ref{A:Exp} we do not specify how to compute $c(S)$, $g(S)$ and $m(S)$  from $S$ neither how to test if an integer is irreducible. 
It also miss  the characterisation of $S^x$ from $S$.
These items depend heavily of the representation of $S$.
Our choice is to use the $S$-decomposition numbers function. 
The first task is to use a finite set of such numbers to characterise the whole semigroup.

\begin{prop}
\label{P:delta}
Let $G$ be an integer and $S$ be a numerical semigroup of genus $g\leq G$.
Then $S$ is entirely described by $\delta_S=(d_S(0),...,d_S(3\,G))$. More precisely we can obtain $c(S), g(S), m(S)$ and $\Irr(S)$ from $\delta_S$.
\end{prop}

\begin{proof}
By~\eqref{E:c} we have $c(S)\leq g(S)$ and so the $S$-decomposition number of $c(S)$ occurs in $\delta_S$.
Since $c(S)$ is equal to $\max(\NN\setminus S)$, Proposition~\ref{P:delta} implies 
\[
c(S)=1+\max\{i\in[0,...,3G],\ d_S(i)=0\}.
\]
As all elements of $\NN\setminus S$ are smaller than $c(S)$, their $S$-decomposition numbers are in $\delta_S$ and we obtain
\[
g(S)=\card\{i\in[0,..,3G],\ d_S(i)=0\}.
\]
By \eqref{E:m} the relation $m(S)\leq g(S)+1$ holds. 
This implies that the $S$-decomposition number of $m(S)$ appears in $\delta_S$ : 
\[
m(S)=\min\{i\in[0,...,3G],\ d_S(i)>0\}.
\]
Since, by \eqref{E:Irr},  all irreducible elements are smaller than $c(S)+m(S)-1$, which is itself smaller than $3G$, equations  \eqref{E:m} and \eqref{E:c} give
\[
\Irr(S)=\{i\in[0,...,3G],\ d_S(i)=1\}.
\]
\end{proof}

Even though it is quite simple, the computation of $c(S),m(S)$ and $g(S)$ from $\delta_S$ has a non negligible cost.
We represent a numerical semigroup $S$ of genus $g\leq G$ by $(c(S),g(S),c(S),\delta_S)$. 
In an algorithmic context, if the variable \texttt{S} stands for a numerical semigroup  we use:

-- \tt{S.c}, \tt{S.g} and \tt{S.m} for the integers $c(S)$, $g(S)$ and $m(S)$;

-- \tt{S.d[i]} for the integer $d_S(i)$.

For example the following Algorithm initializes a representation of the semigroup $\NN$ ready for an exploration up to genus $G$.

\begin{algorithm}
\caption{Return the root of $\mathcal{T}$ for an exploration up to genus \tt{G}}\label{A:Root}
\begin{algorithmic}
\Function{Root}{\tt{G}}
\State \tt{R.c} $\gets$ 1 \Comment{\tt{R} stands for $\NN$}
\State \tt{R.g} $\gets$ 0
\State \tt{R.m} $\gets$ 1 
\For{\tt{x} from $1$ to $3\,\tt{G}$}
	\State\tt{R.d[x]}$\gets1+\left\lfloor\frac{\tt{x}}2\right\rfloor$
\EndFor
\State{\textbf{return} \tt{R}}
\EndFunction
\end{algorithmic}
\end{algorithm}

We can now describe an algorithm that returns the representation of the semigroup $S^x$ from that of the semigroup $S$ where $x$ is an irreducible element of $S$ greater than $c(S)$.

\begin{algorithm}
\caption{Returns the son $S_x$ of $S$ with $x\in\Irr(S)\cap[c(S),c(S)+m(S)[$.}\label{A:Son}
\begin{algorithmic}[1]
\Function{Son}{\tt{S},\tt{x},\tt{G}}
\State \tt{T.c} $\gets$ $\tt{x}+1$ \Comment{\tt{T}  stands for $\tt{S}_x$}
\State \tt{T.g} $\gets$ $\tt{S.g}+1$
\If{$\tt{x}>\tt{S.m}$}
	\State \tt{T.M} $\gets$ \tt{S.m}
\Else
	\State \tt{T.M} $\gets$ $\tt{S.m}+1$
\EndIf
\State\tt{T.d} $\gets$ \tt{S.d}
\For{\tt{y} from \tt{x} to $3\,\tt{G}$}
	\If{$\tt{S.d}[\tt{y}-\tt{x}]>0$}
		\State $\tt{T.d}[\tt{y}]\gets\tt{T.d}[\tt{y}]-1$
	\EndIf
\EndFor
\State{\textbf{return} \tt{T}}
\EndFunction
\end{algorithmic}
\end{algorithm}

\begin{prop}
\label{P:Son}
Running on $(S,x,G)$ with $g(S)\leq G$, $x\in \Irr(S)$ and~$x\leq c(S)$, Algorithm~\ref{A:Son} returns the numerical semigroup $S_x$ in time $O(\log(G)\times G)$.
\end{prop}

\begin{proof}
By construction $S^x$ is the semigroup $S\setminus\{x\}$.
Thus the genus $S^x$ is $g(S)+1$, see Line~\tt{1}.
Every integer of $I=[x+1,+\infty[$ lies in $S$ since $x$ is greater than $c(S)$, so the interval $I$ is included in $S^x$.
As $x$ does not belong to $S$, the conductor of $S^x$ is $x+1$, see Line \tt{2}.
For the multiplicity of $S^x$ we have two cases. 
First, if $x>m(S)$ holds  then $m(S)$ is also in $S^x$ and so $m(S^x)$ is equal to $m(S)$.
Assume now $x=m(S)$. 
The relation $x(S)\geq c(S)$ and the characterisation of~$m(S)$ implies $x=m(S)=c(S)$.
Thus $S^x$ contains $m(S)+1$ which is $m(S^x)$.
The initialisation of $m(S^x)$ is done by Lines \tt{4} to \tt{8}. 
The correctness of the computation of $\delta_{S^x}$ (see Proposition~\ref{P:delta}) done from Line \tt{9} to Line \tt{15} is a direct consequence of~Proposition~\ref{P:Dandd}.

Let us now prove the complexity statement. 
Since by \eqref{E:c} and \eqref{E:m} we have $x\leq 3G$ together with $m(S)\leq G+1$, each line from \tt{2} to \tt{8} is done in time $O(\log(G))$.
The \textbf{for} loop needs $O(G)$ steps and each step is done in time $O(\log(G))$. 
Summarizing, these results give that the algorithm runs in time $O(\log(G)\times G)$.
\end{proof}

\begin{algorithm}
\caption{Returns an array containing the value of $n_g$ for $g\leq G$}\label{A:Count}
\begin{algorithmic}[1]
\Function{Count}{\tt{G}}
\State \tt{n} $\gets [0,...,0]$ \Comment $\tt{n[g]}$ stands for $n_g$ and is initialised to $0$
\State Stack \tt{stack} \Comment{the empty stack}
\State $\tt{S} \gets \textsc{Root}(\tt{G})$
\While{stack is not empty} 
	\State $\tt{S} \gets \tt{stack.top()}$
	\State \tt{stack.pop()}
	\State $\tt{n[S.g]}\gets\tt{n[S.g]}+1$
	\If{$\tt{S.g}< \tt{G}$}
		\For{\tt{x} from $\tt{S.c}$ to $\tt{S.c}+\tt{S.m}$}
			\If{$\tt{S.d[x]\,=\,1}$}
				\State $\tt{S.push}(\textsc{Son}(\tt{S},\tt{x},\tt{G}))$
			\EndIf
		\EndFor
	\EndIf
\EndWhile
\State{\textbf{return} \tt{n}}
\EndFunction
\end{algorithmic}
\end{algorithm}

\begin{prop}
Running on $G\in\NN$, Algorithm \textsc{Count} returns the array~$[n_0,...,n_G]$ in time
\[
O\left(\log(G)\times G \times \sum_{g=0}^G n_g\right)
\]
and its space complexity is $O(\log(G)\times G^3)$.
\end{prop}

\begin{proof}
The correctness of the algorithm is a consequence of Proposition~\ref{P:Son} and of the description of the tree~$\mathcal{T}$ of numerical semigroups.

For the time complexity, let us remark that Algorithm~\textsc{Son} is called for every semigroup of the tree $\mathcal{T}^{\leq G}$ (the restriction of $\mathcal{T}$ to semigroup of genus $\leq G$). 
Since there are exactly $N=\sum_{g=0}^G n_g$ such semigroups, the time complexity of \textsc{Son} established in Proposition~\ref{P:Son} guarantees that the running time of \textsc{Count} is in~$O(\log(G)\times G\times N)$, as stated.

Let us now prove the space complexity statement.
For  this we need to describe the stack through  the run of the algorithm.
Since the stack is filled  with a depth first search algorithm, it has two properties.
The first one is that reading the stack from the bottom to the top, the genus of  semigroup increases.  
The second one is that, for all $g\in[0,G]$, every semigroup of genus $g$ in the stack has the same father.
As the number of sons of a  semigroup $S$ is the number of $S$-irreducible elements in $[c(S),c(S)+m(S)-1]$,  a semigroup $S$ has at most $m(S)$ sons. 
By~\eqref{E:m}, this implies that a semigroup of genus $g$ as at most $g+1$ sons.
Therefore the stack contains at most $g+1$ semigroup of genus $g+1$ for $g\leq G$. 
So the size of the stack is bounded~by 
\[
S=\sum_{g=0}^Gg=\frac{G(G+1)}2
\]
A semigroup is represented by a quadruple $(c(S),g(S),m(S),\delta_S)$. 
By equations~\eqref{E:c} and~\eqref{E:m}, we have $c\leq 2g(S)$ and $m\leq g(S)+1$. 
As $g(S)\leq G$ holds, the integers~$c$, $g$  and $m$ of the representation of $S$ require a memory space in $O(\log(G))$. 
The size of~$\delta_S=(d_S(0),...,d_S(3G))$ is exactly~$3G+1$. 
Each entry of $\delta_S$ is the $S$-decomposition number of an integer smaller than $3G$ and hence
requires $\log(G)$ bytes of memory space.
Therefore the space complexity of $\delta_S$ is in~$O(\log(G)\times G)$, which implies that the space complexity of the \textsc{Count} algorithm is
\[
O(\log(G)\times G\times S)= O(\log(G)\times G^3).
\]
\end{proof}

\section{Technical optimizations and results}
\label{S:Opti}

Assume for example that we want to construct the tree $\mathcal{T}^{\leq 100}$ of all  numerical semigroup of  genus smaller than $100$. 
In this case, the representation of  numerical  semigroup given in Section~\ref{S:DecNumber} uses decomposition numbers of integers smaller than~$300$.
By Lemma~\ref{L:RepN}, such a decomposition number is smaller than $151$ and requires~$1$ byte of memory.
Thus at each \textbf{for} step of Algorithm~\textsc{Son}, the CPU actually works on $1$ byte. 
However current CPUs work on $8$ bytes.
The first optimization uses this point.

To go further we must specify that the array of decomposition numbers in the representation of a semigroup corresponds to consecutive bytes in memory. 
In the \textbf{for} loop of Algorithm~\textsc{Son} we may imagine two cursors: the first one, denoted \texttt{src}  pointing to the memory byte of $\texttt{S.d[0]}$  and the second one, denoted \texttt{dst} pointing to the memory byte $\texttt{T.d[y]}$.
Using these two cursors, Lines \tt{10} to \tt{14} of Algorithm~\textsc{Son} can be rewritten as follows

\begin{algorithmic}
\State $\tt{src}\gets \text{address}(\tt{S.d[0]})$
\State $\tt{dst}\gets \text{address}(\tt{T.d[x]})$
\State $\tt{i}\gets 0$
\While{$\tt{i}\leq \tt{3G}-\tt{x}$}
\If{$\text{content}(\tt{src})>0$}
\State $\textbf{decrease}\ \text{content}(\tt{dst})$ \textbf{by} $1$\EndIf
\State \textbf{increase} \tt{src},\tt{dst},\tt{i} \textbf{by} $1$

\EndWhile
\end{algorithmic}

\noindent In this version we can see that the cursors \texttt{src} and \texttt{dst} move at the same time and that the modification of  the  value pointed by \texttt{dst}   only  needs to access the values pointed by \tt{src} and \tt{dst}. 
We can therefore work in multiple entries at the same time without collision. 
Current CPUs allow this thanks to the SIMD technologies as MMX, SSE, etc.
The acronym SIMD stands for Single Operation Multiple Data.

The MMX technology  permits to work on $8$ bytes in parallel while the SSE works in $16$ bytes. 
As the rest of the CPU works on $8$ bytes, the SSE technology needs some constraint on memory access than cannot be fulfilled in our algorithm. 
This motivate our choice to use the MMX technology. 
More precisely, we use three commands: \tt{pcmpeqb}, \tt{pandn} and \tt{psubb}.
 These commands work on two arrays of $8$ bytes, called~\tt{s} and \tt{d} here. 
 In each case the array \tt{d} is modified. 
 We denote bytes by $\{\tt{0},\tt{1}\}$-words of length $8$.
 After a call to \tt{pcmpeqb(s,d)} the $i$th entry of $d$ contains \tt{11111111} if $\tt{s}[i]=\tt{d}[i]$ holds and \tt{00000000} otherwise. 
 The command \tt{pand(s,d)}  store in $\tt{d}[i]$  the value of the byte-logic operation \tt{s}[i] \tt{and} (\tt{not} \tt{d}[i]).
  When \tt{pand(s,d)} is called the new value of $\tt{d}[i]$ is $\tt{s}[i]-\tt{d}[i]$.  
  Glueing all the pieces together we obtain the following version of the \textbf{for} loop of Algorithm~\textsc{Son} :

\begin{algorithmic}[1]
\State $\tt{src}\gets \text{address}(\tt{S.d[0]})$
\State $\tt{dst}\gets \text{address}(\tt{T.d[x]})$
\State $\tt{i}\gets 0$
\While{$\tt{i}\leq \tt{3G}-\tt{x}$}
\State $\tt{t}\gets [\tt{00000000},...,\tt{00000000}]$ \Comment{$8$ bytes equal to \tt{00000000}}
\State \tt{pcmpeqb(\tt{src},t)}
\State \tt{pandn([00000001,...,00000001],t)}
\State \tt{psub(\tt{dst},t)}
\State $\tt{dst}\gets \tt{t}$ \Comment{copy the array $[\tt{t}[0],...,\tt{t}[7]]$ to $[\tt{d}[0],...,\tt{d}[7]]$}
\State \textbf{increase} \tt{src},\tt{dst},\tt{i} \textbf{by} $8$
\EndWhile
\end{algorithmic}

Let us explain how this works in more details. 
Let $i$ be an integer in $\{0,...,7\}$.  
After Line~\tt{5}, the value of $\tt{t}[i]$ is \tt{0000000}. 
After Line~\tt{6}, we have 
\[
\tt{t}[i]=\begin{cases}
\tt{11111111}&\text{if $\tt{src}[i]=\tt{00000000}$ (corresponding to $0$)}\\
\tt{00000000}&\text{otherwise}
\end{cases}
\]
Line \tt{7} performs a logical \tt{and} between $\tt{00000001}$ and $\tt{not t}[i]$. Hence the result byte is \tt{00000001} if the last byte of \tt{t}[i] is $0$ and \tt{00000000} otherwise: 
\[
\tt{t}[i]=\begin{cases}
\tt{00000000}&\text{if $\tt{src}[i]=\tt{00000000}$ (corresponding to $0$)}\\
\tt{00000001}&\text{otherwise}
\end{cases}
\]
 Line \tt{8} subtracts \tt{t}[i] (which is equal to $0$ or $1$) from $\tt{dst}[i]$ according to Proposition~\ref{P:Sx}. 
Finally we shift the cursor \tt{src} and \tt{dst} by eight cases.

Our second optimization  is to use parallelism on exploration of the tree.
Today, CPU of personal computer have several cores (2, 4 or more). 
The given version of our exploration algorithm  uses a single core and so a fraction only of the power of a CPU.

Our method of parallelism is very simple: for $G\in \NN$,  we cut the tree $\mathcal{T}^{\leq G}$ in sub-trees $\mathcal{T}_1,...,\mathcal{T}_n$ and we launch our algorithm on these sub-trees. 
The advantage of this method is that there is no communication between the instances of our algorithm. 
The disadvantage is that the cutting controls the efficiency of the parallelism. 
Assume for example that we want to explore the tree~$\mathcal{T}^{\leq 40}$.
We first determine all numerical semigroups $S_1,...S_{n_{20}}$ of genus $20$. 
To explore $\mathcal{T}$ it remains to explore the tree $\mathcal{T}_i$ rooted in $S_i$ for $i=1,...,n_{20}$. 
This works but the time to explore $\mathcal{T}_1$ is similar to the time need to explore the tree $\mathcal{T}^{\leq 40}$. 
And in this case, using many cores does not reduce the time to explore the tree.

Let us now explain in more detail how we cut the tree $\mathcal{T}^{\leq G}$ in order to use parallelism.
Semigroups of the form $\left<g+1,...,2g+1\right>$, which are of genus $g$, are called \emph{ordinary} in~\cite{Elizalde}. 
Each ordinary semigroup has a unique son that is also ordinary. 
We define $X$ to be the set of all the non-ordinary sons of an ordinary semigroup and $X^{\leq G}$ the restriction of $X$ to semigroup of genus $\leq G$. 
We then denote by  $\mathcal{T}_i$ the tree rooted on $S_i$ where $X^{\leq G}=\{S_1,...,S_n\}$. 
The time needed to explore $\mathcal{T}_i$ is very heterogeneous but there are many tree $T_i$ with maximal time. 
This cutting is more efficient than the previous one.

Figure~\ref{F:Time} summarizes the time complexity of various exploration algorithms.

The version $\small depth-expl-\delta-mmx$ of our algorithm compute the value of $n_g$ for $g\leq 50$ in $196$ minutes on the \tt{i5-3570K} CPU while the parallel version running on the $4$ cores of the same CPU end the work in $50$ minutes.

Using two \tt{i7} based computers and our  parallel algorithm  we computed in two days the values of $n_g$ for $g\leq 60$, confirming the values given by M.Bras-Amor\'os and M.Delgado :

\begin{minipage}{0.5\textwidth}
\begin{tabular}{|r|r|r|}
\hline
g & $n_g$ & $n_g/n_{g-1}$\\
\hline
0 & 1 &  \\
1 & 1 & 1.0 \\
2 & 2 & 2.0 \\
3 & 4 & 2.0 \\
4 & 7 & 1.75 \\
5 & 12 & 1.71428 \\
6 & 23 & 1.91666 \\
7 & 39 & 1.69565 \\
8 & 67 & 1.71794 \\
9 & 118 & 1.76119 \\
10 & 204 & 1.72881 \\
11 & 343 & 1.68137 \\
12 & 592 & 1.72594 \\
13 & 1001 & 1.69087 \\
14 & 1693 & 1.69130 \\
15 & 2857 & 1.68753 \\
16 & 4806 & 1.68218 \\
17 & 8045 & 1.67394 \\
18 & 13467 & 1.67395 \\
19 & 22464 & 1.66807 \\
20 & 37396 & 1.66470 \\
21 & 62194 & 1.66311 \\
22 & 103246 & 1.66006 \\
23 & 170963 & 1.65588 \\
24 & 282828 & 1.65432 \\
25 & 467224 & 1.65197 \\
26 & 770832 & 1.64981 \\
27 & 1270267 & 1.64791 \\
28 & 2091030 & 1.64613 \\
29 & 3437839 & 1.64408 \\
\hline
\end{tabular}
\end{minipage}
\begin{minipage}{0.5\textwidth}
\begin{tabular}{|r|r|r|}
\hline
g & $n_g$ & $n_g/n_{g-1}$\\
\hline
30 & 5646773 & 1.64253 \\
31 & 9266788 & 1.64107 \\
32 & 15195070 & 1.63973 \\
33 & 24896206 & 1.63843 \\
34 & 40761087 & 1.63724 \\
35 & 66687201 & 1.63605 \\
36 & 109032500 & 1.63498 \\
37 & 178158289 & 1.63399 \\
38 & 290939807 & 1.63304 \\
39 & 474851445 & 1.63212 \\
40 & 774614284 & 1.63127 \\
41 & 1262992840 & 1.63047 \\
42 & 2058356522 & 1.62974 \\
43 & 3353191846 & 1.62906 \\
44 & 5460401576 & 1.62841 \\
45 & 8888486816 & 1.62780 \\
46 & 14463633648 & 1.62723 \\
47 & 23527845502 & 1.62668 \\
48 & 38260496374 & 1.62617 \\
49 & 62200036752 & 1.62569 \\
50 & 101090300128 & 1.62524 \\
51 & 164253200784 & 1.62481 \\
52 & 266815155103 & 1.62441 \\
53 & 433317458741 & 1.62403 \\
54 & 703569992121 & 1.62368 \\
55 & 1142140736859 & 1.62335 \\
56 & 1853737832107 & 1.62303 \\
57 & 3008140981820 & 1.62274 \\
58 & 4880606790010 & 1.62246 \\
59 & 7917344087695 & 1.62220 \\
60 & 12841603251351 & 1.62195 \\
\hline
\end{tabular}
\end{minipage}
\vspace{2em}

In \cite{Zhai}, A.Zhai establishes that the limit of the quotient   $\frac{n_g}{n_{g-1}}$, when $g$ go to $+\infty$, is the golden ratio $\phi\approx 1.618$. 
As the reader can see the convergence is very slow.

\begin{figure}[b]
\begin{tikzpicture}[scale=1.3]
        \begin{semilogyaxis}[xlabel=\small genus,ylabel=time,legend pos=north west,legend cell align=left]
        \addplot [color=blue,mark=x] coordinates {
            (20,0.02)
            (21,0.04)
            (22,0.07)
            (23,0.12)
            (24,0.2)
            (25,0.35)
            (26,0.61)
            (27,1.04)
            (28,1.76)
            (29,2.96)
            (30,4.96)
            (31,8.26)
            (32,13.77)
            (33,22.95)
            (34,38.22)
            (35,1250.88)
        };
       \addplot [color=black,mark=x] coordinates {
			(20,0.02)	
            (21,0.04)
            (22,0.06)
            (23,0.09)
            (24,0.16)
            (25,0.26)
            (26,0.45)
            (27,0.72)
            (28,1.25)
            (29,2.1)
            (30,3.39)
            (31,5.85)
            (32,9.25)
            (33,16.34)
            (34,27.23)
            (35,45.39)
            (36,75.58)
            (37,125.59)
            (38,203.56)
            (39,346.02)
            (40,557.51)
        };
         \addplot [color=red,mark=x] coordinates {
			(20,0.02)	
            (21,0.02)
            (22,0.04)
            (23,0.06)
            (24,0.1)
            (25,0.18)
            (26,0.29)
            (27,0.48)
            (28,0.81)
            (29,1.34)
            (30,2.25)
            (31,3.73)
            (32,6.2)
            (33,10.35)
            (34,17.11)
            (35,28.35)
            (36,46.99)
            (37,77.89)
            (38,128.53)
            (39,212.1)
            (40,349.98)
        };
        \addplot [color=magenta,mark=x] coordinates {
			(20,0.01)	
            (21,0.01)
            (22,0.01)
            (23,0.01)
            (24,0.02)
            (25,0.03)
            (26,0.05)
            (27,0.08)
            (28,0.12)
            (29,0.21)
            (30,0.35)
            (31,0.57)
            (32,0.97)
            (33,1.65)
            (34,2.69)
            (35,4.18)
            (36,7.36)
            (37,11.66)
            (38,20.35)
            (39,32.55)
            (40,74.46)
        };
	%\legend style={anchor=north,at={(20,1000)}}  
        \legend{\tiny breadth-expl,\tiny depth-expl,\tiny depth-expl-$\delta$,\tiny depth-expl-$\delta$-mmx}
        \end{semilogyaxis}
    \end{tikzpicture}
    \label{F:Time}
    \caption{Comparison of time executions of different exploration algorithms of the tree $\mathcal{T}^{\leq g}$ on an Intel \tt{i5-3570K} CPU with $8GB$ of memory. 
The scale is logarithmic in time. 
The first version is based on a breadth-first search exploration and the peak at genus~$35$ is due to the consumption of all the memory and use of swap. 
The second version uses a depth first search exploration.
The third is based on the second one with use of decomposition numbers, it corresponds to the algorithm given in Section~\ref{S:Algo}. 
The last one is an optimisation of the third one with MMX.}
\end{figure}

\footnotesize
\bibliographystyle{ieeetr}
\bibliography{biblio}

\end{document}

### 2013-06-04 src/Cilk++/mon.cpp
Working not ridiculous Cilk++ implementation

@@ -1,0 +1,353 @@
+#include <reducer_opadd.h>
+#include <holder.h>
+#include <iostream>
+#include "monoid.hpp"
+
+
+const unsigned long int target_genus = MAX_GENUS;
+
+
+////////////////////////////////////////////
+
+
+#include <stdio.h>
+#include <assert.h>
+#include <x86intrin.h>
+
+#define nth_block(st, i) (((epi8 *) st)[i])
+#define load_epi8(t) ((epi8) _mm_loadu_si128((__m128i *) (t)))
+#define store_epi8(t, v) (_mm_store_si128((__m128i *) (t), (__m128i) (v)))
+#define storeu_epi8(t, v) (_mm_storeu_si128((__m128i *) (t), (__m128i) (v)))
+#define update_epi8(d,s) (d)=(epi8) _mm_sub_epi8((__m128i) (d),  _mm_andnot_si128(_mm_cmpeq_epi8((__m128i) (s), (__m128i)zero),(__m128i) block1));
+
+
+
+const epi8 zero   = {0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0};
+const epi8 block1 = {1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1};
+const unsigned char m1 = 255;
+const epi8 shift16[16] =
+  { { 0, 1, 2, 3, 4, 5, 6, 7, 8, 9,10,11,12,13,14,15},
+    {m1, 0, 1, 2, 3, 4, 5, 6, 7, 8, 9,10,11,12,13,14},
+    {m1,m1, 0, 1, 2, 3, 4, 5, 6, 7, 8, 9,10,11,12,13},
+    {m1,m1,m1, 0, 1, 2, 3, 4, 5, 6, 7, 8, 9,10,11,12},
+    {m1,m1,m1,m1, 0, 1, 2, 3, 4, 5, 6, 7, 8, 9,10,11},
+    {m1,m1,m1,m1,m1, 0, 1, 2, 3, 4, 5, 6, 7, 8, 9,10},
+    {m1,m1,m1,m1,m1,m1, 0, 1, 2, 3, 4, 5, 6, 7, 8, 9},
+    {m1,m1,m1,m1,m1,m1,m1, 0, 1, 2, 3, 4, 5, 6, 7, 8},
[…]

### 2013-06-04 src/Cilk++/mon.cpp
Descent Cilk++ perfs

@@ -1,8 +1,42 @@
 #include <reducer_opadd.h>
 #include <holder.h>
 #include <iostream>
-#include "monoid.hpp"
 
+using namespace std;
+typedef unsigned char epi8 __attribute__ ((vector_size (16)));
+
+#define MAX_GENUS 40
+#define SIZE_BOUND (3*(MAX_GENUS-1))
+#define NBLOCKS ((SIZE_BOUND+15) >> 4)
+#define SIZE (NBLOCKS << 4)
+
+
+typedef unsigned char nb_decompositions[SIZE] __attribute__ ((aligned (16)));
+typedef unsigned long long int result[MAX_GENUS];
+
+struct monoid
+{
+  nb_decompositions decs;
+  unsigned long int conductor, min, genus;
+};
+
+struct generator_iter
+{
+  unsigned long int iblock, mask, gen, bound;
+};
+
+void init_full_N(monoid &);
+void print_monoid(const monoid &);
+void print_epi8(epi8);
+inline void copy_decs(      nb_decompositions &__restrict__ dst,
+		      const nb_decompositions &__restrict__ src);
+inline monoid remove_generator(const monoid &__restrict__, unsigned long int);
+
+
+inline void init_all_generator_iter(const monoid &, generator_iter &);
+inline generator_iter init_children_generator_iter(const monoid &m);
+inline unsigned long int next_generator_iter(const monoid &, generator_iter &);
+inline unsigned char count_generator_iter(const monoid &, generator_iter &);
 
 const unsigned long int target_genus = MAX_GENUS;
 
@@ -60,16 +94,12 @@ const epi8 mask16[16] =
     { 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0,m1,m1},
     { 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0,m1} };
 
-monoid *alloc_monoid()
-{
-  return new monoid;
-}
 
-void print_monoid(monoid *pm)
+void print_monoid(const monoid &m)
 {
   unsigned int i;
-  cout<<"min = "<<pm->min<<", cond = "<<pm->conductor<<", genus = "<<pm->genus<<", decs = ";
-  for (i=0; i<SIZE; i++) cout<<((int) pm->decs[i])<<' ';
+  cout<<"min = "<<m.min<<", cond = "<<m.conductor<<", genus = "<<m.genus<<", decs = ";
[…]

### 2013-06-04 src/Cilk++/mon.cpp
Better C++ iteration

@@ -20,11 +20,6 @@ struct monoid
   unsigned long int conductor, min, genus;
 };
 
-struct generator_iter
-{
-  unsigned long int iblock, mask, gen, bound;
-};
-
 void init_full_N(monoid &);
 void print_monoid(const monoid &);
 void print_epi8(epi8);
@@ -32,11 +27,6 @@ inline void copy_decs(      nb_decompositions &__restrict__ dst,
 		      const nb_decompositions &__restrict__ src);
 inline monoid remove_generator(const monoid &__restrict__, unsigned long int);
 
-
-inline void init_all_generator_iter(const monoid &, generator_iter &);
-inline generator_iter init_children_generator_iter(const monoid &m);
-inline unsigned long int next_generator_iter(const monoid &, generator_iter &);
-inline unsigned char count_generator_iter(const monoid &, generator_iter &);
 
 const unsigned long int target_genus = MAX_GENUS;
 
@@ -94,6 +84,73 @@ const epi8 mask16[16] =
     { 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0,m1,m1},
     { 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0,m1} };
 
+class generator_iter
+{
+  const monoid &m;
+  unsigned long int iblock, mask, gen, bound;
+
+  generator_iter(const monoid &mon) : m(mon), bound((mon.conductor+mon.min+15) >> 4) {};
+
+    public:
+  static inline generator_iter all(const monoid &m)
+  {
+    epi8 block;
+    generator_iter it(m);
+    it.iblock = 0;
+    block = nth_block(m.decs, 0);
+    ((unsigned char*)(&block))[0]=0; // 0 is not a generator
+    it.mask  = _mm_movemask_epi8((__m128i) _mm_cmpeq_epi8((__m128i) block, (__m128i) block1));
+    it.gen = - 1;
+    it.iblock++;
+    return it;
+  };
+
+  static inline generator_iter children(const monoid &m)
+  {
+    generator_iter it(m);
+    epi8 block;
+    it.iblock = m.conductor >> 4;
+    block = nth_block(m.decs, it.iblock) & mask16[m.conductor & 0xF];
+    it.mask  = _mm_movemask_epi8(_mm_cmpeq_epi8((__m128i) block, (__m128i) block1));
[…]

### 2013-06-04 src/Cilk++/monoid.cpp

@@ -1,284 +1,0 @@
-#include <stdio.h>
-#include <assert.h>
-#include <x86intrin.h>
-#include "monoid.hpp"
-
-#define nth_block(st, i) (((epi8 *) st)[i])
-#define load_epi8(t) ((epi8) _mm_loadu_si128((__m128i *) (t)))
-#define store_epi8(t, v) (_mm_store_si128((__m128i *) (t), (__m128i) (v)))
-#define storeu_epi8(t, v) (_mm_storeu_si128((__m128i *) (t), (__m128i) (v)))
-#define update_epi8(d,s) (d)=(epi8) _mm_sub_epi8((__m128i) (d),  _mm_andnot_si128(_mm_cmpeq_epi8((__m128i) (s), (__m128i)zero),(__m128i) block1));
-
-
-
-epi8 zero   = {0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0};
-epi8 block1 = {1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1};
-epi8 shift16[16] =
-  { { 0, 1, 2, 3, 4, 5, 6, 7, 8, 9,10,11,12,13,14,15},
-    {-1, 0, 1, 2, 3, 4, 5, 6, 7, 8, 9,10,11,12,13,14},
-    {-1,-1, 0, 1, 2, 3, 4, 5, 6, 7, 8, 9,10,11,12,13},
-    {-1,-1,-1, 0, 1, 2, 3, 4, 5, 6, 7, 8, 9,10,11,12},
-    {-1,-1,-1,-1, 0, 1, 2, 3, 4, 5, 6, 7, 8, 9,10,11},
-    {-1,-1,-1,-1,-1, 0, 1, 2, 3, 4, 5, 6, 7, 8, 9,10},
-    {-1,-1,-1,-1,-1,-1, 0, 1, 2, 3, 4, 5, 6, 7, 8, 9},
-    {-1,-1,-1,-1,-1,-1,-1, 0, 1, 2, 3, 4, 5, 6, 7, 8},
-    {-1,-1,-1,-1,-1,-1,-1,-1, 0, 1, 2, 3, 4, 5, 6, 7},
-    {-1,-1,-1,-1,-1,-1,-1,-1,-1, 0, 1, 2, 3, 4, 5, 6},
[…]

### 2013-06-04 src/Cilk++/monoid.h

@@ -1,4 +1,0 @@
-extern "C"
-{
-  #include "../monoid.h"
-}

### 2013-06-05 src/Cilk++/mon.cpp
Fast cilk version using an array reducer

@@ -12,7 +12,6 @@ typedef unsigned char epi8 __attribute__ ((vector_size (16)));
 
 
 typedef unsigned char nb_decompositions[SIZE] __attribute__ ((aligned (16)));
-typedef unsigned long long int result[MAX_GENUS];
 
 struct monoid
 {
@@ -148,9 +147,37 @@ class generator_iter
     return nbr;
   };
 
-  inline bool const is_finished() { return  (iblock > bound); };
-  inline unsigned long int const get_gen() {return gen;}
+  inline bool is_finished() const { return  (iblock > bound); };
+  inline unsigned long int get_gen() const {return gen; };
 };
+
+extern "C++"
+{
+  struct Results
+  {
+    unsigned long int values[MAX_GENUS];
+    Results() {for(int i=0; i<MAX_GENUS; i++) values[i] = 0;};
+  };
+
+  class ResultsReducer
+  {
+    struct Monoid: cilk::monoid_base<Results>
+    {
+      static void reduce (Results *left, Results *right){
+	for(int i=0; i<MAX_GENUS; i++) left->values[i] += right->values[i];
+      };
+    };
+  private:
+    cilk::reducer<Monoid> imp_;
+  public:
+    ResultsReducer() : imp_() {};
+    inline unsigned long int & operator[](int i) {
+      return imp_.view().values[i];
+    };
+    inline Results &get_array() {return imp_.view();}
+  };
+};
+
 
 void print_monoid(const monoid &m)
 {
@@ -289,39 +316,41 @@ void init_full_N(monoid &m)
 }
 
 
-cilk::reducer_opadd<unsigned long int> cilk_result[target_genus];
+ResultsReducer cilk_results;
 
+#define STACK_SIZE 100
 inline void walk_children_stack(const monoid &m, unsigned long int bound)
 {
   unsigned long int stack_pointer = 1, nbr;
-  monoid current, stack[100];
+  monoid data[STACK_SIZE-1], *stack[STACK_SIZE], *tmp;
+  Results & res = cilk_results.get_array();
 
-  stack[0] = m;
+  for (int i=1; i<STACK_SIZE; i++) stack[i] = &(data[i-1]); // Nathann's trick to avoid copy
+  stack[0] = & ((monoid &)m);
   while (stack_pointer)
     {
       --stack_pointer;
-      if (stack[stack_pointer].genus < bound - 1)
+      if (stack[stack_pointer]->genus < bound - 1)
 	{
 	  nbr = 0;
-
[…]

### 2013-06-05 src/Cilk++/mon.cpp
Cycling the stack

@@ -318,11 +318,11 @@ void init_full_N(monoid &m)
 
 ResultsReducer cilk_results;
 
-#define STACK_SIZE 100
+#define STACK_SIZE 40
 inline void walk_children_stack(const monoid &m, unsigned long int bound)
 {
   unsigned long int stack_pointer = 1, nbr;
-  monoid data[STACK_SIZE-1], *stack[STACK_SIZE], *tmp;
+  monoid data[STACK_SIZE-1], *stack[STACK_SIZE], *current;
   Results & res = cilk_results.get_array();
 
   for (int i=1; i<STACK_SIZE; i++) stack[i] = &(data[i-1]); // Nathann's trick to avoid copy
@@ -330,39 +330,39 @@ inline void walk_children_stack(const monoid &m, unsigned long int bound)
   while (stack_pointer)
     {
       --stack_pointer;
-      if (stack[stack_pointer]->genus < bound - 1)
+      current = stack[stack_pointer];
+      if (current->genus < bound - 1)
 	{
 	  nbr = 0;
-	  for (auto it = generator_iter::children(*stack[stack_pointer]).next();
+	  for (auto it = generator_iter::children(*current).next();
 	       not it.is_finished();
 	       it.next())
 	    {
 	      // exchange top with top+1
-	      tmp = stack[stack_pointer];
 	      stack[stack_pointer] = stack[stack_pointer+1];
-	      stack[stack_pointer+1] = tmp;
-	      remove_generator(*stack[stack_pointer], *stack[stack_pointer+1], it.get_gen());
+	      remove_generator(*stack[stack_pointer], *current, it.get_gen());
 	      stack_pointer++;
 	      nbr++;
 	    }
-	  res.values[stack[stack_pointer]->genus] += nbr;
+	  stack[stack_pointer] = current;
+	  res.values[current->genus] += nbr;
 	}
       else
 	{
-	  res.values[stack[stack_pointer]->genus] +=
-	    generator_iter::children(*stack[stack_pointer]).count();
+	  res.values[current->genus] +=
+	    generator_iter::children(*current).count();
 	}
     }
 }
 
 
 ////////////////////////////////////////////
-
+const int stack_bound = 11;
 void walk_children(const monoid &m)
 {
   unsigned long int nbr = 0;
 
-  if (m.genus < target_genus - 10)
+  if (m.genus < target_genus - stack_bound)
     {
       for (auto it = generator_iter::children(m).next();
 	   not it.is_finished();

### 2013-06-05 src/Cilk++/mon.cpp
Cleanup in interator interface

@@ -91,20 +91,20 @@ class generator_iter
   generator_iter(const monoid &mon) : m(mon), bound((mon.conductor+mon.min+15) >> 4) {};
 
     public:
-  static inline generator_iter all(const monoid &m)
+  static inline generator_iter all_count(const monoid &m)
   {
     epi8 block;
     generator_iter it(m);
     it.iblock = 0;
     block = nth_block(m.decs, 0);
-    ((unsigned char*)(&block))[0]=0; // 0 is not a generator
     it.mask  = _mm_movemask_epi8((__m128i) _mm_cmpeq_epi8((__m128i) block, (__m128i) block1));
+    it.mask &= 0xFE; // 0 is not a generator
     it.gen = - 1;
     it.iblock++;
     return it;
   };
 
-  static inline generator_iter children(const monoid &m)
+  static inline generator_iter children_count(const monoid &m)
   {
     generator_iter it(m);
     epi8 block;
@@ -116,7 +116,12 @@ class generator_iter
     return it;
   };
 
-  inline generator_iter next()
+  static inline generator_iter all(const monoid &m)
+  { return ++generator_iter::all_count(m); };
+  static inline generator_iter children(const monoid &m)
+  { return ++generator_iter::children_count(m); };
+
+  inline generator_iter operator++()
   {
     unsigned long int shift;
     epi8 block;
@@ -318,15 +323,15 @@ void init_full_N(monoid &m)
 
 ResultsReducer cilk_results;
 
-#define STACK_SIZE 40
-inline void walk_children_stack(const monoid &m, unsigned long int bound)
+#define STACK_SIZE 50
+inline void walk_children_stack(monoid m, unsigned long int bound)
 {
   unsigned long int stack_pointer = 1, nbr;
   monoid data[STACK_SIZE-1], *stack[STACK_SIZE], *current;
   Results & res = cilk_results.get_array();
 
   for (int i=1; i<STACK_SIZE; i++) stack[i] = &(data[i-1]); // Nathann's trick to avoid copy
-  stack[0] = & ((monoid &)m);
+  stack[0] = &m;
   while (stack_pointer)
     {
       --stack_pointer;
@@ -334,9 +339,9 @@ inline void walk_children_stack(const monoid &m, unsigned long int bound)
       if (current->genus < bound - 1)
 	{
 	  nbr = 0;
-	  for (auto it = generator_iter::children(*current).next();
+	  for (auto it = generator_iter::children(*current);
[…]

### 2013-06-05 src/Cilk++/mon.cpp
Back to epi8 as they seems faster + cleanup

@@ -3,6 +3,19 @@
 #include <iostream>
 
 using namespace std;
+
+
+/*
+Note: for some reason the code using gcc vector variables is 2-3% faster than
+the code using gcc intrinsic instructions.
+Here are the results of two runs:
+data_mmx =  [9.757, 9.774, 9.757, 9.761, 9.811, 9.819, 9.765, 9.888, 9.774, 9.771]
+data_epi8 = [9.592, 9.535, 9.657, 9.468, 9.460, 9.679, 9.458, 9.461, 9.629, 9.474]
+*/
+
+extern "C++"
+{
+
 typedef unsigned char epi8 __attribute__ ((vector_size (16)));
 
 #define MAX_GENUS 40
@@ -38,10 +51,7 @@ const unsigned long int target_genus = MAX_GENUS;
 #include <x86intrin.h>
 
 #define nth_block(st, i) (((epi8 *) st)[i])
-#define load_epi8(t) ((epi8) _mm_loadu_si128((__m128i *) (t)))
-#define store_epi8(t, v) (_mm_store_si128((__m128i *) (t), (__m128i) (v)))
-#define storeu_epi8(t, v) (_mm_storeu_si128((__m128i *) (t), (__m128i) (v)))
-#define update_epi8(d,s) (d)=(epi8) _mm_sub_epi8((__m128i) (d),  _mm_andnot_si128(_mm_cmpeq_epi8((__m128i) (s), (__m128i)zero),(__m128i) block1));
+#define load_unaligned_epi8(t) ((epi8) _mm_loadu_si128((__m128i *) (t)))
 
 
 
@@ -97,7 +107,7 @@ class generator_iter
     generator_iter it(m);
     it.iblock = 0;
     block = nth_block(m.decs, 0);
-    it.mask  = _mm_movemask_epi8((__m128i) _mm_cmpeq_epi8((__m128i) block, (__m128i) block1));
+    it.mask  = _mm_movemask_epi8((__m128i) (block == block1));
     it.mask &= 0xFE; // 0 is not a generator
     it.gen = - 1;
     it.iblock++;
@@ -110,7 +120,7 @@ class generator_iter
     epi8 block;
     it.iblock = m.conductor >> 4;
     block = nth_block(m.decs, it.iblock) & mask16[m.conductor & 0xF];
-    it.mask  = _mm_movemask_epi8(_mm_cmpeq_epi8((__m128i) block, (__m128i) block1));
[…]

### 2013-06-05 src/Cilk++/mon.cpp
Cleanup the block access

@@ -15,29 +15,33 @@ data_epi8 = [9.592, 9.535, 9.657, 9.468, 9.460, 9.679, 9.458, 9.461, 9.629, 9.47
 
 extern "C++"
 {
-
-typedef unsigned char epi8 __attribute__ ((vector_size (16)));
-
 #define MAX_GENUS 40
 #define SIZE_BOUND (3*(MAX_GENUS-1))
 #define NBLOCKS ((SIZE_BOUND+15) >> 4)
 #define SIZE (NBLOCKS << 4)
 
-
-typedef unsigned char nb_decompositions[SIZE] __attribute__ ((aligned (16)));
+typedef unsigned char epi8 __attribute__ ((vector_size (16)));
+typedef unsigned char dec_numbers[SIZE] __attribute__ ((aligned (16)));
+typedef epi8 dec_blocks[NBLOCKS];
 
 struct monoid
 {
-  nb_decompositions decs;
+  union {
+    dec_numbers decs;
+    dec_blocks blocks;
+  };
   unsigned long int conductor, min, genus;
 };
 
 void init_full_N(monoid &);
 void print_monoid(const monoid &);
 void print_epi8(epi8);
-inline void copy_decs(      nb_decompositions &__restrict__ dst,
-		      const nb_decompositions &__restrict__ src);
-inline monoid remove_generator(const monoid &__restrict__, unsigned long int);
+inline void copy_blocks(      dec_blocks &__restrict__ dst,
+			const dec_blocks &__restrict__ src);
+inline void remove_generator(monoid &__restrict__ dst,
+			     const monoid &__restrict__ src,
+			     unsigned long int gen)  __attribute__((always_inline));
+monoid remove_generator(const monoid &src, unsigned long int gen);
 
 
 const unsigned long int target_genus = MAX_GENUS;
@@ -50,13 +54,17 @@ const unsigned long int target_genus = MAX_GENUS;
 #include <assert.h>
 #include <x86intrin.h>
 
-#define nth_block(st, i) (((epi8 *) st)[i])
-#define load_unaligned_epi8(t) ((epi8) _mm_loadu_si128((__m128i *) (t)))
+extern inline epi8 load_unaligned_epi8(const unsigned char *t)
+{ return (epi8) _mm_loadu_si128((__m128i *) (t)); };
 
+extern inline epi8 shuffle_epi8(epi8 b, epi8 sh)
[…]

### 2013-06-06 src/Cilk++/monoid.hpp
Cleanup un the cilk++ code

@@ -1,9 +1,20 @@
-#include <reducer_opadd.h>
-#include <holder.h>
+#ifndef MONOID_HPP
+#define MONOID_HPP
+
 #include <iostream>
 
 using namespace std;
 
+#define MAX_GENUS 40
+#define SIZE_BOUND (3*(MAX_GENUS-1))
+#define NBLOCKS ((SIZE_BOUND+15) >> 4)
+#define SIZE (NBLOCKS << 4)
+
+#include <x86intrin.h>
+
+typedef unsigned char epi8 __attribute__ ((vector_size (16)));
+typedef unsigned char dec_numbers[SIZE] __attribute__ ((aligned (16)));
+typedef epi8 dec_blocks[NBLOCKS];
 
 /*
 Note: for some reason the code using gcc vector variables is 2-3% faster than
@@ -12,47 +23,6 @@ Here are the results of two runs:
 data_mmx =  [9.757, 9.774, 9.757, 9.761, 9.811, 9.819, 9.765, 9.888, 9.774, 9.771]
 data_epi8 = [9.592, 9.535, 9.657, 9.468, 9.460, 9.679, 9.458, 9.461, 9.629, 9.474]
 */
-
-extern "C++"
-{
-#define MAX_GENUS 40
-#define SIZE_BOUND (3*(MAX_GENUS-1))
-#define NBLOCKS ((SIZE_BOUND+15) >> 4)
-#define SIZE (NBLOCKS << 4)
-
-typedef unsigned char epi8 __attribute__ ((vector_size (16)));
-typedef unsigned char dec_numbers[SIZE] __attribute__ ((aligned (16)));
-typedef epi8 dec_blocks[NBLOCKS];
-
-struct monoid
-{
-  union {
-    dec_numbers decs;
-    dec_blocks blocks;
-  };
-  unsigned long int conductor, min, genus;
-};
-
-void init_full_N(monoid &);
-void print_monoid(const monoid &);
-void print_epi8(epi8);
-inline void copy_blocks(      dec_blocks &__restrict__ dst,
-			const dec_blocks &__restrict__ src);
-inline void remove_generator(monoid &__restrict__ dst,
-			     const monoid &__restrict__ src,
-			     unsigned long int gen)  __attribute__((always_inline));
-monoid remove_generator(const monoid &src, unsigned long int gen);
-
-
-const unsigned long int target_genus = MAX_GENUS;
-
-
-////////////////////////////////////////////
-
-
[…]

### 2013-06-06 src/Cilk++/treewalk.cpp

@@ -1,0 +1,112 @@
+#include <cilk/reducer.h>
+#include <iostream>
+
+using namespace std;
+
+#include "monoid.hpp"
+
+struct Results
+{
+  typedef unsigned long int type[MAX_GENUS];
+  type values;
+  Results() {for(int i=0; i<MAX_GENUS; i++) values[i] = 0;};
+};
+
+class ResultsReducer
+{
+  struct Monoid: cilk::monoid_base<Results>
+  {
+    static void reduce (Results *left, Results *right){
+      for(int i=0; i<MAX_GENUS; i++) left->values[i] += right->values[i];
+    };
+  };
+private:
+  cilk::reducer<Monoid> imp_;
+public:
+  ResultsReducer() : imp_() {};
+  inline unsigned long int & operator[](int i) {
+    return imp_.view().values[i];
+  };
+  inline Results::type &get_array() {return imp_.view().values;}
+};
+
+
+#define STACK_SIZE 50
+void walk_children_stack(monoid m, unsigned long int bound, Results::type &res)
+{
+  unsigned long int stack_pointer = 1, nbr;
+  monoid data[STACK_SIZE-1], *stack[STACK_SIZE], *current;
+
+  for (int i=1; i<STACK_SIZE; i++) stack[i] = &(data[i-1]); // Nathann's trick to avoid copy
+  stack[0] = &m;
+  while (stack_pointer)
+    {
+      --stack_pointer;
+      current = stack[stack_pointer];
+      if (current->genus < bound - 1)
+	{
+	  nbr = 0;
+	  for (auto it = generator_iter::children(*current);
+	       not it.is_finished();
+	       ++it)
+	    {
+	      // exchange top with top+1
+	      stack[stack_pointer] = stack[stack_pointer+1];
+	      remove_generator(*stack[stack_pointer], *current, it.get_gen());
+	      stack_pointer++;
+	      nbr++;
+	    }
+	  stack[stack_pointer] = current;
+	  res[current->genus] += nbr;
+	}
+      else
+	{
+	  res[current->genus] +=
+	    generator_iter::children_count(*current).count();
+	}
+    }
+}
+
+ResultsReducer cilk_results;
+////////////////////////////////////////////
+#define STACK_BOUND 11
+void walk_children(const monoid &m)
+{
+  unsigned long int nbr = 0;
+
+  if (m.genus < target_genus - STACK_BOUND)
+    {
[…]

### 2013-06-06 src/Cilk++/treewalk.cpp
Remove unneeded spawn in main

@@ -100,8 +100,7 @@ int main(void)
 
   init_full_N(N);
   cilk_results[0]++;
-  cilk_spawn walk_children(N);
-  cilk_sync;
+  walk_children(N);
 
   std::cout << std::endl << "============================" << std::endl << std::endl;
   for (unsigned int i=0; i<target_genus; i++)

### 2013-06-06 src/Cilk++/treewalk.cpp
Corrected genus 0 result

@@ -99,7 +99,6 @@ int main(void)
 	    << target_genus << std::endl;
 
   init_full_N(N);
-  cilk_results[0]++;
   walk_children(N);
 
   std::cout << std::endl << "============================" << std::endl << std::endl;

### 2013-06-06 src/Cilk++/treewalk.cpp
Adding setting of nworkers

@@ -95,9 +95,11 @@ int main(void)
 {
   monoid N;
 
-  std::cout << "Computing number of numeric monoids for genus <= "
-	    << target_genus << std::endl;
+  if (__cilkrts_set_param("nworkers", "0") != __CILKRTS_SET_PARAM_SUCCESS)
+    std::cerr << "Failed to set the number of workers" << endl;
 
+  std::cout << "Computing number of numeric monoids for genus <= "
+	    << target_genus << " using " << __cilkrts_get_nworkers() << " workers" << endl;
   init_full_N(N);
   walk_children(N);
 

### 2013-06-07 src/Cilk++/treewalk.cpp
Removed unusefull std

@@ -96,17 +96,17 @@ int main(void)
   monoid N;
 
   if (__cilkrts_set_param("nworkers", "0") != __CILKRTS_SET_PARAM_SUCCESS)
-    std::cerr << "Failed to set the number of workers" << endl;
+    cerr << "Failed to set the number of workers" << endl;
 
-  std::cout << "Computing number of numeric monoids for genus <= "
+  cout << "Computing number of numeric monoids for genus <= "
 	    << target_genus << " using " << __cilkrts_get_nworkers() << " workers" << endl;
   init_full_N(N);
   walk_children(N);
 
-  std::cout << std::endl << "============================" << std::endl << std::endl;
+  cout << endl << "============================" << endl << endl;
   for (unsigned int i=0; i<target_genus; i++)
-    std::cout << cilk_results[i] << " ";
-  std::cout << std::endl;
+    cout << cilk_results[i] << " ";
+  cout << endl;
   return EXIT_SUCCESS;
 }
 

### 2013-06-07 src/Cilk++/treewalk.cpp
Fixed stack size

@@ -30,14 +30,12 @@ public:
   inline Results::type &get_array() {return imp_.view().values;}
 };
 
-
-#define STACK_SIZE 50
 void walk_children_stack(monoid m, unsigned long int bound, Results::type &res)
 {
   unsigned long int stack_pointer = 1, nbr;
-  monoid data[STACK_SIZE-1], *stack[STACK_SIZE], *current;
+  monoid data[bound-1], *stack[bound], *current;
 
-  for (int i=1; i<STACK_SIZE; i++) stack[i] = &(data[i-1]); // Nathann's trick to avoid copy
+  for (unsigned long int i=1; i<bound; i++) stack[i] = &(data[i-1]); // Nathann's trick to avoid copy
   stack[0] = &m;
   while (stack_pointer)
     {

### 2013-10-07 Papier/article.tex
Rereading + more computed values

@@
 
-$iii)$ $c(S)=1+\max(\NN\setminus S)$, the \emph{conductor} of $S$ for $S$ different from $\NN$. By convention the conductor of $\NN$ is $0$. 
+$iii)$ $c(S)=1+\max(\NN\setminus S)$, the \emph{conductor} of $S$ for $S$ different from $\NN$. By convention, the conductor of $\NN$ is $0$. 
 \end{defi}
@@
 If~$S$ is a semigroup in the tree,  the sons of $S$ are exactly the semigroup $S^x$ where $x$ belongs to $\Irr(S)\cap[c(s),+\infty[$. 
-By convention, when depicting the tree, the ,numerical semigroup $S^x$ is in the left of $S^y$ if $x$ is smaller than $y$.
+By convention, when depicting the tree, the numerical semigroup $S^x$ is in the left of $S^y$ if $x$ is smaller than $y$.
 
@@
 Since $\psi$ is injective we must have $k\leq g(S)$. 
->From the relation 
+From the relation 
 \[
@@
 The main advantage in our approach is the small memory needs. 
-The cost to pay is that, if we want to explore the tree deepero, we must restart from the root.
+The cost to pay is that, if we want to explore the tree deeper, we must restart from the root.
 
@@
 29 & 3437839 & 1.64408 \\
+30 & 5646773 & 1.64253 \\
+31 & 9266788 & 1.64107 \\
+32 & 15195070 & 1.63973 \\
+33 & 24896206 & 1.63843 \\
+34 & 40761087 & 1.63724 \\
+35 & 66687201 & 1.63605 \\
 \hline
@@
 \hline
-30 & 5646773 & 1.64253 \\
-31 & 9266788 & 1.64107 \\
-32 & 15195070 & 1.63973 \\
-33 & 24896206 & 1.63843 \\
-34 & 40761087 & 1.63724 \\
-35 & 66687201 & 1.63605 \\
 36 & 109032500 & 1.63498 \\
@@
 60 & 12841603251351 & 1.62195 \\
-\hline
+61 & 20825558002053 & 1.62173 \\
+62 & 33768763536686 & 1.62151 \\
+63 & 54749244915730 & 1.62130 \\
+64 & 88754191073328 & 1.62110 \\
+65 & 143863484925550 & 1.62092 \\
+66 & 233166577125714 & 1.62075 \\
+67 & 377866907506273 & 1.62059 \\\hline
 \end{tabular}

### 2013-10-08 Papier/article.tex
Florent's rant

@@
 \section{Technical optimizations and results}
+
+Though there are asymptotically faster algorithm than the one presented here,
+thank to careful optimizations, we were able to compute $n_g$ for much larger
+genus than before. This is due to the fact that our algorithm is very well
+suited for the current processor architecture. In particular, it allows to use
+parallelism at various scale (parallel branch exploration, vectorization)...
+
+\subsection{Florent's rant}
+
+We use the following optimization tricks
+\begin{itemize}
+\item Vectorisation (MMX, SSE, GCC's 
+\verb|__attribute__ ((vector_size(16)));|
+) and careful memory alignment.
+\item Shared memory multicore computing using \texttt{CILK++} for low level
+  enumerating tree branching;
+\item Un-recursived algorithm using a stack;
+\item Avoiding all dynamic allocation during the computation: everything is
+  computed ``in place'';
+\item Avoiding all unnecessary copy (the cost of the \textsc{Son} algorithm is
+  roughly the same (even less) as for as copy !
+\item Aggressive loop unrolling. The main loop is unrolled by hand using some
+  kind of Duff's device.
+\item Carefull choice of data type (\verb|uint_fast8_t| for decomposition
+  number, vs \verb|uint_fast64_t| for all indexes).
+\end{itemize}
+
 \label{S:Opti}

### 2013-10-08 Papier/article.tex
Rewrote figure using tikz

@@
 \usepackage{url}
-\usepackage{pst-tree}
 \usepackage{algorithm}
@@
 \renewcommand{\tt}[1]{\texttt{#1}}
-\newcommand{\sgnode}[1]{\TR{\bf \left<#1\right>}}
+\newcommand{\sgnode}[1]{{\bf \left<#1\right>}}
 \renewcommand{\ie}{\emph{i.e}}
@@
  \begin{figure}[htf!]
+ % \[
+ % \pstree[,nodesep=3pt]{\sgnode{1}}{
+ % 	\pstree[nodesep=3pt]{\sgnode{2,3} \tlput1}{
+ % 		\pstree[treesep=4pt,nodesep=2pt]{\sgnode{3,4,5} \tlput2}{
+ % 			\pstree[treesep=3pt,nodesep=3pt]{\sgnode{4,5,6,7} \tlput3}{
+ % 				\pstree{\sgnode{5,6,7,8,9} \tlput4}{}
+ % 				\pstree{\sgnode{\gr 4,6,7,9} \tlput5}{}
+ % 				\pstree{\sgnode{\gr 4,\gr 5,7} \trput6}{}
+ % 				\pstree{\sgnode{\gr 4,\gr 5,\gr 6} \trput7}{}
+ % 			}
+ % 			\pstree[treesep=3pt,nodesep=3pt]{\sgnode{\gr 3,5,7} \tlput4}{
+ % 				\pstree{\sgnode{\gr 3,7,8} \tlput 5}{}
+ % 				\pstree{\sgnode{\gr 3,\gr 5} \trput 7}{}
+ % 			}
+ % 			\pstree[treesep=10pt]{\sgnode{\gr 3,\gr 4} \trput5}{\Tn \Tn \Tn}
+ % 		}
+ % 		\pstree[nodesep=3pt]{~\hspace{-5em}\sgnode{\gr2,5} \trput3}{
+ % 			\pstree[nodesep=3pt]{\sgnode{\gr2,7} \trput5}{
+ % 				\pstree[nodesep=3pt]{\sgnode{\gr2,9} \trput7}{}
+ % 			}
+ % 		}
+ % 	}
+ % }
+ % \vspace{-5em}
+ % \]
+
 \[
- \pstree[,nodesep=3pt]{\sgnode{1}}{
- 	\pstree[nodesep=3pt]{\sgnode{2,3} \tlput1}{
- 		\pstree[treesep=4pt,nodesep=2pt]{\sgnode{3,4,5} \tlput2}{
- 			\pstree[treesep=3pt,nodesep=3pt]{\sgnode{4,5,6,7} \tlput3}{
- 				\pstree{\sgnode{5,6,7,8,9} \tlput4}{}
- 				\pstree{\sgnode{\gr 4,6,7,9} \tlput5}{}
- 				\pstree{\sgnode{\gr 4,\gr 5,7} \trput6}{}
- 				\pstree{\sgnode{\gr 4,\gr 5,\gr 6} \trput7}{}					
+\begin{tikzpicture}[level distance=2cm, inner sep=2mm,
+  level/.style={sibling distance=1.55cm},
+%  edge from parent/.style={draw}
+  ]
+    \node {$\sgnode{1}$}
+    child {node {$\sgnode{2,3}$}
+      child [sibling distance=4.5cm] {node {$\sgnode{3,4,5}$}
+        child {node {$\sgnode{4,5,6,7}$}
+          child [sibling distance=2cm]{node {$\sgnode{5,6,7,8,9}$}  edge from parent node [left] {4}}
+          child [sibling distance=1.9cm]{node {$\sgnode{\gr 4,6,7,9}$}  edge from parent node [left] {5}}
+          child {node {$\sgnode{\gr 4,\gr 5,7}$}  edge from parent node [right] {6}}
+          child {node {$\sgnode{\gr 4,\gr 5,\gr 6}$}  edge from parent node [right] {7}}
+          edge from parent node [left] {3}
         }
- 			\pstree[treesep=3pt,nodesep=3pt]{\sgnode{\gr 3,5,7} \tlput4}{
- 				\pstree{\sgnode{\gr 3,7,8} \tlput 5}{}
- 				\pstree{\sgnode{\gr 3,\gr 5} \trput 7}{}
+        child{{} edge from parent[draw=none]}
+        child{{} edge from parent[draw=none]}
+        child {node {$\sgnode{\gr 3,5,7}$}
+          child {node {$\sgnode{\gr 3,7,8}$} edge from parent node [left] {5}}
+          child {node {$\sgnode{\gr 3,\gr 5}$} edge from parent node [right] {7}}
+          edge from parent node [left] {4}
         }
- 			\pstree[treesep=10pt]{\sgnode{\gr 3,\gr 4} \trput5}{\Tn \Tn \Tn}
+        child {node {$\sgnode{\gr 3,\gr 4}$}
+          edge from parent node [right] {5}
         }
- 		\pstree[nodesep=3pt]{~\hspace{-5em}\sgnode{\gr2,5} \trput3}{
- 			\pstree[nodesep=3pt]{\sgnode{\gr2,7} \trput5}{
- 				\pstree[nodesep=3pt]{\sgnode{\gr2,9} \trput7}{}
+        edge from parent node [left] {2}
       }
+      child [sibling distance=4.5cm] {node {$\sgnode{\gr2,5}$}
+        child {node {$\sgnode{\gr2,7}$}
+          child {node {$\sgnode{\gr2,9}$}
+            edge from parent node [right] {7}
           }
+          edge from parent node [right] {5}
         }
+        edge from parent node [right] {3}
       }
- \vspace{-5em}
+      edge from parent node [left] {1}
+    }
+    ;
+  \end{tikzpicture}
   \]

### 2013-10-11 src/Cilk++/setup.py
Fix Cilk++/g++ on Darwin

@@ -1,7 +1,7 @@
 # call with
 # sage -python setup.py build_ext --inplace
 
-import os, sys
+import os, sys, platform
 
 try:
     CILK_ROOT = os.environ['CILK_ROOT']
@@ -9,6 +9,7 @@ except KeyError:
     raise EnvironmentError, "Please define the CILK_ROOT environment variable !"
 # setup the gcc/cilk compiler
 os.environ['CC'] = os.path.join(CILK_ROOT, 'bin', 'g++')
+
 
 from distutils.core import setup
 from distutils.extension import Extension
@@ -19,8 +20,18 @@ from sage.env import *
 SAGE_INC = os.path.join(SAGE_LOCAL, 'include')
 SAGE_C   = os.path.join(SAGE_SRC, 'c_lib', 'include')
 SAGE_DEV = os.path.join(SAGE_ROOT, 'devel', 'sage-main')
-CILK_LIB = os.path.join(CILK_ROOT, 'lib64')
 
+
+if platform.system()=="Darwin":
+    MARCH='-march=corei7'
+    MTUNE='-march=corei7'   
+    CILK_LIB = os.path.join(CILK_ROOT, 'lib') 
+    
+else:
+    MARCH='-march=native'
+    MTUNE='-march=native'
+    CILK_LIB = os.path.join(CILK_ROOT, 'lib64')
+ 
 import Cython.Compiler.Options
 Cython.Compiler.Options.annotate = True
 
@@ -32,7 +43,7 @@ setup(
                   depends = ['numeric_monoid.pxd', 'monoid.hpp', 'treewalk.hpp'],
                   language="c++",
                   extra_compile_args = ['-std=c++11', '-O3',
-                                        '-march=native', '-mtune=native',
+                                        MARCH, MTUNE,
                                         '-fcilkplus'],
                   define_macros = [('NDEBUG', '1'), ('MAX_GENUS','86')],
                   include_dirs = [SAGE_C,SAGE_DEV],

### 2013-10-22 Papier/article.tex
Reduction of section 2

@@
 \title{Exploring the tree of numerical semigroups}
-\author{XXXX}
+\author{Jean Fromentin and  Florent Hivert}
 
@@
 These values have been obtain by M. Bras-Amor\'os (view \cite{BrasAmoros2008} for more details for $g\leq 50)$. 
-On his home page \cite{Delgado}, M. Delgado  gives the value of $n_{55}$ without specifying the values of $n_{53}$ and $n_{54}$. 
-M.~-Bras Amor\'os used a depth first search exploration of the tree of numerical semigroups up to a given genus.
+On his home page~\cite{Delgado}, M. Delgado  gives the value of $n_{55}$ without specifying the values of $n_{53}$ and~$n_{54}$.
+
+M.~Bras-Amor\'os used a depth first search exploration of the tree of numerical semigroups up to a given genus.
 This tree was introduced by J.C. Rosales and al. in \cite{Rosales} and it is the subject of the Section~\ref{S:Tree}.
-
 Starting with all the numerical semigroups of genus $49$ she obtained the number of numerical semigroups of genus $50$ in $18$ days on a pentium D runing at $3$GHz. 
@@
 The paper is divided as follows.
-In section~\ref{S:Tree} we describe the tree of numerical semigroups and give bounds for some parameters attached to a numerical semigroup.
+In Section~\ref{S:Tree} we describe the tree of numerical semigroups and give bounds for some parameters attached to a numerical semigroup.
 In Section~\ref{S:DecNumber} we describe a new representation of numerical semigroups that is well suited to the construction of the tree. 
@@
 \label{S:Tree}
-We first start by some notations.
+
+We start this section with definitions and properties of numerical semigroups that will be used in the sequel.
+For a more complete introduction, the reader can usefully consults the book \emph{Numerical Semigroups} of J.C.~Rosales and P.A.~Garc\'ia-S\'anchez \cite{BookNS} or the book \emph{The Diophantine Frobenius Problem} of J.L.~Ram\'irez~Alfons\'in \cite{BookDFP}.
 
@@
 \begin{defi}
-A subset $X$ of a semigroup is a \emph{generating set} of $S$ if every element of $S$ can be express as a sum of elements in $X$. 
+A subset $X=\{x_1<x_2<...<x_n\}$ of a semigroup is a \emph{generating set} of $S$ if every element of $S$ can be express as a sum of elements in $X$. 
+In this case we write $S=\left<x_1,...,x_\ell\right>$.
 \end{defi}
-
-\begin{nota}
-A numerical semigroup admitting $X=\{x_1<x_2<...<x_n\}$ as generating set is denoted  by $S=\left<x_1,...,x_\ell\right>$.
-\end{nota}
-
-We now introduce a specific generating set.
-
-\begin{defi}
-A non-zero element $x$ of a numerical semigroup $S$ is said to be \emph{irreducible} if it cannot be expressed as a sum of two non-zeros elements of $S$.  
-We denote by $\Irr(S)$ the set of all irreducible elements of $S$.
-\end{defi}
-
-\begin{prop}
-Let $S$ be a numerical semigroup.
-Then $\Irr(S)$ is the minimal generating set of $S$.
-\end{prop}
-
-\begin{proof}
-Assume for a contradiction, that there exists an integer  $x$ in $S$ than cannot be decomposed as a sum of irreducible elements. 
-We may assume that $x$ is  minimal with this property. 
-As  $x$ cannot be irreducible, there exist $y$ and $z$ in~$S\setminus\{0\}$ satisfying $x=y+z$. 
-Since we have $y<x$ and $z<x$, the integers $y$ and $z$ can be expressed as a sum of irreducible elements of $S$ and so $x=y+z$ is a sum of irreducible elements, in contradiction to hypothesis.
-
-As irreducible elements of $S$ cannot be decomposed as the sum of two non-zeros integers in $S$, they must occur in each generating set of $S$.
-\end{proof}
-
-We recall that Ap\'ery elements of a numerical semigroup $S$ associated to $m(S)$ are the integers $x$ in $S$ such that $x-m(S)$ is no longer in $S$.
-We denote by~$\App(S)$ the set of these elements. 
-It is well known that the cardinality of $\App(S)$ is exactly $m(S)$ (see~\cite{Book} for example). 
-Note that the set $\Irr(S)$ is in included in $\App(S)$.
-In particular, the set $\Irr(S)$ is finite and its cardinality is at most $m(S)$.
 
@@
 
+A non-zero element $x$ of a numerical semigroup $S$ is said to be \emph{irreducible} if it cannot be expressed as a sum of two non-zeros elements of $S$.  
+We denote by $\Irr(S)$ the set of all irreducible elements of $S$.
+
+\begin{lem}[Lemma 2.3, Chap. I, \cite{BookNS}]
+Let $S$ be a numerical semigroup. The set $\Irr(S)$ is the minimal generating set of $S$.
+\end{lem}
+
+%\begin{proof}
+%Assume for a contradiction, that there exists an integer  $x$ in $S$ than cannot be decomposed as a sum of irreducible elements. 
+%We may assume that $x$ is  minimal with this property. 
+%As  $x$ cannot be irreducible, there exist $y$ and $z$ in~$S\setminus\{0\}$ satisfying $x=y+z$. 
+%Since we have $y<x$ and $z<x$, the integers $y$ and $z$ can be expressed as a sum of irreducible elements of $S$ and so $x=y+z$ is a sum of irreducible elements, in contradiction to hypothesis. 
+%
+%As irreducible elements of $S$ cannot be decomposed as the sum of two non-zeros integers in $S$, they must occur in each generating set of $S$.
+%\end{proof}
+
+%Let $S$ be a numericla semigroup. As $m(S)$ lies in $S$ the irreducible elements are include in
+%\[
+%\App(S)=\{x\in S \, |\, x-m(S)\not\in S\}
+%\]
+%The set $\App(S)$ is the so-called Ap\'ery set relativey to the element $m(S)$.
+%
+%\begin{lem}[Lemma 2.4, Chap. I, \cite{BookNS}]
+%Let $S$ be a numerical semigroup. The Apéry set $\App(S)$ is finite of cardinality $m(S)$.
+%\end{lem}
+%
+%The following result is a direct consequence of the previous Lemma and the Inclusion of $\Irr(S)$ in $m(S)$.
+%
+%\begin{prop}
+%\label{P:Res}
+%A numerical semigroup has at most $m(S)$ irreducible elements.
+%\end{prop}
+
+The differents parameters we have defined on a numerical semigroup, statisfy the following relations.
+
+\begin{prop}
+\label{P:Res}
+Let $S$ be a numerical semigroup. We have
+
+$i)$  $x\in \Irr(S)$ implies $x\leq c(S)+m(S)-1$;
+
+$ii)$ $m(S)\leq g(S)+1$;
+
+$iii)$ $c(S)\leq 2 g(S)$.
+\end{prop}
+
+\begin{proof}
+$i)$ Let $x$ ba a positive integer satisfying $x \geq c(S)+m(S)$. 
+As the relations $x\geq c(S)$ and $x-m(S)$ hold, the integers $x$ and $x-m(S)$ belong to $S$.
+Since $x=(x-m(S))+m(S)$ hold with $x-m(S)\not=0$ and $m(S)\not=0$, the integer $x$ could not be irreducible.
+
+$ii)$ The set $\NN\setminus{S}$ of cardinality $g(S)$ contains $\{1,..,m(S)-1\}$ and so the relation $m(S)\leq g(S)+1$ holds.
+
+$iii)$ Let $x$ be an element of $S\cap[0,c(S)-1]$. 
+The integer $y=c(S)-1-x$ lies in $[0,c(S)-1]$. 
+Moreover $y$ is not in $S$, for otherwise  we write $c(S)-1=y+x$ where $x,y$ are elements of $S$ implying $c(S)-1\in S$. 
+In contradiction with the definition of conductor. 
+Thus we define an involution 
+\[
+\begin{array}{rcl}
+\psi:[0,c(S)-1]&\to&[0,c(S)-1]\\
+x&\mapsto&c(S)-1-x
+\end{array}
+\]
+that send $S\cap[0,c(S)-1]$ into $[0,c(S)-1]\setminus S$.
+The cardinality of $[0,c(S)[\setminus S$ is exactly $g(S)$. Let us denote by $k$ the cardinality  of $S\cap[0,c(S)-1]$. 
+Since $\psi$ is injective we must have $k\leq g(S)$. 
+From the relation 
+\[
+[0,c(S)[=S\cap[0,c(S)[\sqcup[0,c(S)[\setminus S
+\]
+ we obtain $c(S)=k+g(S)$ and so $c(S)\leq 2 g(S)$.
+\end{proof}
+
+A consequence of Proposition~ \ref{P:Res}~$i)$ is that $\Irr(S)$ is finite of cardinality at most $c(S)+m(S)-1$.
+In fact, using Apéry set, we can prove that  the cardinality of $\Irr(S)$ is at most $m(S)$. See Section 2 of Chapter I of \cite{BookNS} for more details.
+
 Let  $S$ be a numerical semigroup. The set $T=S\cup\{c(S)-1\}$ is also a numerical  semigroup and its genus is $g(S)-1$. 
 As  $[c(S)-1,+\infty[$ is included in $T$ we have $c(T)\leq c(S)-1$. 
-Therefore every semigroup $S$ of genus $g$ can be obtained from a semigroup $T$ of genus $g-1$ by removing an element of $T$ greater than or equal to $c(T)$.
+Therefore every semigroup $S$ of genus $g$ can be obtained from a semigroup $T$ of genus $g-1$ by removing an element of~$T$ greater than or equal to $c(T)$.
 
@@
 Let $a$ and $b$ be two integers belonging to $S^x$. 
-The set $S$ is stable under addition, hence $a+b$ lies in $S$. 
+The set~$S$ is stable under addition, hence $a+b$ lies in $S$. 
 As $S$ is equal to $S^x\cup\{x\}$, the integer $a+b$ also lies in $S^x$ except if it is equal to $x$. 
@@
 The root of the tree is the unique semigroup of genus $0$, \ie, $\left<1\right>$ that is equal to $\NN$. 
-If~$S$ is a semigroup in the tree,  the sons of $S$ are exactly the semigroup $S^x$ where $x$ belongs to $\Irr(S)\cap[c(s),+\infty[$. 
-By convention, when depicting the tree, the numerical semigroup $S^x$ is in the left of $S^y$ if $x$ is smaller than $y$.
+If~$S$ is a semigroup in the tree,  the sons of $S$ are exactly the semigroup~$S^x$ where $x$ belongs to $\Irr(S)\cap[c(s),+\infty[$. 
+By convention, when depicting the tree, the numerical semigroup $S^x$ is in the left of $S^y$ if $x$ is smaller than~$y$.
 
-The above remarks imply that a semigroup $S$ has depth $g$ in $≠mathcal{T}$ if and only if its genus is~$g$, see Figure~\ref{F:Tree}.
+The above remarks imply that a semigroup $S$ has depth $g$ in $\mathcal{T}$ if and only if its genus is~$g$, see Figure~\ref{F:Tree}.
 We denote by $\mathcal{T}^{\leq g}$ the subtree of $\mathcal{T}$ resticted to all semigroup of genus $\leq g$.
 
- \begin{figure}[htf!]
+ \begin{figure}[hf!]
  % \[
@@
 
-As the reader can check, the main difficulty to characterize the sond of a semigroup is to determine its irreducible elements.
- In \cite{NumericalSgps}, the semigroup are given by their Ap\'ery set $\App(S)$ and then the main difficulty is to describe $\App(S_x)$ from $\App(S)$. 
+As the reader can check, the main difficulty to characterize the son of a semigroup is to determine its irreducible elements.
+ In \cite{NumericalSgps}, the semigroup are given by their Ap\'ery set and then the main difficulty is to describe the Ap\'ery set of $S_x$  from this of $S$. 
  This approach is elegant but not sufficiently \emph{basic} for our optimisations.
-
-We conclude this section with some basic results on numerical semigroups of a given genus.
-Let $S$ be a numerical semigroup. 
-We first prove :
-\begin{equation}
-\label{E:Irr}
-x\in \Irr(S)\quad\textrm{implies}\quad x\leq c(S)+m(S).
-\end{equation}
-If $y$ is a positive integer with $y \geq c(S)+m(S)$ then $y$ lies in $S$ (as $y\geq c(S)$ holds). Moreover, we always have $y-m(S)\geq c(s)$ and so $y$ is not irreducible.
-
-As $\NN\setminus{S}$ contains $\{1,..,m(S)-1\}$ we have 
-\begin{equation}
-\label{E:m}
-m(S)\leq g(S)+1.
-\end{equation}
-Let $x$ be an element of $S\cap[0,c(S)-1]$. 
-The integer $y=c(S)-1-x$ lies in $[0,c(S)-1]$. 
-Moreover $y$ is not in $S$, for otherwise  we write $c(S)-1=y+x$ where $x,y$ are elements of $S$ implying $c(S)-1\in S$. 
-In contradiction with the definition of conductor. 
-Thus we define an involution 
-\[
-\begin{array}{rcl}
-\psi:[0,c(S)-1]&\to&[0,c(S)-1]\\
-x&\mapsto&c(S)-1-x
-\end{array}
-\]
-that send $S\cap[0,c(S)-1]$ into $[0,c(S)-1]\setminus S$.
-The cardinality of $[0,c(S)[\setminus S$ is exactly $g(S)$. Let us denote by $k$ the cardinality  of $S\cap[0,c(S)-1]$. 
-Since $\psi$ is injective we must have $k\leq g(S)$. 
-From the relation 
-\[
-[0,c(S)[=S\cap[0,c(S)[\sqcup[0,c(S)[\setminus S
-\]
- we obtain $c(S)=k+g(S)$ and so 
-\begin{equation}
-\label{E:c}
-c(S)\leq 2 g(S).
-\end{equation}
 
@@
 and $d_S(x)=\card D_S(x)$.
-We called $d_S(x)$ the $S$-\emph{decomposition number} of $x$.
+We called $d_S(x)$ the $S$-\emph{decomposition number} of~$x$.
 The application $d_S:\NN\to\NN$ is the $S$-\emph{decomposition numbers function}.
@@
 In other words if we define~$D'_S(x)$ to be the set of all $(y,z)\in S\times S$ with $x=y+z$ and $y\leq z$ then~$D_S(x)$ is the image of $D'_S(x)$ under the projection on the first coordinate.
- Hence~$D_S(x)$ describes how~$x$ can be decomposed as sums of two elements of $S$,-.
+ Hence~$D_S(x)$ describes how~$x$ can be decomposed as sums of two elements of $S$.
  This is why $d_S(x)$ is called the $S$-decomposition number of $x$.
@@
 \begin{proof}
-By~\eqref{E:c} we have $c(S)\leq g(S)$ and so the $S$-decomposition number of $c(S)$ occurs in $\delta_S$.
+By Proposition~\ref{P:Res} $iii)$ we have $c(S)\leq 2g(S)$ and so the $S$-decomposition number of $c(S)$ occurs in $\delta_S$.
 Since $c(S)$ is equal to $\max(\NN\setminus S)$, Proposition~\ref{P:delta} implies 
@@
 \]
-By \eqref{E:m} the relation $m(S)\leq g(S)+1$ holds. 
+By Proposition~\ref{P:Res} $ii)$, the relation $m(S)\leq g(S)+1$ holds. 
 This implies that the $S$-decomposition number of $m(S)$ appears in $\delta_S$ : 
@@
 \]
-Since, by \eqref{E:Irr},  all irreducible elements are smaller than $c(S)+m(S)-1$, which is itself smaller than $3G$, equations  \eqref{E:m} and \eqref{E:c} give
+Since, by Proposition~\ref{P:Res} $i)$,  all irreducible elements are smaller than $c(S)+m(S)-1$, which is itself smaller than $3G$ the relations $ii)$ and $iii)$ of Proposition~\ref{P:Res} give
 \[
@@
 Let us now prove the complexity statement. 
-Since by \eqref{E:c} and \eqref{E:m} we have $x\leq 3G$ together with $m(S)\leq G+1$, each line from \tt{2} to \tt{8} is done in time $O(\log(G))$.
+Since by relations $ii)$ and $iii)$ of Proposition~\ref{P:Res} we have $x\leq 3G$ together with $m(S)\leq G+1$, each line from \tt{2} to \tt{8} is done in time $O(\log(G))$.
 The \textbf{for} loop needs $O(G)$ steps and each step is done in time $O(\log(G))$. 
@@
 As the number of sons of a  semigroup $S$ is the number of $S$-irreducible elements in $[c(S),c(S)+m(S)-1]$,  a semigroup $S$ has at most $m(S)$ sons. 
-By~\eqref{E:m}, this implies that a semigroup of genus $g$ as at most $g+1$ sons.
+By~Proposition~\ref{P:Res} $ii)$, this implies that a semigroup of genus $g$ as at most $g+1$ sons.
 Therefore the stack contains at most $g+1$ semigroup of genus $g+1$ for $g\leq G$. 
@@
 A semigroup is represented by a quadruple $(c(S),g(S),m(S),\delta_S)$. 
-By equations~\eqref{E:c} and~\eqref{E:m}, we have $c\leq 2g(S)$ and $m\leq g(S)+1$. 
+By relations $ii)$ and $iii)$ of Proposition~\ref{P:Res}, we have $c\leq 2g(S)$ and $m\leq g(S)+1$. 
 As $g(S)\leq G$ holds, the integers~$c$, $g$  and $m$ of the representation of $S$ require a memory space in $O(\log(G))$. 

### 2014-01-06 Papier/article.tex
Modification of proof of Proposition 3.5 and small correction in sections <=3

@@
 is a numerical semigroup.
-The \emph{genus} of a numerical semigroup $S$, denoted by~$g(S)$, is the cardinality of $\NN\setminus S$, \ie, $g(S)=\card(\NN\setminus S)$. 
- For example the genus of $S_E$ defined in \eqref{E:NSG} is $6$,  the cardinality of $\{1,2,4,5,8,11\}$
+The \emph{genus} of a numerical semigroup $S$, denoted by~$g(S)$, is the cardinality of $\NN\setminus S$.
+ For example the genus of $S_E$ is $6$,  the cardinality of $\{1,2,4,5,8,11\}$
 
-For a given positive integer $g$, the number of numerical semigroups  is finite and is denoted by $n_g$. 
+For a given positive integer $g$, the number of numerical semigroups of genus $g$ is finite and is denoted by $n_g$. 
 In  J.A. Sloane's \emph{on-line encyclopedia of integer}~\cite{OEIS} we find the values of $n_g$ for $g\leq 52$. 
@@
 M.~Bras-Amor\'os used a depth first search exploration of the tree of numerical semigroups up to a given genus.
-This tree was introduced by J.C. Rosales and al. in \cite{Rosales} and it is the subject of the Section~\ref{S:Tree}.
+This tree was introduced by J.C.~Rosales and al. in \cite{Rosales} and it is the subject of the Section~\ref{S:Tree}.
 Starting with all the numerical semigroups of genus $49$ she obtained the number of numerical semigroups of genus $50$ in $18$ days on a pentium D runing at $3$GHz. 
@@
 A subset $X=\{x_1<x_2<...<x_n\}$ of a semigroup is a \emph{generating set} of $S$ if every element of $S$ can be express as a sum of elements in $X$. 
-In this case we write $S=\left<x_1,...,x_\ell\right>$.
+In this case we write $S=\left<x_1,...,x_n\right>$.
 \end{defi}
@@
 \begin{lem}[Lemma 2.3, Chap. I, \cite{BookNS}]
-Let $S$ be a numerical semigroup. The set $\Irr(S)$ is the minimal generating set of $S$.
+Let $S$ be a numerical semigroup. The set $\Irr(S)$ is the minimal generating set of $S$ relatively to the inclusion ordering.
 \end{lem}
@@
 $i)$ Let $x$ ba a positive integer satisfying $x \geq c(S)+m(S)$. 
-As the relations $x\geq c(S)$ and $x-m(S)$ hold, the integers $x$ and $x-m(S)$ belong to $S$.
-Since $x=(x-m(S))+m(S)$ hold with $x-m(S)\not=0$ and $m(S)\not=0$, the integer $x$ could not be irreducible.
+As the integers $x$ and $x-m(S)$ are greater than $c(S)$, they belong to $S$.
+Since $x$ equals $(x-m(S))+m(S)$ with $x-m(S)\not=0$ and $m(S)\not=0$, it could not be irreducible.
 
@@
 \]
-that send $S\cap[0,c(S)-1]$ into $[0,c(S)-1]\setminus S$.
-The cardinality of $[0,c(S)[\setminus S$ is exactly $g(S)$. Let us denote by $k$ the cardinality  of $S\cap[0,c(S)-1]$. 
+that send $S\cap[0,c(S)-1]$ into $[0,c(S)-1]\setminus S$ which is  of cardinality $g(S)$. 
+Let us denote by $k$ the cardinality  of $S\cap[0,c(S)-1]$. 
 Since $\psi$ is injective we must have $k\leq g(S)$. 
@@
 
-A consequence of Proposition~ \ref{P:Res}~$i)$ is that $\Irr(S)$ is finite of cardinality at most $c(S)+m(S)-1$.
-In fact, using Apéry set, we can prove that  the cardinality of $\Irr(S)$ is at most $m(S)$. See Section 2 of Chapter I of \cite{BookNS} for more details.
+A consequence of Proposition~ \ref{P:Res}~$i)$ is that $\Irr(S)$ is finite ant its cardinality is at most $c(S)+m(S)-1$.
+More precisely, using Ap\'ery set, we can prove that  the cardinality of $\Irr(S)$ is at most $m(S)$. See Section 2 of Chapter I of~\cite{BookNS} for more details.
 
-Let  $S$ be a numerical semigroup. The set $T=S\cup\{c(S)-1\}$ is also a numerical  semigroup and its genus is $g(S)-1$. 
+We now explain the construction of the tree of numerical semigroups.
+Let~$S$ be a numerical semigroup. The set $T=S\cup\{c(S)-1\}$ is also a numerical  semigroup and its genus is $g(S)-1$. 
 As  $[c(S)-1,+\infty[$ is included in $T$ we have $c(T)\leq c(S)-1$. 
@@
 By convention, when depicting the tree, the numerical semigroup $S^x$ is in the left of $S^y$ if $x$ is smaller than~$y$. 
-
-The above remarks imply that a semigroup $S$ has depth $g$ in $\mathcal{T}$ if and only if its genus is~$g$, see Figure~\ref{F:Tree}.
-We denote by $\mathcal{T}^{\leq g}$ the subtree of $\mathcal{T}$ resticted to all semigroup of genus $\leq g$.
+With this construction, a semigroup $S$ has depth $g$ in $\mathcal{T}$ if and only if its genus is~$g$, see Figure~\ref{F:Tree}.
+We denote by $\mathcal{T}_{\leq g}$ the subtree of $\mathcal{T}$ resticted to all semigroup of genus $\leq g$.
 
@@
   \]
-\caption{The first five layers of the tree $\mathcal{T}$ of numerical semigroups, corresponding to $\mathcal{T}^{\leq 5}$.
+\caption{The first four layers of the tree $\mathcal{T}$ of numerical semigroups, corresponding to $\mathcal{T}_{\leq 4}$.
 A generator of a semigroup is in gray if it is not in $[c(S),+\infty[$.
@@
 As the reader can check, the main difficulty to characterize the son of a semigroup is to determine its irreducible elements.
- In \cite{NumericalSgps}, the semigroup are given by their Ap\'ery set and then the main difficulty is to describe the Ap\'ery set of $S_x$  from this of $S$. 
+ In \cite{NumericalSgps}, the semigroup are given by their Ap\'ery set and then the difficulty is to describe the Ap\'ery set of $S_x$  from this of $S$. 
  This approach is elegant but not sufficiently \emph{basic} for our optimisations.
@@
 The aim of this section is to describe a new representation of numerical semigroups.
-which is well suited to a efficient exploration of the tree~$\mathcal{T}$ of numerical semigroups.
+which is well suited to an efficient exploration of the tree~$\mathcal{T}$ of numerical semigroups.
 
@@
 Hence, the set $D_S(x)$ is equal to~$\{0\}$ and we have $d_S(x)=1$. 
-Conversely, let $x$ such that $d_S(x)=1$. 
+Conversely, let $x$ such that $d_S(x)$ is $1$. 
 By  $i)$ the  integer $x$ must be in $S$. 
@@
 
-We now explain how to compute the $S$-decomposition numbers function of a numerical semigroup  from these of its father.
+We now explain how to compute the $S$-decomposition numbers function of a numerical semigroup  from this of its father.
 
@@
 \[
-D_{S^x}(y)=\{z \in S^x\ |\ y-z\in S^x\ \text{and}\  2z\leq y\}
+D_{S^x}(y)=\{z \in S^x\ |\ y-z\in S^x\ \text{and}\  2z\leq y\}.
 \]
-and $D_{S^x}(y)$ is a subset of $D_S(y)$.
-We have $D_S(y)\setminus D_{S^x}(y)=E\sqcup F$ where
+We construct three sets as follow:
 \begin{align*}
-E&=\begin{cases}
-\{x\} &\text{if $y-x\in S^x$ and $2x\leq y$,}\\
-\emptyset & \text{otherwise.}
-\end{cases}\\
-F&=\{z\in S\ |\ y-z=x\ \text{and}\ 2z\leq y\}
+E&=\{z\in S^x\ |\ y-z=x\ \text{and}\ 2z\leq y\}\\
+F&=\begin{cases}
+\{x\}&\text{if $y-x\in S$ and $2x\leq y$},\\
+\emptyset&\text{otherwise}.
+\end{cases}
 \end{align*}
-Since  the relation $y-x\in S^x$ can be rewritten as the conjunction of $y-x\in S$ and $y\not=2x$, we obtain
+We have the relation $D_{S}(y)=D_{S^x}(y)\sqcup E\sqcup F$ and so $d_S(y)$ is equal to $d_{S^x}(y)+\card(E)+\card(F)$.
+and $D_{S^x}(y)$ is a subset of $D_S(y)$. We start by determining $E$. 
+The relation $y-z=x$ implies $z=y-x$ and $2z\leq y$ becomes $y\leq 2x$. Thus we obtain
 \[
 E=\begin{cases}
-\{x\} &\text{if $y-x\in S$ and $2x< y$,}\\
-\emptyset & \text{otherwise.}
+\{y-x\}&\text{if $y-x\in S^x$ and $y\leq 2x$}\\
+\emptyset&\text{otherwise}
 \end{cases}
 \] 
-By Proposition~\ref{P:Dandd}, the relation $y-x\in S$ is equivalent to $y\geq x$ and $d_S(y-x)>0$, we have
+The relation $y-x\in S^x$ together with $y\leq2x$ is equivalent to $y-x\in S$ together with $y<2x$.
+Therefore we obtain
 \[
 E=\begin{cases}
-\{x\} &\text{if $y\geq x$ and $d_S(y-x)>0$ and $2x< y$,}\\
-\emptyset & \text{otherwise.}
+\{y-x\}&\text{if $y-x\in S$ and $y< 2x$}\\
+\emptyset&\text{otherwise}
 \end{cases}
 \] 
-On the other hand, we have 
-\begin{align*}
-F&=\{z\in S\ |\ y-z=x\ \text{and}\ 2z\leq y\}\\
-&=\{z\in S\ |\ z=y-x\ \text{and}\ y\leq 2x\}\\
-&=\begin{cases}
-\{y-x\} & \text{if $y-x\in S$ and $y\leq 2x$,}\\
-\emptyset & \text{otherwise.}
-\end{cases}
-\end{align*}
-As in the case of $E$, we get 
-\[
-F=\begin{cases}
-\{y-x\} & \text{if $y\geq x$ and $d_S(y-x)>0$ and $y\leq 2x$,}\\
-\emptyset & \text{otherwise.}
-\end{cases}
-\]
-In $E$ we have the constraint $2x<y$ while in $F$ we have $y\leq 2x$, hence only one of the sets $E$ or $F$ can be non-empty. 
-This implies
-\[
-D_S(y)\setminus D_{S^x}(y)=
-\begin{cases}
-\{x\} & \text{if $y\geq x$ and $d_S(y-x)>0$ and $y\leq 2x$,}\\
-\{y-x\} & \text{if $y\geq x$ and $d_S(y-x)>0$ and $y>2x$,}\\
-\emptyset & \text{otherwise.}
-\end{cases}
-\]
-Therefore we obtain
-\[
-d_{S^x}(y)=\begin{cases}
-d_S(y)-1&\text{if $y\geq x$ and $d_S(y-x)>0$,}\\
-d_S(y)& \text{otherwise.}
-\end{cases}
-\]
+Assume $y-x$ lies in $S$.  If $y<2x$ holds, we have $E=\{y-x\}$ and $F=\emptyset$. Otherwise, we obtain $E=\emptyset$ and $F=\{x\}$. 
+Then if $y-x\in S$ holds, we have $\card(E)+\card(F)=1$ and so $d_{S}(y)$ is equal to $d_{S^x}(y)+1$.
+If $y-x$ not belong to $S$, then $E$ and $F$ are empty implying $d_{S}(y)=d_{S^x}(y)$.
+As the condition $y-x\in S$ is equivalent to $y\geq x$ together with $d_S(y-x)>0$ by Proposition~\ref{P:Dandd} the result holds.
 \end{proof}
@@
 
+\label{S:Opti}
+
 Though there are asymptotically faster algorithm than the one presented here,
@@
 \end{itemize}
-
-\label{S:Opti}
 

### 2014-08-31 Papier/article.tex
Florent's modifications.

@@
 \usepackage{url}
+\usepackage{moreverb}
 \usepackage{algorithm}
@@
 
+\newcommand{\TODO}[1]{\marginpar{\small To do: #1}}
+
 \begin{document}
@@
 
+\TODO{Present the goal of the paper.}
+
 The paper is divided as follows.
@@
 \begin{defi}
-A subset $X=\{x_1<x_2<...<x_n\}$ of a semigroup is a \emph{generating set} of $S$ if every element of $S$ can be express as a sum of elements in $X$. 
+A subset $X=\{x_1<x_2<...<x_n\}$ of a semigroup is a \emph{generating set} of $S$ if every element of $S$ can be expressed as a sum of elements in $X$. 
 In this case we write $S=\left<x_1,...,x_n\right>$.
@@
 
-A consequence of Proposition~ \ref{P:Res}~$i)$ is that $\Irr(S)$ is finite ant its cardinality is at most $c(S)+m(S)-1$.
+A consequence of Proposition~ \ref{P:Res}~$i)$ is that $\Irr(S)$ is finite and its cardinality is at most $c(S)+m(S)-1$.
 More precisely, using Ap\'ery set, we can prove that  the cardinality of $\Irr(S)$ is at most $m(S)$. See Section 2 of Chapter I of~\cite{BookNS} for more details.
@@
  Hence~$D_S(x)$ describes how~$x$ can be decomposed as sums of two elements of $S$.
- This is why $d_S(x)$ is called the $S$-decomposition number of $x$.
+ This is we decide to call $d_S(x)$ the $S$-decomposition number of $x$.
 
@@
 \label{S:Algo}
-We can easily explore the tree of numerical semigroups up to a genus $G$ using a depth search first algorithm using a stack (see~\ref{A:Exp}). 
-This approach does not seem to have been used before. 
-In particular,  M. Bras-Amor\'os and M. Delgado use instead a breadth search first exploration. 
-The main advantage in our approach is the small memory needs. 
-The cost to pay is that, if we want to explore the tree deeper, we must restart from the root.
-
+We can easily explore the tree of numerical semigroups up to a genus $G$ using
+a depth first search algorithm (see Algorithm~\ref{A:ExpRec}). This approach does not seem to
+have been used before.  In particular, M. Bras-Amor\'os and M. Delgado use
+instead a breadth first search exploration.  The main advantage in our
+approach is the small memory needs.  The cost to pay is that, if we want to
+explore the tree deeper, we must restart from the root.
 \begin{algorithm}
-\caption{Depth search first exploration of the tree of numerical semigroups}\label{A:Exp}
+\caption{Recursive Depth first search exploration of the tree of numerical semigroups}\label{A:ExpRec}
+\begin{algorithmic}[1]
+\Procedure{ExploreRec}{\tt{S}, \tt{G}}
+	\If{$g(\tt{S})< \tt{G}$}
+		\For{\tt{x} from $c(\tt{S})$ to $c(\tt{S})+m(\tt{S})$}
+			\If{$\tt{x}\in \Irr(\tt{S})$}
+				\State \Call{ExploreRec}{$\tt{S}^\tt{x}, \tt{G}$}
+			\EndIf
+		\EndFor
+	\EndIf
+\EndProcedure
+\State \Call{ExploreRec}{$\NN, \tt{G}$}
+\end{algorithmic}
+\end{algorithm}
+
+Equivalently, we can use an iterative version which uses a stack (see Algorithm~\ref{A:Exp}).
+\begin{algorithm}
+\caption{Iterative Depth first search exploration of the tree of numerical semigroups}\label{A:Exp}
 \begin{algorithmic}[1]
@@
 
-In Algorithm~\ref{A:Exp} we do not specify how to compute $c(S)$, $g(S)$ and $m(S)$  from $S$ neither how to test if an integer is irreducible. 
+In Algorithm~\ref{A:ExpRec} we do not specify how to compute $c(S)$, $g(S)$ and $m(S)$  from $S$ neither how to test if an integer is irreducible. 
 It also miss  the characterisation of $S^x$ from $S$.
@@
 
-\subsection{Florent's rant}
-
-We use the following optimization tricks
+To get the greatest speed from modern processor we used several optimization
+tricks, which we will elaborate in the following section:
 \begin{itemize}
-\item Vectorisation (MMX, SSE, GCC's 
-\verb|__attribute__ ((vector_size(16)));|
-) and careful memory alignment.
-\item Shared memory multicore computing using \texttt{CILK++} for low level
+\item Vectorization (MMX, SSE instructions sets) and careful memory alignment.
+\item Shared memory multi-core computing using \texttt{CILK++} for low level
   enumerating tree branching;
@@
 
+\subsection{Vectorization}
+
 Assume for example that we want to construct the tree $\mathcal{T}^{\leq 100}$ of all  numerical semigroup of  genus smaller than $100$. 
@@
 
-\noindent In this version we can see that the cursors \texttt{src} and \texttt{dst} move at the same time and that the modification of  the  value pointed by \texttt{dst}   only  needs to access the values pointed by \tt{src} and \tt{dst}.
-We can therefore work in multiple entries at the same time without collision. 
-Current CPUs allow this thanks to the SIMD technologies as MMX, SSE, etc.
-The acronym SIMD stands for Single Operation Multiple Data.
-
-The MMX technology  permits to work on $8$ bytes in parallel while the SSE works in $16$ bytes. 
-As the rest of the CPU works on $8$ bytes, the SSE technology needs some constraint on memory access than cannot be fulfilled in our algorithm. 
-This motivate our choice to use the MMX technology. 
-More precisely, we use three commands: \tt{pcmpeqb}, \tt{pandn} and \tt{psubb}.
- These commands work on two arrays of $8$ bytes, called~\tt{s} and \tt{d} here. 
- In each case the array \tt{d} is modified. 
- We denote bytes by $\{\tt{0},\tt{1}\}$-words of length $8$.
- After a call to \tt{pcmpeqb(s,d)} the $i$th entry of $d$ contains \tt{11111111} if $\tt{s}[i]=\tt{d}[i]$ holds and \tt{00000000} otherwise. 
- The command \tt{pand(s,d)}  store in $\tt{d}[i]$  the value of the byte-logic operation \tt{s}[i] \tt{and} (\tt{not} \tt{d}[i]).
-  When \tt{pand(s,d)} is called the new value of $\tt{d}[i]$ is $\tt{s}[i]-\tt{d}[i]$.  
-  Glueing all the pieces together we obtain the following version of the \textbf{for} loop of Algorithm~\textsc{Son} :
-
+\noindent In this version we can see that the cursors \texttt{src} and
+\texttt{dst} move at the same time and that the modification of the value
+pointed by \texttt{dst} only needs to access the values pointed by \tt{src}
+and \tt{dst}.  We can therefore work in multiple entries at the same time
+without collision.  Current CPUs allow this thanks to the SIMD technologies as
+MMX, SSE, etc.  The acronym SIMD stands for Single Operation Multiple Data.
+We used SSE4.1 technology as it allows for the largest speedup. This need to
+respect some constraints in the memory organization of the data, namely
+``memory alignment''. Recall that an address is 16~bytes aligned if it is a
+multiple of 16. SSE memory access are much faster for aligned access.
+
+The computation of the children is then performed as follows. First, the
+parent's decomposition numbers are copied in the children's using the
+following C++ code
+\begin{listing}{1}
+void copy_blocks(dec_blocks &dst, dec_blocks const &src) {
+  for (ind_t i=0; i<NBLOCKS; i++) dst[i] = src[i];
+}
+\end{listing}
+Here \tt{dec\_blocks} is a type for arrays of 16~bytes blocks whose size
+\tt{NBLOCKS} are just large enough to store the decomposition numbers (that is
+$3G$ rounded up to a multiple of 16). The instruction \tt{dst[i] = src[i]}
+actually copy a full 16~bytes block.
+
+Then the core of the \textbf{while} loop in the preceding algorithm is
+translated as a for loop as follows (recall that $x$ denotes the generator if the
+father which is to be removed in the children):
+\begin{listing}{1}
+start = x >> 4;      // index of the block containing x
+shift = x & 0xF;     // offset of x inside the block
+...                  // some specific instructions to
+...                  // handle the first incomplete block.
+for (long int i=start+1; i<NBLOCKS; i++) {
+  block = load_unaligned_epi8(src + ((i-start)<<4) - shift);
+  dst[i] -= ((block != zero) & one);
+}
+\end{listing}
+The instruction \tt{load\_unaligned\_epi8} (specific to SSE technology) loads
+in the variable \tt{block} 16 consecutive entries of the decomposition number
+of the parent (called \tt{src} monoid). Those entries will be used to compute
+the entries $16i, \dots 16i+15$ of the children monoid. Since the removed
+generator $x$ is not necessarily a multiple of $16$, the data are not aligned
+in memory, hence the use of a specific instruction. The \tt{zero}
+(resp. \tt{one}) constants are initialized as 16 bytes equal to $0$
+(resp. $1$). The comparison \tt{(block != zero)} therefore returns a block
+which contains $0$ in the bytes corresponding the $0$ entries of block and
+$255$ in the non zero one. This results is then bitwise anded with one so that
+the instruction actually performs a 16~bytes parallel version of
+\begin{algorithmic}
+\State $\tt{dst} \gets \tt{dst} -
+\textbf{ if } \tt{block} \neq 0 \textbf{ then } 1 \textbf{ else } 0$
+\end{algorithmic}
+As we previously said, to gain more speed this core loop is actually unrolled
+using some kind of Duff device as
+\begin{algorithmic}
+\State \textbf{Jump to case } $\tt{start}+1$
+\State \textbf{case 1:} \tt{block = ...; dst[1] -= ((block != zero) \& one)}
+\State \textbf{case 2:} \tt{block = ...; dst[2] -= ((block != zero) \& one)}
+\State \textbf{[...]}
+\State \textbf{case NBLOCK-1:} \tt{...}
+\end{algorithmic}
+
+% The MMX technology permits to work on $8$ bytes in parallel while the SSE
+% works in $16$ bytes.  As the rest of the CPU works on $8$ bytes, the SSE
+% technology needs some constraint on memory access than cannot be fulfilled in
+% our algorithm.
+
+% This motivate our choice to use the MMX technology. 
+% More precisely, we use three commands: \tt{pcmpeqb}, \tt{pandn} and \tt{psubb}.
+%  These commands work on two arrays of $8$ bytes, called~\tt{s} and \tt{d} here. 
+%  In each case the array \tt{d} is modified. 
+%  We denote bytes by $\{\tt{0},\tt{1}\}$-words of length $8$.
+%  After a call to \tt{pcmpeqb(s,d)} the $i$th entry of $d$ contains \tt{11111111} if $\tt{s}[i]=\tt{d}[i]$ holds and \tt{00000000} otherwise. 
+%  The command \tt{pand(s,d)}  store in $\tt{d}[i]$  the value of the byte-logic operation \tt{s}[i] \tt{and} (\tt{not} \tt{d}[i]).
+%   When \tt{pand(s,d)} is called the new value of $\tt{d}[i]$ is $\tt{s}[i]-\tt{d}[i]$.  
+%   Glueing all the pieces together we obtain the following version of the \textbf{for} loop of Algorithm~\textsc{Son} :
+
+% \begin{algorithmic}[1]
+% \State $\tt{src}\gets \text{address}(\tt{S.d[0]})$
+% \State $\tt{dst}\gets \text{address}(\tt{T.d[x]})$
+% \State $\tt{i}\gets 0$
+% \While{$\tt{i}\leq \tt{3G}-\tt{x}$}
+% \State $\tt{t}\gets [\tt{00000000},...,\tt{00000000}]$ \Comment{$8$ bytes equal to \tt{00000000}}
+% \State \tt{pcmpeqb(\tt{src},t)}
+% \State \tt{pandn([00000001,...,00000001],t)}
+% \State \tt{psub(\tt{dst},t)}
+% \State $\tt{dst}\gets \tt{t}$ \Comment{copy the array $[\tt{t}[0],...,\tt{t}[7]]$ to $[\tt{d}[0],...,\tt{d}[7]]$}
+% \State \textbf{increase} \tt{src},\tt{dst},\tt{i} \textbf{by} $8$
+% \EndWhile
+% \end{algorithmic}
+
+% Let us explain how this works in more details. 
+% Let $i$ be an integer in $\{0,...,7\}$.  
+% After Line~\tt{5}, the value of $\tt{t}[i]$ is \tt{0000000}. 
+% After Line~\tt{6}, we have 
+% \[
+% \tt{t}[i]=\begin{cases}
+% \tt{11111111}&\text{if $\tt{src}[i]=\tt{00000000}$ (corresponding to $0$)}\\
+% \tt{00000000}&\text{otherwise}
+% \end{cases}
+% \]
+% Line \tt{7} performs a logical \tt{and} between $\tt{00000001}$ and $\tt{not t}[i]$. Hence the result byte is \tt{00000001} if the last byte of \tt{t}[i] is $0$ and \tt{00000000} otherwise: 
+% \[
+% \tt{t}[i]=\begin{cases}
+% \tt{00000000}&\text{if $\tt{src}[i]=\tt{00000000}$ (corresponding to $0$)}\\
+% \tt{00000001}&\text{otherwise}
+% \end{cases}
+% \]
+%  Line \tt{8} subtracts \tt{t}[i] (which is equal to $0$ or $1$) from $\tt{dst}[i]$ according to Proposition~\ref{P:Sx}. 
+% Finally we shift the cursor \tt{src} and \tt{dst} by eight cases.
+
+\subsection{Parallel tree exploration using \texttt{Cilk+}}
+
+Our second optimization is to use parallelism on exploration of the tree.
+Today, CPU of personal computer have several cores (2, 4 or more).  The given
+version of our exploration algorithm uses a single core and so a fraction only
+of the power of a CPU. The idea here is that different branches of the tree can
+be explored in parallel by different cores of the computer. The tricky part is
+to ensure that each cores are busy, giving a new branch when a core is done
+with a former one. Fortunately there is a technology calle \texttt{Cilk+}
+which is particularly well suited for those kind of problems.
+
+\tt{Cilk} is a general-purpose language designed for multithreaded parallel
+computing. The \tt{C++} incarnation is called \tt{Cilk+}. The biggest
+principle behind the design of the Cilk language is that the programmer should
+be responsible for \emph{exposing} the parallelism, identifying elements that
+can safely be executed in parallel; it should then be left to the run-time
+environment, particularly the scheduler, to decide during execution how to
+actually divide the work between processors.
+
+The crucial thing is that two keywords are all that are needed to start
+using the parallel features of \tt{Cilk+}:
+\begin{itemize}
+\item\tt{cilk\_spawn}: used on a procedure call, indicates that the call can
+  safely operate in parallel with the remaining code if the current
+  function. Note that the scheduler is not obligated to run this procedure in
+  parallel; the keyword merely alerts the scheduler that it can do so.
+\item\tt{cilk\_sync}: indicates that execution of the current procedure cannot
+  proceed until all previously spawned procedures have completed and returned
+  their results to the current frame.
+\end{itemize}
+As a consequence, to get a parallel version of the recursive
+Algorithm~\ref{A:ExpRec}, you only need to modify it as
 \begin{algorithmic}[1]
-\State $\tt{src}\gets \text{address}(\tt{S.d[0]})$
-\State $\tt{dst}\gets \text{address}(\tt{T.d[x]})$
-\State $\tt{i}\gets 0$
-\While{$\tt{i}\leq \tt{3G}-\tt{x}$}
-\State $\tt{t}\gets [\tt{00000000},...,\tt{00000000}]$ \Comment{$8$ bytes equal to \tt{00000000}}
-\State \tt{pcmpeqb(\tt{src},t)}
-\State \tt{pandn([00000001,...,00000001],t)}
-\State \tt{psub(\tt{dst},t)}
-\State $\tt{dst}\gets \tt{t}$ \Comment{copy the array $[\tt{t}[0],...,\tt{t}[7]]$ to $[\tt{d}[0],...,\tt{d}[7]]$}
-\State \textbf{increase} \tt{src},\tt{dst},\tt{i} \textbf{by} $8$
-\EndWhile
+\Procedure{ExploreRec}{\tt{S}, \tt{G}}
+	\If{$g(\tt{S})< \tt{G}$}
+		\For{\tt{x} from $c(\tt{S})$ to $c(\tt{S})+m(\tt{S})$}
+			\If{$\tt{x}\in \Irr(\tt{S})$}
+				\State\textbf{CilkSpawn} \Call{ExploreRec}{$\tt{S}^\tt{x}, \tt{G}$}
+			\EndIf
+		\EndFor
+	\EndIf
+\EndProcedure
 \end{algorithmic}
-
-Let us explain how this works in more details. 
-Let $i$ be an integer in $\{0,...,7\}$.  
-After Line~\tt{5}, the value of $\tt{t}[i]$ is \tt{0000000}. 
-After Line~\tt{6}, we have 
-\[
-\tt{t}[i]=\begin{cases}
-\tt{11111111}&\text{if $\tt{src}[i]=\tt{00000000}$ (corresponding to $0$)}\\
-\tt{00000000}&\text{otherwise}
-\end{cases}
-\]
-Line \tt{7} performs a logical \tt{and} between $\tt{00000001}$ and $\tt{not t}[i]$. Hence the result byte is \tt{00000001} if the last byte of \tt{t}[i] is $0$ and \tt{00000000} otherwise: 
-\[
-\tt{t}[i]=\begin{cases}
-\tt{00000000}&\text{if $\tt{src}[i]=\tt{00000000}$ (corresponding to $0$)}\\
-\tt{00000001}&\text{otherwise}
-\end{cases}
-\]
- Line \tt{8} subtracts \tt{t}[i] (which is equal to $0$ or $1$) from $\tt{dst}[i]$ according to Proposition~\ref{P:Sx}. 
-Finally we shift the cursor \tt{src} and \tt{dst} by eight cases.
-
-Our second optimization  is to use parallelism on exploration of the tree.
-Today, CPU of personal computer have several cores (2, 4 or more). 
-The given version of our exploration algorithm  uses a single core and so a fraction only of the power of a CPU.
-
-Our method of parallelism is very simple: for $G\in \NN$,  we cut the tree $\mathcal{T}^{\leq G}$ in sub-trees $\mathcal{T}_1,...,\mathcal{T}_n$ and we launch our algorithm on these sub-trees. 
-The advantage of this method is that there is no communication between the instances of our algorithm. 
-The disadvantage is that the cutting controls the efficiency of the parallelism. 
-Assume for example that we want to explore the tree~$\mathcal{T}^{\leq 40}$.
-We first determine all numerical semigroups $S_1,...S_{n_{20}}$ of genus $20$. 
-To explore $\mathcal{T}$ it remains to explore the tree $\mathcal{T}_i$ rooted in $S_i$ for $i=1,...,n_{20}$. 
-This works but the time to explore $\mathcal{T}_1$ is similar to the time need to explore the tree $\mathcal{T}^{\leq 40}$. 
-And in this case, using many cores does not reduce the time to explore the tree.
-
-Let us now explain in more detail how we cut the tree $\mathcal{T}^{\leq G}$ in order to use parallelism.
-Semigroups of the form $\left<g+1,...,2g+1\right>$, which are of genus $g$, are called \emph{ordinary} in~\cite{Elizalde}. 
-Each ordinary semigroup has a unique son that is also ordinary. 
-We define $X$ to be the set of all the non-ordinary sons of an ordinary semigroup and $X^{\leq G}$ the restriction of $X$ to semigroup of genus $\leq G$. 
-We then denote by  $\mathcal{T}_i$ the tree rooted on $S_i$ where $X^{\leq G}=\{S_1,...,S_n\}$. 
-The time needed to explore $\mathcal{T}_i$ is very heterogeneous but there are many tree $T_i$ with maximal time. 
-This cutting is more efficient than the previous one.
+We just tells \tt{Cilk+} that the subtrees rooted at various children can be
+explored in parallel.
+
+The thing are actually only a little bit more complicated. First we have to
+gather the results of the exploration. If we simply write
+\begin{algorithmic}[1]
+\State \tt{result}[$g(\tt{S}$] $\gets \tt{result}[g(\tt{S}] + 1$
+\end{algorithmic}
+then we face the problem of to core incrementing the same variable at the same
+time. This is called a data race and leads to nondeterministic wrong
+result. Cilk provide the notion of \emph{reducer} which are variables local to
+each thread which are gathered (here added) when a thread is finishing its
+job.
+
+A more important problem is that the cost of a recursive call in non
+negligible. Using \tt{Cilk+} recursive calls makes it even worse. The solution
+we use is to switch back to the non recursive version using a stack when the
+genus is close to the target genus. This leads to the following \tt{Cilk+} code:
+\begin{listing}{1}
+void explore(const monoid &m) {
+  unsigned long int nbr = 0;
+  if (m.genus < MAX_GENUS - STACK_BOUND) {
+    auto it = generator_iter<CHILDREN>(m);
+    while (it.move_next()) {
+      auto child = remove_generator(m, it.get_gen()).
+      cilk_spawn explore(child);
+      nbr++;
+    }
+    cilk_results[m.genus] += nbr;
+  }
+  else explore_stack(m, cilk_results.get_array());
+}
+\end{listing}
+Note that in our version, we found that \tt{STACK\_BOUND} optimal value was
+around 10-12 for genus un the range $45\dots 67$ so that \tt{explore\_stack}
+is used more than $99\%$ of the time, \tt{Cilk+} is used to balance the work
+between the different cores.
+
+\subsection{Various technical optimizations}
+
+Avoiding all dynamic allocation during the computation: everything is computed
+``in place'';
+
+% Our method of parallelism is very simple: for $G\in \NN$,  we cut the tree $\mathcal{T}^{\leq G}$ in sub-trees $\mathcal{T}_1,...,\mathcal{T}_n$ and we launch our algorithm on these sub-trees. 
+% The advantage of this method is that there is no communication between the instances of our algorithm. 
+% The disadvantage is that the cutting controls the efficiency of the parallelism. 
+% Assume for example that we want to explore the tree~$\mathcal{T}^{\leq 40}$.
+% We first determine all numerical semigroups $S_1,...S_{n_{20}}$ of genus $20$. 
+% To explore $\mathcal{T}$ it remains to explore the tree $\mathcal{T}_i$ rooted in $S_i$ for $i=1,...,n_{20}$. 
+% This works but the time to explore $\mathcal{T}_1$ is similar to the time need to explore the tree $\mathcal{T}^{\leq 40}$. 
+% And in this case, using many cores does not reduce the time to explore the tree.
+
+% Let us now explain in more detail how we cut the tree $\mathcal{T}^{\leq G}$ in order to use parallelism.
+% Semigroups of the form $\left<g+1,...,2g+1\right>$, which are of genus $g$, are called \emph{ordinary} in~\cite{Elizalde}. 
+% Each ordinary semigroup has a unique son that is also ordinary. 
+% We define $X$ to be the set of all the non-ordinary sons of an ordinary semigroup and $X^{\leq G}$ the restriction of $X$ to semigroup of genus $\leq G$. 
+% We then denote by  $\mathcal{T}_i$ the tree rooted on $S_i$ where $X^{\leq G}=\{S_1,...,S_n\}$. 
+% The time needed to explore $\mathcal{T}_i$ is very heterogeneous but there are many tree $T_i$ with maximal time. 
+% This cutting is more efficient than the previous one.
 

### 2014-09-02 Papier/article.tex
article.tex

@@
 
-Though there are asymptotically faster algorithm than the one presented here,
+Though there are asymptotically faster algorithms than the one presented here,
 thank to careful optimizations, we were able to compute $n_g$ for much larger
-genus than before. This is due to the fact that our algorithm is very well
-suited for the current processor architecture. In particular, it allows to use
-parallelism at various scale (parallel branch exploration, vectorization)...
-
-To get the greatest speed from modern processor we used several optimization
+genus than before. This is due to the fact that our algorithm is particularly
+well suited for the current processor architecture. In particular, it allows
+to use parallelism at various scale (parallel branch exploration,
+vectorization)...
+
+To get the greatest speed from modern processor, we used several optimization
 tricks, which we will elaborate in the following section:
@@
   enumerating tree branching;
-\item Un-recursived algorithm using a stack;
+\item Partially derecursived algorithm using a stack;
 \item Avoiding all dynamic allocation during the computation: everything is
@@
 \item Avoiding all unnecessary copy (the cost of the \textsc{Son} algorithm is
-  roughly the same (even less) as for as copy !
-\item Aggressive loop unrolling. The main loop is unrolled by hand using some
+  roughly the same as copying) !
+\item Aggressive loop unrolling: the main loop is unrolled by hand using some
   kind of Duff's device.
@@
 
-Assume for example that we want to construct the tree $\mathcal{T}^{\leq 100}$ of all  numerical semigroup of  genus smaller than $100$. 
-In this case, the representation of  numerical  semigroup given in Section~\ref{S:DecNumber} uses decomposition numbers of integers smaller than~$300$.
-By Lemma~\ref{L:RepN}, such a decomposition number is smaller than $151$ and requires~$1$ byte of memory.
-Thus at each \textbf{for} step of Algorithm~\textsc{Son}, the CPU actually works on $1$ byte. 
-However current CPUs work on $8$ bytes.
-The first optimization uses this point.
-
-To go further we must specify that the array of decomposition numbers in the representation of a semigroup corresponds to consecutive bytes in memory. 
-In the \textbf{for} loop of Algorithm~\textsc{Son} we may imagine two cursors: the first one, denoted \texttt{src}  pointing to the memory byte of $\texttt{S.d[0]}$  and the second one, denoted \texttt{dst} pointing to the memory byte $\texttt{T.d[y]}$.
-Using these two cursors, Lines \tt{10} to \tt{14} of Algorithm~\textsc{Son} can be rewritten as follows
+Assume for example that we want to construct the tree $\mathcal{T}^{\leq 100}$
+of all numerical semigroup of genus smaller than $100$.  In this case, the
+representation of numerical semigroup given in Section~\ref{S:DecNumber} uses
+decomposition numbers of integers smaller than~$300$.  By Lemma~\ref{L:RepN},
+such a decomposition number is smaller than $151$ and requires~$1$ byte of
+memory.  Thus at each \textbf{for} step of Algorithm~\textsc{Son}, the CPU
+actually works on $1$ byte. However current CPUs usually work on $8$ bytes and
+even on $16$~bytes using vector extensions. The first optimization uses this
+point.
+
+To go further we must specify that the array of decomposition numbers in the
+representation of a semigroup corresponds to consecutive bytes in memory. In
+the \textbf{for} loop of Algorithm~\textsc{Son} we may imagine two cursors:
+the first one, denoted \texttt{src} pointing to the memory byte of
+$\texttt{S.d[0]}$ and the second one, denoted \texttt{dst} pointing to the
+memory byte $\texttt{T.d[y]}$.  Using these two cursors, Lines \tt{10} to
+\tt{14} of Algorithm~\textsc{Son} can be rewritten as follows
 
@@
 without collision.  Current CPUs allow this thanks to the SIMD technologies as
-MMX, SSE, etc.  The acronym SIMD stands for Single Operation Multiple Data.
-We used SSE4.1 technology as it allows for the largest speedup. This need to
-respect some constraints in the memory organization of the data, namely
-``memory alignment''. Recall that an address is 16~bytes aligned if it is a
-multiple of 16. SSE memory access are much faster for aligned access.
+MMX, SSE, etc.  The acronym SIMD~\cite{SIMD} stands for Single Operation
+Multiple Data.  We used SSE4.1~\cite{SSE, IntelSSE} technology as it allows
+for the largest speedup. This need to respect some constraints in the memory
+organization of the data, namely ``memory alignment''. Recall that an address
+is 16~bytes aligned if it is a multiple of 16. SSE memory access are much
+faster for aligned access.
 
@@
 Then the core of the \textbf{while} loop in the preceding algorithm is
-translated as a for loop as follows (recall that $x$ denotes the generator if the
-father which is to be removed in the children):
+translated as a for loop as follows (recall that $x$ denotes the generator of
+the father which is to be removed in the children):
 \begin{listing}{1}
@@
 \end{algorithmic}
+which is equivalent to Lines \tt{10} to \tt{14} of Algorithm~\textsc{Son}.
+\medskip
+
 As we previously said, to gain more speed this core loop is actually unrolled
-using some kind of Duff device as
+using some kind of Duff device~\cite{Duff} as
 \begin{algorithmic}
@@
 to ensure that each cores are busy, giving a new branch when a core is done
-with a former one. Fortunately there is a technology calle \texttt{Cilk+}
+with a former one. Fortunately there is a technology called
+\texttt{Cilk+}~\cite{CilkIntel, CilkRefman}
 which is particularly well suited for those kind of problems.
@@
 \end{algorithmic}
-then we face the problem of to core incrementing the same variable at the same
-time. This is called a data race and leads to nondeterministic wrong
-result. Cilk provide the notion of \emph{reducer} which are variables local to
-each thread which are gathered (here added) when a thread is finishing its
-job.
+then we face the problem of two core incrementing the same variable at the
+same time. Incrementing a variables is actually done in 3 steps: reading the
+value from the memory, adding one, storing back the results. Since there is by
+default non synchronization, the following sequence of actions for two core
+and is possible: Read1 / Read2 / Add1 / Add2 / Store1 / Store2. Then the two
+cores performs the same modification resulting in incrementing the variable
+only once. This is called a data race and leads to nondeterministic wrong
+result. To cope with those kinds of synchronization problems, Cilk provide the
+notion of \emph{reducer} which are variables local to each thread which are
+gathered (here added) when a thread is finishing its job.
 
 A more important problem is that the cost of a recursive call in non
-negligible. Using \tt{Cilk+} recursive calls makes it even worse. The solution
-we use is to switch back to the non recursive version using a stack when the
-genus is close to the target genus. This leads to the following \tt{Cilk+} code:
+negligible. Using \tt{Cilk+} recursive calls instead of \tt{C++} calls makes
+it even worse. The solution we use is to switch back to the non recursive
+version using a stack when the genus is close to the target genus. This leads
+to the following \tt{Cilk+} code:
 \begin{listing}{1}
@@
 Note that in our version, we found that \tt{STACK\_BOUND} optimal value was
-around 10-12 for genus un the range $45\dots 67$ so that \tt{explore\_stack}
-is used more than $99\%$ of the time, \tt{Cilk+} is used to balance the work
-between the different cores.
+around 10-12 for genus in the range $45\dots 67$ so that \tt{explore\_stack}
+is used more than $99\%$ of the time. The \tt{Cilk+} recursive function does
+actually very little work but ensure that the work is balanced between the
+different cores.
 
@@
 
-Avoiding all dynamic allocation during the computation: everything is computed
-``in place'';
-
-% Our method of parallelism is very simple: for $G\in \NN$,  we cut the tree $\mathcal{T}^{\leq G}$ in sub-trees $\mathcal{T}_1,...,\mathcal{T}_n$ and we launch our algorithm on these sub-trees. 
-% The advantage of this method is that there is no communication between the instances of our algorithm. 
-% The disadvantage is that the cutting controls the efficiency of the parallelism. 
-% Assume for example that we want to explore the tree~$\mathcal{T}^{\leq 40}$.
-% We first determine all numerical semigroups $S_1,...S_{n_{20}}$ of genus $20$. 
-% To explore $\mathcal{T}$ it remains to explore the tree $\mathcal{T}_i$ rooted in $S_i$ for $i=1,...,n_{20}$. 
-% This works but the time to explore $\mathcal{T}_1$ is similar to the time need to explore the tree $\mathcal{T}^{\leq 40}$. 
-% And in this case, using many cores does not reduce the time to explore the tree.
-
-% Let us now explain in more detail how we cut the tree $\mathcal{T}^{\leq G}$ in order to use parallelism.
-% Semigroups of the form $\left<g+1,...,2g+1\right>$, which are of genus $g$, are called \emph{ordinary} in~\cite{Elizalde}. 
-% Each ordinary semigroup has a unique son that is also ordinary. 
-% We define $X$ to be the set of all the non-ordinary sons of an ordinary semigroup and $X^{\leq G}$ the restriction of $X$ to semigroup of genus $\leq G$. 
-% We then denote by  $\mathcal{T}_i$ the tree rooted on $S_i$ where $X^{\leq G}=\{S_1,...,S_n\}$. 
-% The time needed to explore $\mathcal{T}_i$ is very heterogeneous but there are many tree $T_i$ with maximal time. 
-% This cutting is more efficient than the previous one.
+Using vectorization and loop enrolling as described previously leads to an
+extremely fast \textsc{Son} algorithm. Indeed, its cost is comparable to the cost
+of copying a monoid. It is therefore crucial for performance to avoid any
+extra cost. We list here various places where unnecessary cost can be avoided.
+
+\subsubsection*{Avoiding dynamic allocation} Compared to the \textsc{Son}
+algorithm, dynamic allocation cost orders of magnitude more. Therefore, during
+the derecursived stack algorithm, we only allocate (on the system stack rather
+than on the heap), the stack of monoid. No further allocation are done.
+
+\subsubsection*{Avoiding all unnecessary copy} We also used a trick to avoid
+copying from and to the top of the stack. Indeed, then main loop perform the
+following sequences of operations:
+\begin{algorithmic}[1]
+  \State $\tt{S} \gets \tt{stack.top()}$
+  \State \tt{stack.pop()}
+  \For{all children $\tt{S}^\tt{x}$ of \tt{S}}
+  \State $\tt{S.push}(\tt{S}^\tt{x})$
+  \EndFor
+\end{algorithmic}
+If we use a stack of monoid, we can construct $\tt{S}^\tt{x}$ directly into
+the stack memory but we have to copy the top of the stack to \tt{S}. Since the
+average number of children is the golden ratio, this single copy in far from
+being negligible. The trick is to use a level of indirection, replacing the
+stack by an array of monoid and an array of indexes pointing to the array of
+monoid. We keep the invariant that the array of indexes is a permutation of
+the array of monoid. Now instead of copying \tt{S} out of the stack, we keep
+it on the stack, pushing the children in the second position by exchanging the
+indexes. Here is the relevant part of the code:
+\begin{listing}{1}
+  monoid data[MAX_GENUS-1], *stack[MAX_GENUS], *current;
+  monoid **stack_pointer = stack + 1;
+  for (ind_t i=1; i<MAX_GENUS; i++) stack[i] = &(data[i-1]);
+  [...]
+  while (stack_pointer != stack) {
+      --stack_pointer;
+      current = *stack_pointer;
+      [...] for each children {
+        *stack_pointer = *(stack_pointer + 1);
+        [...] construct the children in **stack_pointer
+        [...] using the parent in *current
+        stack_pointer++;
+        [...]
+      }
+      *stack_pointer = current;
+  }
+\end{listing}
+
+\subsubsection*{Pointer arithmetic and indexes}
+
+Due to the way, the C++ does its pointer arithmetic, even if the index in the
+array are less than $3G$ and therefore fits in $8$~bits, we use $64$~bits
+indexes (namely \verb|uint_fast64_t|) to avoid conversion and sign extension
+when computing addresses of indexed element. This single standard trick save
+20\% of speed.
+
+\subsection{Timings}
 

### 2014-09-02 src/Cilk++/monoid.hpp

@@ -46,9 +46,12 @@ inline void remove_generator(monoid &__restrict__ dst,
 inline monoid remove_generator(const monoid &src, ind_t gen);
 
 
-typedef enum { ALL, CHILDREN } generator_type;
+// typedef enum { ALL, CHILDREN } generator_type;
+class ALL {};
+class CHILDREN {};
 
-template <generator_type T> class generator_iter
+// template <generator_type T> class generator_iter
+template <class T> class generator_iter
 {
 private:
 
@@ -149,7 +152,7 @@ template<> inline generator_iter<CHILDREN>::generator_iter(const monoid &mon)
   gen = (iblock << 4) - 1;
 };
 
-template <generator_type T> inline uint8_t generator_iter<T>::count()
+template <class T> inline uint8_t generator_iter<T>::count()
 {
   uint8_t nbr = _mm_popcnt_u32(mask); // popcnt returns a 8 bits value
   for (iblock++; iblock < bound; iblock++)
@@ -157,7 +160,7 @@ template <generator_type T> inline uint8_t generator_iter<T>::count()
   return nbr;
 };
 
-template <generator_type T> inline bool generator_iter<T>::move_next()
+template <class T> inline bool generator_iter<T>::move_next()
 {
   while (!mask)
     {

### 2014-09-02 src/Cilk++/setup.py

@@ -19,7 +19,7 @@ from sage.env import *
 
 SAGE_INC = os.path.join(SAGE_LOCAL, 'include')
 SAGE_C   = os.path.join(SAGE_SRC, 'c_lib', 'include')
-SAGE_DEV = os.path.join(SAGE_ROOT, 'devel', 'sage-main')
+SAGE_DEV = os.path.join(SAGE_ROOT, 'src')
 
 
 if platform.system()=="Darwin":

### 2014-09-03 src/others/ngtree-v1/main.cpp
Add other programs for time computation graph

@@ -1,0 +1,97 @@
+#include <iostream>
+#include <list>
+#include <stdlib.h>
+
+using namespace std;
+
+typedef unsigned char uchar; 
+typedef unsigned short int usint;
+
+class SemiGroup{
+public:
+  uchar* tab;
+  uchar c,m;
+  SemiGroup();
+  SemiGroup(const SemiGroup& S,uchar x);
+  ~SemiGroup();
+
+};
+
+SemiGroup::SemiGroup(){
+  c=m=1;
+  tab=new uchar[c+m];
+  tab[0]=2;
+  tab[1]=1;
+}
+
+SemiGroup::~SemiGroup(){
+  delete[] tab;
+}
+
+SemiGroup::SemiGroup(const SemiGroup& S,uchar x){
+  usint i,j;
+  uchar nb_irr=0;
+  uchar s=S.m+S.c;
+  c=x+1;
+  m=(x==S.m)?m=S.m+1:S.m;
+  uchar irr[m];
+  uchar v;
+  tab=new uchar[c+m];
+  for(i=0;i<x;++i){
+    v=S.tab[i];
+    tab[i]=v;
+    if(v==1){
+      irr[nb_irr++]=i;
+    }
+  }
+  tab[x]=0;
+  for(i=x+1;i<c+m;++i){
+    v=(i<s)?S.tab[i]:2;
+    if(v==1){
+      tab[i]=1;
+      irr[nb_irr++]=i;
+    }
+    else{
+      for(j=0;j<nb_irr;++j){
+	if(tab[i-irr[j]]!=0) break;
+      }
+      if(j==nb_irr){
+	tab[i]=1;
+	irr[nb_irr++]=i;
+      }
+      else tab[i]=2;
+    }
+  }
+}
+
+int main(int argc,char** argv){
+  if(argc!=2){
+    cout<<"You must sepcify uniquley the genus"<<endl;
+    return -1;
+  }
+  int g,G;
+  uchar c,m,x;
+  SemiGroup *S,*Sx;
+  list<SemiGroup*> lg;
+  list<SemiGroup*> next;
+  list<SemiGroup*>::iterator it;
+  G=atoi(argv[1]);
+  lg.push_back(new SemiGroup);
+  for(g=1;g<=G;++g){
+    for(it=lg.begin();it!=lg.end();++it){
+      S=*it;
+      c=S->c;
+      m=S->m;
+      for(x=c;x<c+m;++x){
[…]

### 2014-09-03 src/others/ngtree-v2/main.cpp

@@ -1,0 +1,102 @@
+#include <iostream>
+#include <list>
+#include <stdlib.h>
+
+using namespace std;
+
+typedef unsigned char uchar; 
+typedef unsigned short int usint;
+typedef unsigned long long integer;
+
+class SemiGroup{
+public:
+  uchar* tab;
+  uchar c,g,m;
+  SemiGroup();
+  SemiGroup(const SemiGroup& S,uchar x);
+  ~SemiGroup();
+};
+
+SemiGroup::SemiGroup(){
+  c=m=1;
+  g=0;
+  tab=new uchar[c+m];
+  tab[0]=2;
+  tab[1]=1;
+}
+
+SemiGroup::~SemiGroup(){
+  delete[] tab;
+}
+
+SemiGroup::SemiGroup(const SemiGroup& S,uchar x){
+  usint i,j;
+  uchar nb_irr=0;
+  uchar s=S.m+S.c;
+  g=S.g+1;
+  c=x+1;
+  m=(x==S.m)?m=S.m+1:S.m;
+  uchar irr[m];
+  uchar v;
+  tab=new uchar[c+m];
+  for(i=0;i<x;++i){
+    v=S.tab[i];
+    tab[i]=v;
+    if(v==1){
+      irr[nb_irr++]=i;
+    }
+  }
+  tab[x]=0;
+  for(i=x+1;i<c+m;++i){
+    v=(i<s)?S.tab[i]:2;
+    if(v==1){
+      tab[i]=1;
+      irr[nb_irr++]=i;
+    }
+    else{
+      for(j=0;j<nb_irr;++j){
+	if(tab[i-irr[j]]!=0) break;
+      }
+      if(j==nb_irr){
+	tab[i]=1;
+	irr[nb_irr++]=i;
+      }
+      else tab[i]=2;
+    }
+  }
+}
+
+int main(int argc,char** argv){
+  if(argc!=2){
+    cout<<"You must sepcify uniquley the genus"<<endl;
+    return -1;
+  }
+  int g,i,G;
+  uchar c,m,x;
+  SemiGroup *S,*Sx;
+  list<SemiGroup*> st;
+  G=atoi(argv[1]);
+  integer ng[G+1];
+  for(i=0;i<=G;++i) ng[i]=0;
+  st.push_back(new SemiGroup);
+  ng[0]=1;
+  while(not st.empty()){
+    S=st.back();
+    st.pop_back();
[…]

### 2014-09-03 src/others/ngtree-v3/main.cpp

@@ -1,0 +1,82 @@
+#include <iostream>
+#include <list>
+#include <stdlib.h>
+#include <string.h>
+
+using namespace std;
+
+typedef unsigned char uchar; 
+typedef unsigned short int usint;
+typedef unsigned long long integer;
+
+usint size;
+
+class SemiGroup{
+public:
+  uchar* tab;
+  uchar c,g,m;
+  SemiGroup();
+  SemiGroup(const SemiGroup& S,uchar x);
+  ~SemiGroup();
+};
+
+SemiGroup::SemiGroup(){
+  c=m=1;
+  g=0;
+  tab=new uchar[size];
+  for(int i=0;i<size;++i) tab[i]=i/2+1;
+}
+
+SemiGroup::~SemiGroup(){
+  delete[] tab;
+}
+
+SemiGroup::SemiGroup(const SemiGroup& S,uchar x){
+  usint i;
+  g=S.g+1;
+  c=x+1;
+  m=(x==S.m)?m=S.m+1:S.m;
+  uchar v;
+  tab=new uchar[size];
+  memcpy(tab,S.tab,size);
+  tab[x]=0;
+  for(i=x+1;i<size;++i){
+    if(S.tab[i-x]>0) --tab[i];
+  }
+}
+
+int main(int argc,char** argv){
+  if(argc!=2){
+    cout<<"You must sepcify uniquely the genus"<<endl;
+    return -1;
+  }
+  int g,i,G;
+  uchar c,m,x;
+  SemiGroup *S,*Sx;
+  list<SemiGroup*> st;
+  G=atoi(argv[1]);
+  size=3*G;
+  integer ng[G+1];
+  for(i=0;i<=G;++i) ng[i]=0;
+  st.push_back(new SemiGroup);
+  ng[0]=1;
+  while(not st.empty()){
+    S=st.back();
+    st.pop_back();
+    if(S->g<G){
+      c=S->c;
+      m=S->m;
+      for(x=c;x<c+m;++x){
+	if(S->tab[x]==1){
+	  Sx=new SemiGroup(*S,x);
+	  ng[Sx->g]++;
+	  st.push_back(Sx);
+	}
+      }
+    }
+    delete S;
+  }
+  /*for(i=0;i<=G;++i){
+    cout<<i<<"->"<<ng[i]<<endl;
+    }*/
+}

### 2014-09-03 src/others/ngtree-v4/config.hpp

@@ -1,0 +1,4 @@
+#define GMAX 40
+#define MMAX (GMAX+1)
+#define SIZE 3*GMAX+3
+#define STACK_CAPACITY MMAX*(GMAX+1)

### 2014-09-03 src/others/ngtree-v4/main-asm.cpp

@@ -1,0 +1,61 @@
+#include <iostream>
+#include "stack.hpp"
+#include <string.h>
+
+using namespace std;
+
+
+int main(){
+  Stack stack;
+  init(*stack.data[0]);
+  stack.size=1;
+  byte i,g,m,x,s,t;
+  ulong ng[GMAX+1];
+  for(i=0;i<=GMAX;++i) ng[i]=0;
+  ng[1]=1;
+  byte *data,*ndata;
+  SemiGroup *sg,*nsg;
+  while(stack.size>0){
+    sg=stack.pop(); 
+    g=sg->g;
+    if(g<GMAX){
+      ++g;
+      t=0;
+      data=sg->data;
+      s=(x=sg->c)+(m=sg->m);
+      for(;x<s;++x){
+	if(data[x]==1){
+	  ++t;
+	  nsg=stack.push();
+	  ndata=nsg->data;
+	  memcpy(ndata,data,SIZE);
+	  ndata[x]=0;
+	  nsg->c=x+1;
+	   __asm__ ("movq %0,%%rax \n"           //&tab[1]--> rax
+	  "movq %1,%%rdx \n"           //&ntab[x+1] --> rdx
+	  "mov %2,%%cl \n"             //size-x-1->cl
+	  "LOOP: \n"                   //start of the loop
+	  "cmpb $0,(%%rax) \n"         //compare *rax with 0 ?
+	  "jz END_LOOP \n"             //if *rax==0 goto REM_END_LOOP
+	  "decb (%%rdx) \n"            //else (*rdx)--;
+	  "END_LOOP: \n"           
+	  "incq %%rax \n"              //rax++
+	  "incq %%rdx \n"              //rdx++
+	  "decb %%cl\n"                //cl--
+	  "jnz LOOP\n"                 //if cl!=0 goto REM_LOOP
+	  :
+	  :"g"(&data[1]),"g"(&ndata[x+1]),"g"(byte(SIZE-x-1))
+	  :"rax","cl","rdx"
+	  );
+	  nsg->m=(x==m)?x+1:m;
+	  nsg->g=g;
+	}
+      }
+      ng[g]+=t;
+    } 
+  }
+  for(i=0;i<=GMAX;++i){
+    cout<<i<<" -> "<<ng[i]<<endl;
+  }
+  
+}

### 2014-09-03 src/others/ngtree-v4/main-mmx.cpp

@@ -1,0 +1,85 @@
+#include <iostream>
+#include "stack.hpp"
+#include <string.h>
+#include <cstdlib>
+
+using namespace std;
+
+int main(int argc,char** argv){
+  cout<<"Computing until genus "<<GMAX<<endl;
+  Stack stack;
+  int i;
+  init(*stack.data[0]);
+  stack.size=1;
+
+  //Init mmx
+  __asm__("movq $0x0101010101010101,%rax \n"
+	  "movq %rax,%mm2 \n"
+	  "movq $0x0000000000000000,%rax \n" 
+	  "movq %rax,%mm3 \n");
+  byte g,m,y,x,s,t;
+  ulong ng[GMAX+1];
+  for(i=0;i<=GMAX;++i) ng[i]=0;
+  ng[1]=1;
+  byte *data,*ndata;
+  SemiGroup *sg,*nsg;
+  while(stack.size>0){
+    sg=stack.pop(); 
+    g=sg->g;
+    if(g<GMAX){
+      ++g;
+      t=0;
+      data=sg->data;
+      s=(x=sg->c)+(m=sg->m);
+      for(;x<s;++x){
+	if(data[x]==1){
+	  ++t;
+	  nsg=stack.push();
+	  ndata=nsg->data;
+	  memcpy(ndata,data,SIZE);
+	  ndata[x]=0;
+	  nsg->c=x+1;
+	  nsg->m=(x==m)?x+1:m;
+	  nsg->g=g;
+	  __asm__(//"mms \n"
+	  "movq %0,%%rax \n"            //&tab[1] --> roax
+	  "movq %1,%%rdx \n"            //&ntab[x+1] --> rdx
+	  "movb %2,%%cl \n"             //size-x-1 --> cl
+	  "LOOP_MMX: \n"                //start of the loop using mmx operation
+	  "movq %%mm3,%%mm0 \n"         //0x00|00|00|00|00|00|00|00 --> mm0
+	  "pcmpeqb (%%rax),%%mm0 \n"    //test per byte if *rax==0 
+	                                //if true the corresping byte in mm0 is 0xFF else is 0x00 
+	  "pandn %%mm2,%%mm0 \n"        //0x01|01|01|01|01|01|01|01 and (not mm0)--> mm0
+	  "movq (%%rdx),%%mm1 \n"       //*rdx --> mm1
+	  "psubb %%mm0,%%mm1 \n"        //mm1-mm0 --> mm1
[…]

### 2014-09-03 src/others/ngtree-v4/main-sse.cpp

@@ -1,0 +1,89 @@
+#include <iostream>
+#include "stack.hpp"
+#include <string.h>
+
+using namespace std;
+
+
+int main(){
+  ulong c1[2]={0x0101010101010101,0x0101010101010101};
+  ulong c2[2]={0,0};
+  __asm__("movq %0,%%rax \n"
+	  "movaps (%%rax),%%xmm2 \n"
+	  "movq %1,%%rax \n"
+	  "movaps (%%rax),%%xmm3 \n"
+	  :
+	  :"g"(c1),"g"(c2)
+	  :"rax","xmm2","xmm3");
+  Stack stack;
+  init(*stack.data[0]);
+  stack.size=1;
+  byte i,g,m,x,s,t;
+  ulong ng[GMAX+1];
+  for(i=0;i<=GMAX;++i) ng[i]=0;
+  ng[1]=1;
+  byte *data,*ndata;
+  SemiGroup *sg,*nsg;
+  while(stack.size>0){
+    sg=stack.pop(); 
+    g=sg->g;
+    if(g<GMAX){
+      ++g;
+      t=0;
+      data=sg->data;
+      s=(x=sg->c)+(m=sg->m);
+      for(;x<s;++x){
+	if(data[x]==1){
+	  ++t;
+	  nsg=stack.push();
+	  ndata=nsg->data;
+	  memcpy(ndata,data,SIZE);
+	  ndata[x]=0;
+	  nsg->c=x+1;
+ __asm__(//"mms \n"
+	  "movq %0,%%rax \n"            //&tab[1] --> rax
+	  "movq %1,%%rdx \n"            //&ntab[x+1] --> rdx
+	  "movb %2,%%cl \n"             //size-x-1 --> cl
+	  "cmp $16,%%cl \n"             //compare cl with 16
+	  "jna REM_LOOP \n"             //if cl<16 goto REM_LOOP
+	  "LOOP_SSE: \n"                //start of the loop using SSE operation
+	  "movups %%xmm3,%%xmm0 \n"     //0x00|...|00 --> xmm0"
+	  "movupd (%%rax),%%xmm4 \n"
+	  "pcmpeqb %%xmm4,%%xmm0 \n"    //test per byte if *rax==0 
+	                                //if true the corresping byte in xmm0 is 0xFF else is 0x00 
+	  "pandn %%xmm2,%%xmm0 \n"      //0x01|...|01 and (not mm0)--> mm0
+	  "movupd (%%rdx),%%xmm1 \n"
[…]

### 2014-09-03 src/others/ngtree-v4/semigroup.hpp

@@ -1,0 +1,49 @@
+#ifndef SEMIGROUP_HPP
+#define SEMIGROUP_HPP
+#include "config.hpp"
+#include "types.hpp"
+
+struct SemiGroup{
+  byte c,m,g;
+  byte data[SIZE];
+};
+
+void init(SemiGroup& sg);
+
+void disp(SemiGroup& sg);
+
+
+inline void init(SemiGroup& sg){
+  sg.c=sg.m=2;
+  sg.g=1;
+  for(byte i=0;i<SIZE;++i) sg.data[i]=i/2;
+}
+
+inline SemiGroup* copy(SemiGroup* sg){
+  SemiGroup* res=new SemiGroup;
+  res->c=sg->c;
+  res->m=sg->m;
+  res->g=sg->g;
+  for(byte i=0;i<SIZE;++i) res->data[i]=sg->data[i];
+  return res;
+}
+
+inline void disp(SemiGroup& sg){
+  bool begin=true;
+  cout<<'<';
+  for(byte i=0;i<SIZE;++i){
+    if(sg.data[i]==1){
+      if(not begin) cout<<',';
+      else begin=false;
+      cout<<i;
+    }
+  }
+  cout<<"> g="<<sg.g<<", m="<<sg.m<<", c="<<sg.c<<endl;
+}
+
+inline void full_disp(SemiGroup& sg){
+  for(byte i=0;i<SIZE;++i) cout<<sg.data[i]<<' ';
+  cout<<": g="<<sg.g<<", m="<<sg.m<<", c="<<sg.c<<endl;
+}
+
+#endif

### 2014-09-03 src/others/ngtree-v4/stack.hpp

@@ -1,0 +1,33 @@
+#ifndef STACK_HPP
+#define STACK_HPP
+#include "semigroup.hpp"
+
+class Stack{
+public:
+  SemiGroup *cache;
+  SemiGroup* data[STACK_CAPACITY];
+  ulong size;
+  SemiGroup* pop();
+  SemiGroup* push();
+  Stack();
+};
+
+inline Stack::Stack(){
+  cache=new SemiGroup;
+  for(uint i=0;i<STACK_CAPACITY;++i){
+    data[i]=new SemiGroup;
+  }
+}
+
+inline SemiGroup* Stack::push(){
+  return data[size++];
+}
+
+
+inline SemiGroup* Stack::pop(){
+  SemiGroup* temp=data[--size];
+  data[size]=cache;
+  return cache=temp;
+}
+
+#endif

### 2014-09-03 src/others/ngtree-v4/types.hpp

@@ -1,0 +1,20 @@
+#ifndef TYPES_HPP
+#define TYPES_HPP
+
+#include <iostream>
+
+using namespace std;
+
+typedef short int sint;
+typedef unsigned char uchar;
+typedef unsigned char byte;
+typedef unsigned short int usint;
+typedef unsigned int uint;
+typedef unsigned long ulong;
+
+inline ostream& operator<<(ostream& os,byte b){
+  os<<(usint)b;
+  return os;
+}
+
+#endif

### 2014-09-03 src/others/ngtree-v5/config.hpp
Add an SSE-intrinsics version without Cilk

@@ -1,0 +1,2 @@
+#define MAX_GENUS 35
+

### 2014-09-03 src/others/ngtree-v5/main.cpp

@@ -1,0 +1,82 @@
+#include <iostream>
+#include <iomanip>
+#include <cpuid.h>
+#include "monoid.hpp"
+
+using namespace std;
+
+void walk_stack(monoid m, unsigned long int* res)
+{
+  unsigned long int nbr;
+  monoid data[SIZE_BOUND], *stack[SIZE_BOUND], *current;
+  monoid **stack_pointer = stack + 1;
+
+  for (ind_t i=1; i<SIZE_BOUND; i++) stack[i] = &(data[i-1]); // Nathann's trick to avoid copy
+  stack[0] = &m;
+  while (stack_pointer != stack)
+    {
+      --stack_pointer;
+      current = *stack_pointer;
+      if (current->genus < MAX_GENUS - 1)
+	{
+	  nbr = 0;
+	  auto it = generator_iter<CHILDREN>(*current);
+	  while (it.move_next())
+	    {
+	      // exchange top with top+1
+	      stack_pointer[0] = stack_pointer[1];
+	      remove_generator(**stack_pointer, *current, it.get_gen());
+	      stack_pointer++;
+	      nbr++;
+	    }
+	  *stack_pointer = current;
+	  res[current->genus] += nbr;
+	}
+      else
+	{
+	  auto it = generator_iter<CHILDREN>(*current);
+	  res[current->genus] += it.count();
+	}
+    }
+}
+
+
+
+
+int main(int argc, char **argv)
+{
+  monoid N;
+  unsigned int ax, bx, cx, dx;
+  if (!__get_cpuid(0x00000001, &ax, &bx, &cx, &dx))
+    {
+      cerr << "Unable to determine the processor type !" << endl;
+      return EXIT_FAILURE;
+    }
+  if (!(cx & bit_SSSE3))
+    {
+      cerr << "This programm require sse3 instructions set !" << endl;
+      return EXIT_FAILURE;
+    }
+  if (!(cx & bit_POPCNT))
+    {
+      cerr << "This programm require popcount instruction !" << endl;
+      return EXIT_FAILURE;
+    }
+
+  init_full_N(N);
+  unsigned long int ng[MAX_GENUS];
+  for(unsigned int i=0;i<MAX_GENUS;++i){
+    ng[i]=0;
+  }
+  walk_stack(N,ng);
+
+  
+  cout << endl << "============================" << endl << endl;
+  for (unsigned int i=0; i<MAX_GENUS; i++)
+    {
[…]

### 2014-09-03 src/others/ngtree-v5/monoid.hpp

@@ -1,0 +1,290 @@
+#ifndef MONOID_HPP
+#define MONOID_HPP
+
+#include <cstdint>
+#include "config.hpp"
+
+// We cant have those as C++ constant because of the #define used below in
+// remove_generator manual loop unrolling. I don't know if the unrolling is
+// doable using template metaprogamming. I didn't manage to figure out how.
+
+#ifndef MAX_GENUS
+#error "Please define the MAX_GENUS macro"
+#endif
+#define SIZE_BOUND (3*(MAX_GENUS-1))
+#define NBLOCKS ((SIZE_BOUND+15) >> 4)
+#define SIZE (NBLOCKS << 4)
+// const uint_fast8_t MAX_GENUS = 40;
+// const uint_fast8_t SIZE_BOUND = (3*(MAX_GENUS-1));
+// const uint_fast8_t NBLOCKS = ((SIZE_BOUND+15) >> 4);
+// const uint_fast8_t SIZE = (NBLOCKS << 4);
+
+#include <x86intrin.h>
+
+typedef uint8_t epi8 __attribute__ ((vector_size (16)));
+typedef uint8_t dec_numbers[SIZE] __attribute__ ((aligned (16)));
+typedef epi8 dec_blocks[NBLOCKS];
+typedef uint_fast64_t ind_t;  // The type used for array indexes
+
+struct monoid
+{
+  union {
+    dec_numbers decs;
+    dec_blocks blocks;
+  };
+  // Dont use char as they have to be promoted to 64 bits to do pointer arithmetic.
+  ind_t conductor, min, genus;
+};
+
+void init_full_N(monoid &);
+void print_monoid(const monoid &);
+void print_epi8(epi8);
+inline void copy_blocks(      dec_blocks &__restrict__ dst,
+			      const dec_blocks &__restrict__ src);// __attribute__((always_inline));
+inline void remove_generator(monoid &__restrict__ dst,
+		      const monoid &__restrict__ src,
+			     ind_t gen);// __attribute__((always_inline));
[…]

### 2014-09-04 Papier/article.tex
Add goal of the paper in the introduction and do some stuff in last section

@@
 \usepackage{pgfplots}
+\usepackage{xspace}
 \theoremstyle{plain}
@@
 \newcommand{\ie}{\emph{i.e.}}
+\newcommand{\Cilk}{\texttt{Cilk}\xspace}
+\newcommand{\CilkP}{\texttt{Cilk++}\xspace}
+\newcommand{\CPP}{\texttt{C++}\xspace}
 \newcommand{\kkp}{{k+1}}
@@
 \renewcommand{\geq}{\geqslant}
+\newcommand{\MMX}{\texttt{MMX}\xspace}
+\newcommand{\SIMD}{\texttt{SIMD}\xspace}
+\newcommand{\SSE}{\texttt{SSE}\xspace}
+\newcommand{\SSEV}{\texttt{SSE4.1}\xspace}
 \newcommand{\NN}{\mathbb{N}}
@@
 \DeclareMathOperator{\card}{card}
-
+\newcommand{\Jean}[1]{{\color{blue}#1}}
+\newcommand{\Florent}[1]{{\color{green}#1}}
+\newcommand{\Delete}[1]{}
+\newcommand{\REM}[1]{\marginpar{\small #1}}
 \newcommand{\TODO}[1]{\marginpar{\small To do: #1}}
@@
 
-M.~Bras-Amor\'os used a depth first search exploration of the tree of numerical semigroups up to a given genus.
+M.~Bras-Amor\'os used a depth first search exploration of the tree of numerical semigroups $\mathcal{T}$ up to a given genus.
 This tree was introduced by J.C.~Rosales and al. in \cite{Rosales} and it is the subject of the Section~\ref{S:Tree}.
@@
 
-\TODO{Present the goal of the paper.}
+\Jean{Here we describe a new algorithm for the exploration of the tree of numerical semigroup $\mathcal{T}$ and achieve the computation of $n_g$ for $g\leq 67$. 
+The cornerstone of our method is a new combinatorial representation of numerical semigroup that is well suited for exploration of the tree~$\mathcal{T}$ and allow code optimization essentially based on use of vectorial instructions and parallelization.}
 
@@
 In Section~\ref{S:Tree} we describe the tree of numerical semigroups and give bounds for some parameters attached to a numerical semigroup.
-In Section~\ref{S:DecNumber} we describe a new representation of numerical semigroups that is well suited to the construction of the tree. 
+\Jean{The decription of our new representation of numerical semigroup is done in the third section.}
+\Delete{In Section~\ref{S:Tree} we describe a new representation of numerical semigrouos that is well suited to the construction of the tree.}
 In Section~\ref{S:Algo} we describe an algorithm based on the representation given in Section~\ref{S:DecNumber} and give its complexity. 
@@
 tricks, which we will elaborate in the following section:
+
+\REM{C'est un peu moche c'est gros points, non? Des tirets peut etre.}
 \begin{itemize}
-\item Vectorization (MMX, SSE instructions sets) and careful memory alignment.
-\item Shared memory multi-core computing using \texttt{CILK++} for low level
+\item Vectorization (\MMX, \SSE instructions sets) and careful memory alignment. 
+\item Shared memory multi-core computing using \CilkP for low level
   enumerating tree branching;
@@
 
-Assume for example that we want to construct the tree $\mathcal{T}^{\leq 100}$
+Assume for example that we want to construct the tree~$\mathcal{T}^{\leq 100}$
 of all numerical semigroup of genus smaller than $100$.  In this case, the
@@
 actually works on $1$ byte. However current CPUs usually work on $8$ bytes and
-even on $16$~bytes using vector extensions. The first optimization uses this
+even on $16$~bytes using vector extensions. \REM{Peut etre rajouter qqchose sur l'AVX; en fait on devrait pouvoir utiliser l'AVX2.} The first optimization uses this
 point.
@@
 and \tt{dst}.  We can therefore work in multiple entries at the same time
-without collision.  Current CPUs allow this thanks to the SIMD technologies as
-MMX, SSE, etc.  The acronym SIMD~\cite{SIMD} stands for Single Operation
-Multiple Data.  We used SSE4.1~\cite{SSE, IntelSSE} technology as it allows
+without collision.  Current CPUs allow this thanks to the \SIMD technologies as
+\MMX, \SSE, etc.  The acronym \SIMD~\cite{SIMD} stands for Single Operation
+Multiple Data.  We used \SSEV~\cite{SSE, IntelSSE} technology as it allows
 for the largest speedup. This need to respect some constraints in the memory
 organization of the data, namely ``memory alignment''. Recall that an address
-is 16~bytes aligned if it is a multiple of 16. SSE memory access are much
+is 16~bytes aligned if it is a multiple of 16. \SSE memory access are much
 faster for aligned access.
@@
 parent's decomposition numbers are copied in the children's using the
-following C++ code
+following \CPP code
 \begin{listing}{1}
@@
 Then the core of the \textbf{while} loop in the preceding algorithm is
-translated as a for loop as follows (recall that $x$ denotes the generator of
+translated as a \texttt{for} loop as follows (recall that $x$ denotes the generator of
 the father which is to be removed in the children):
@@
 \end{listing}
-The instruction \tt{load\_unaligned\_epi8} (specific to SSE technology) loads
-in the variable \tt{block} 16 consecutive entries of the decomposition number
-of the parent (called \tt{src} monoid). Those entries will be used to compute
+The instruction \tt{load\_unaligned\_epi8} (specific to \SSE technology) loads
+ 16 consecutive entries of the decomposition number
+of the parent (called \tt{src} monoid) in the variable \tt{block}.
+Those entries will be used to compute
 the entries $16i, \dots 16i+15$ of the children monoid. Since the removed
-generator $x$ is not necessarily a multiple of $16$, the data are not aligned
+generator~$x$ is not necessarily a multiple of $16$, the data are not aligned
 in memory, hence the use of a specific instruction. The \tt{zero}
 (resp. \tt{one}) constants are initialized as 16 bytes equal to $0$
-(resp. $1$). The comparison \tt{(block != zero)} therefore returns a block
+(resp. $1$). The comparison \tt{(block != zero)} therfore returns a block
 which contains $0$ in the bytes corresponding the $0$ entries of block and
@@
 
-\subsection{Parallel tree exploration using \texttt{Cilk+}}
+\subsection{Parallel tree exploration using \CilkP}
 
@@
 with a former one. Fortunately there is a technology called
-\texttt{Cilk+}~\cite{CilkIntel, CilkRefman}
+\CilkP~\cite{CilkIntel, CilkRefman}
 which is particularly well suited for those kind of problems.
 
-\tt{Cilk} is a general-purpose language designed for multithreaded parallel
-computing. The \tt{C++} incarnation is called \tt{Cilk+}. The biggest
-principle behind the design of the Cilk language is that the programmer should
+\Cilk is a general-purpose language designed for multithreaded parallel
+computing. The \CPP incarnation is called \CilkP. The biggest
+principle behind the design of the \Cilk language is that the programmer should
 be responsible for \emph{exposing} the parallelism, identifying elements that
@@
 The crucial thing is that two keywords are all that are needed to start
-using the parallel features of \tt{Cilk+}:
+using the parallel features of \CilkP:
 \begin{itemize}
@@
 \end{algorithmic}
-We just tells \tt{Cilk+} that the subtrees rooted at various children can be
+We just tells \CilkP that the subtrees rooted at various children can be
 explored in parallel.
@@
 \begin{algorithmic}[1]
-\State \tt{result}[$g(\tt{S}$] $\gets \tt{result}[g(\tt{S}] + 1$
+\State \tt{result}[$g(\tt{S})$] $\gets \tt{result}[g(\tt{S})] + 1$
 \end{algorithmic}
-then we face the problem of two core incrementing the same variable at the
-same time. Incrementing a variables is actually done in 3 steps: reading the
-value from the memory, adding one, storing back the results. Since there is by
-default non synchronization, the following sequence of actions for two core
-and is possible: Read1 / Read2 / Add1 / Add2 / Store1 / Store2. Then the two
+then we face the problem of two cores incrementing the same variable at the
+same time.
+ Incrementing a variable is actually done in 3 steps: reading the
+value from the memory, adding one, storing back the result. 
+Since there is by
+default non synchronization, the following sequence of actions for two cores
+ is possible: Read1 / Read2 / Add1 / Add2 / Store1 / Store2. Then the two
 cores performs the same modification resulting in incrementing the variable
 only once. This is called a data race and leads to nondeterministic wrong
-result. To cope with those kinds of synchronization problems, Cilk provide the
+result. To cope with those kinds of synchronization problems, \Cilk provide the
 notion of \emph{reducer} which are variables local to each thread which are
@@
 A more important problem is that the cost of a recursive call in non
-negligible. Using \tt{Cilk+} recursive calls instead of \tt{C++} calls makes
+negligible. Using \CilkP recursive calls instead of \CPP calls makes
 it even worse. The solution we use is to switch back to the non recursive
 version using a stack when the genus is close to the target genus. This leads
-to the following \tt{Cilk+} code:
+to the following \CilkP code:
 \begin{listing}{1}
@@
 around 10-12 for genus in the range $45\dots 67$ so that \tt{explore\_stack}
-is used more than $99\%$ of the time. The \tt{Cilk+} recursive function does
+is used more than $99\%$ of the time. The \CilkP recursive function does
 actually very little work but ensure that the work is balanced between the
@@
 the derecursived stack algorithm, we only allocate (on the system stack rather
-than on the heap), the stack of monoid. No further allocation are done.
+than on the heap) the stack of monoid. No further allocation are done.
 
@@
 the stack memory but we have to copy the top of the stack to \tt{S}. Since the
-average number of children is the golden ratio, this single copy in far from
+average number of children is the golden ratio, this single copy is far from
 being negligible. The trick is to use a level of indirection, replacing the
@@
 
-Due to the way, the C++ does its pointer arithmetic, even if the index in the
+Due to the way, the \CPP does its pointer arithmetic, even if the index in the
 array are less than $3G$ and therefore fits in $8$~bits, we use $64$~bits
@@
 The third is based on the second one with use of decomposition numbers, it corresponds to the algorithm given in Section~\ref{S:Algo}. 
-The last one is an optimisation of the third one with MMX.}
+The last one is an optimisation of the third one with \MMX.}
 \end{figure}

### 2014-09-05 Papier/article.tex
Biblio + new attempt of curves.

@@
 without collision.  Current CPUs allow this thanks to the \SIMD technologies as
-\MMX, \SSE, etc.  The acronym \SIMD~\cite{SIMD} stands for Single Operation
-Multiple Data.  We used \SSEV~\cite{SSE, IntelSSE} technology as it allows
+\MMX, \SSE, etc.  The acronym \SIMD~\cite{WikipediaSIMD} stands for Single Operation
+Multiple Data.  We used \SSEV~\cite{WikipediaSSE, IntelSSE} technology as it allows
 for the largest speedup. This need to respect some constraints in the memory
@@
 As we previously said, to gain more speed this core loop is actually unrolled
-using some kind of Duff device~\cite{Duff} as
+using some kind of Duff device~\cite{WikipediaDuff} as
 \begin{algorithmic}
@@
 version of our exploration algorithm uses a single core and so a fraction only
-of the power of a CPU. The idea here is that different branches of the tree can
-be explored in parallel by different cores of the computer. The tricky part is
-to ensure that each cores are busy, giving a new branch when a core is done
-with a former one. Fortunately there is a technology called
-\CilkP~\cite{CilkIntel, CilkRefman}
-which is particularly well suited for those kind of problems.
+of the power of a CPU. The idea here is that different branches of the tree
+can be explored in parallel by different cores of the computer. The tricky
+part is to ensure that each cores are busy, giving a new branch when a core is
+done with a former one. Fortunately there is a technology called
+\CilkP~\cite{CilkIntel, CilkRefman} which is particularly well suited for
+those kind of problems. For our computation, we used the free version which is
+integrated in the latest version of the GNU C compiler~\cite{GCCcilk}.
 
 \Cilk is a general-purpose language designed for multithreaded parallel
-computing. The \CPP incarnation is called \CilkP. The biggest
-principle behind the design of the \Cilk language is that the programmer should
-be responsible for \emph{exposing} the parallelism, identifying elements that
-can safely be executed in parallel; it should then be left to the run-time
-environment, particularly the scheduler, to decide during execution how to
-actually divide the work between processors.
+computing. The \CPP incarnation is called \CilkP. The biggest principle behind
+the design of the \Cilk language is that the programmer should be responsible
+for \emph{exposing} the parallelism, identifying elements that can safely be
+executed in parallel; it should then be left to the run-time environment,
+particularly the scheduler, to decide during execution how to actually divide
+the work between processors.
 
@@
 only once. This is called a data race and leads to nondeterministic wrong
-result. To cope with those kinds of synchronization problems, \Cilk provide the
+result. To cope with those kinds of synchronization problems, \CilkP provide the
 notion of \emph{reducer} which are variables local to each thread which are
@@
     \end{tikzpicture}
+    \label{F:Time:new}
+    \caption{Comparison of time executions of different exploration algorithms of the tree $\mathcal{T}^{\leq g}$ on an Intel \tt{i5-3570K} CPU with $8GB$ of memory. 
+The scale is logarithmic in time. 
+The first version is based on a breadth-first search exploration and the peak at genus~$35$ is due to the consumption of all the memory and use of swap. 
+The second version uses a depth first search exploration.
+The third is based on the second one with use of decomposition numbers, it corresponds to the algorithm given in Section~\ref{S:Algo}. 
+The last one is an optimisation of the third one with \MMX.}
+\end{figure}
+
+\begin{figure}[b]
+\begin{tikzpicture}[scale=1.3]
+        \begin{semilogyaxis}[xlabel=\small genus,ylabel=time,legend pos=north west,legend cell align=left]
+        \addplot [color=blue,mark=x] coordinates {
+(20,0.045767)
+(21,0.067697)
+(22,0.112697)
+(23,0.194646)
+(24,0.335412)
+(25,0.581481)
+(26,1.01088)
+(27,1.72613)
+(28,2.9324)
+(29,4.91856)
+(30,8.39383)
+(31,14.0308)
+(32,23.7138)
+(33,39.6461)
+(34,66.1154)
+(35,109.406)
+(36,182.331)
+(37,302.765)
+        };
+       \addplot [color=black,mark=x] coordinates {
+			(20,0.02)	
+(20,0.043697)
+(21,0.065487)
+(22,0.113301)
+(23,0.186903)
+(24,0.31748)
+(25,0.53251)
+(26,0.899982)
+(27,1.5205)
+(28,2.55423)
+(29,4.28809)
+(30,7.30236)
+(31,12.1284)
+(32,20.2132)
+(33,33.707)
+(34,56.5415)
+(35,93.7614)
+(36,157.09)
+(37,259.997)
+(38,427.497)
+(39,709.564)
+(40,1192.81)
+        };
+         \addplot [color=red,mark=x] coordinates {
+(20,0.027708)
+(21,0.044416)
+(22,0.074789)
+(23,0.122006)
+(24,0.203155)
+(25,0.344143)
+(26,0.594073)
+(27,1.07134)
+(28,1.622)
+(29,2.82573)
+(30,4.53882)
+(31,7.60975)
+(32,12.5442)
+(33,21.0466)
+(34,34.7638)
+(35,57.7621)
+(36,95.6878)
+(37,158.95)
+(38,264.495)
+(39,436.146)
+(40,719.347)
+        };
+        \addplot [color=magenta,mark=x] coordinates {
+
+        };
+	%\legend style={anchor=north,at={(20,1000)}}  
+        \legend{\tiny breadth-expl,\tiny depth-expl,\tiny depth-expl-$\delta$}
+        \end{semilogyaxis}
+    \end{tikzpicture}
     \label{F:Time}

### 2014-09-05 Papier/article.tex
Courbes + timming scripts.

@@
 when computing addresses of indexed element. This single standard trick save
-20\% of speed.
+10\% of speed.
 
@@
        \addplot [color=black,mark=x] coordinates {
-			(20,0.02)	
 (20,0.043697)
@@
         \addplot [color=magenta,mark=x] coordinates {
-
+(20,0.001562)
+(21,0.001957)
+(22,0.002711)
+(23,0.003811)
+(24,0.005605)
+(25,0.00861)
+(26,0.013127)
+(27,0.020991)
+(28,0.036053)
+(29,0.05879)
+(30,0.095882)
+(31,0.158853)
+(32,0.259237)
+(33,0.441074)
+(34,0.750349)
+(35,1.2067)
+(36,1.9608)
+(37,3.19713)
+(38,5.33436)
+(39,9.2136)
+(40,14.5226)
         };
@@
     \label{F:Time}
-    \caption{Comparison of time executions of different exploration algorithms of the tree $\mathcal{T}^{\leq g}$ on an Intel \tt{i5-3570K} CPU with $8GB$ of memory. 
+    \caption{Comparison of time executions of different exploration algorithms of the tree $\mathcal{T}^{\leq g}$ on an Intel \tt{i7-3840QM} CPU with $32GB$ of memory. 
 The scale is logarithmic in time. 
@@
 The third is based on the second one with use of decomposition numbers, it corresponds to the algorithm given in Section~\ref{S:Algo}. 
-The last one is an optimisation of the third one with \MMX.}
+The last one is an optimisation of the third one with \SSE but not using \CilkP.}
 \end{figure}

### 2014-09-05 src/others/ngtree-v1/main.cpp

@@ -1,3 +1,4 @@
+#include <time.h>
 #include <iostream>
 #include <list>
 #include <stdlib.h>
@@ -94,4 +95,5 @@ int main(int argc,char** argv){
     lg.clear();
     lg.swap(next);
   }
+  cout<<1.0*long(clock())/CLOCKS_PER_SEC<<endl;
 }

### 2014-09-05 src/others/ngtree-v2/main.cpp

@@ -1,3 +1,4 @@
+#include <time.h>
 #include <iostream>
 #include <list>
 #include <stdlib.h>
@@ -99,4 +100,5 @@ int main(int argc,char** argv){
   /*for(i=0;i<=G;++i){
     cout<<i<<"->"<<ng[i]<<endl;
     }*/
+  cout<<1.0*long(clock())/CLOCKS_PER_SEC<<endl;
 }

### 2014-09-05 src/others/ngtree-v3/main.cpp

@@ -1,3 +1,4 @@
+#include <time.h>
 #include <iostream>
 #include <list>
 #include <stdlib.h>
@@ -79,4 +80,5 @@ int main(int argc,char** argv){
   /*for(i=0;i<=G;++i){
     cout<<i<<"->"<<ng[i]<<endl;
     }*/
+  cout<<1.0*long(clock())/CLOCKS_PER_SEC<<endl;
 }

### 2014-09-05 src/others/ngtree-v4/config.hpp

@@ -1,4 +1,4 @@
-#define GMAX 40
+#define GMAX 30
 #define MMAX (GMAX+1)
 #define SIZE 3*GMAX+3
 #define STACK_CAPACITY MMAX*(GMAX+1)

### 2014-09-05 src/others/ngtree-v5/config.hpp

@@ -1,2 +1,1 @@
-#define MAX_GENUS 35
-
+#define MAX_GENUS 40

### 2014-09-05 src/others/ngtree-v5/main.cpp

@@ -1,3 +1,4 @@
+#include <time.h>
 #include <iostream>
 #include <iomanip>
 #include <cpuid.h>
@@ -71,12 +72,13 @@ int main(int argc, char **argv)
   walk_stack(N,ng);
 
   
-  cout << endl << "============================" << endl << endl;
-  for (unsigned int i=0; i<MAX_GENUS; i++)
-    {
-      cout << ng[i] << " ";
-    }
-  cout << endl;
+  // cout << endl << "============================" << endl << endl;
+  // for (unsigned int i=0; i<MAX_GENUS; i++)
+  //   {
+  //     cout << ng[i] << " ";
+  //   }
+  // cout << endl;
+  cout<<1.0*long(clock())/CLOCKS_PER_SEC<<endl;
   return EXIT_SUCCESS;
 }
 

### 2014-09-08 Papier/article.tex
Remove Bibtex Reduce size of algorithm Replace bullet of itemize

@@
 \usepackage{pgfplots}
+\usepackage{alltt}
+\usepackage{listings}
 \usepackage{xspace}
@@
 
+%Replace bullet in item by dash
+\def\labelitemi{--}
 \begin{document}
@@
 \EndProcedure
-\State \Call{ExploreRec}{$\NN, \tt{G}$}
 \end{algorithmic}
@@
 \caption{Iterative Depth first search exploration of the tree of numerical semigroups}\label{A:Exp}
+\small
 \begin{algorithmic}[1]
@@
 \begin{algorithm}
-\caption{Return the root of $\mathcal{T}$ for an exploration up to genus \tt{G}}\label{A:Root}
+\caption{Returns the root of $\mathcal{T}$ for an exploration up to genus \tt{G}}\label{A:Root}
 \begin{algorithmic}
+\small
 \Function{Root}{\tt{G}}
@@
 \begin{algorithmic}[1]
+\small
 \Function{Son}{\tt{S},\tt{x},\tt{G}}
@@
 \begin{algorithmic}[1]
+\small
 \Function{Count}{\tt{G}}
@@
 
-\REM{C'est un peu moche c'est gros points, non? Des tirets peut etre.}
 \begin{itemize}
@@
 actually works on $1$ byte. However current CPUs usually work on $8$ bytes and
-even on $16$~bytes using vector extensions. \REM{Peut etre rajouter qqchose sur l'AVX; en fait on devrait pouvoir utiliser l'AVX2.} The first optimization uses this
-point.
+even on $16$~bytes using vector extensions. 
+The first optimization uses this point.
 
@@
 
+{\small
 \begin{algorithmic}
@@
 \EndWhile
-\end{algorithmic}
+\end{algorithmic}}
 
@@
 without collision.  Current CPUs allow this thanks to the \SIMD technologies as
-\MMX, \SSE, etc.  The acronym \SIMD~\cite{WikipediaSIMD} stands for Single Operation
-Multiple Data.  We used \SSEV~\cite{WikipediaSSE, IntelSSE} technology as it allows
-for the largest speedup. This need to respect some constraints in the memory
+\MMX, \SSE, etc.  The acronym \SIMD~\cite{WikipediaSIMD} stands for Single Operation Multiple Data.  
+We used \SSEV~\cite{WikipediaSSE, IntelSSE} technology as it allows for the largest speedup\footnote{A much more speedup can be certainly obtained using \texttt{AVX2} technology \cite{AVX2}.
+However, at this time, we can't access a performent computer with this set of instructions.}.
+This need to respect some constraints in the memory
 organization of the data, namely ``memory alignment''. Recall that an address
@@
 following \CPP code
-\begin{listing}{1}
+{
+\small
+\begin{alltt}
 void copy_blocks(dec_blocks &dst, dec_blocks const &src) {
@@
 }
-\end{listing}
+\end{alltt}}
 Here \tt{dec\_blocks} is a type for arrays of 16~bytes blocks whose size
@@
 the father which is to be removed in the children):
-\begin{listing}{1}
+{
+\small
+\begin{alltt}
 start = x >> 4;      // index of the block containing x
@@
 }
-\end{listing}
+\end{alltt}
+}
 The instruction \tt{load\_unaligned\_epi8} (specific to \SSE technology) loads
@@
 the instruction actually performs a 16~bytes parallel version of
-\begin{algorithmic}
-\State $\tt{dst} \gets \tt{dst} -
+
+\begin{center}
+$\tt{dst} \gets \tt{dst} -
 \textbf{ if } \tt{block} \neq 0 \textbf{ then } 1 \textbf{ else } 0$
-\end{algorithmic}
+\end{center}
+
 which is equivalent to Lines \tt{10} to \tt{14} of Algorithm~\textsc{Son}.
@@
 using some kind of Duff device~\cite{WikipediaDuff} as
+
+{\small
 \begin{algorithmic}
@@
 \State \textbf{case NBLOCK-1:} \tt{...}
-\end{algorithmic}
+\end{algorithmic}}
 
@@
 Algorithm~\ref{A:ExpRec}, you only need to modify it as
-\begin{algorithmic}[1]
+
+{\small
+
+\begin{algorithm}
+\caption{\Cilk version of Algorithm~\ref{A:ExpRec}}
+\begin{algorithmic}
 \Procedure{ExploreRec}{\tt{S}, \tt{G}}
@@
 \end{algorithmic}
+\end{algorithm}
+}
+
 We just tells \CilkP that the subtrees rooted at various children can be
@@
 gather the results of the exploration. If we simply write
-\begin{algorithmic}[1]
-\State \tt{result}[$g(\tt{S})$] $\gets \tt{result}[g(\tt{S})] + 1$
-\end{algorithmic}
+\begin{center}
+\tt{result}[$g(\tt{S})$] $\gets \tt{result}[g(\tt{S})] + 1$
+\end{center}
 then we face the problem of two cores incrementing the same variable at the
@@
 to the following \CilkP code:
-\begin{listing}{1}
+{\small
+\begin{verbatim}
 void explore(const monoid &m) {
@@
 }
-\end{listing}
+\end{verbatim}
+}
 Note that in our version, we found that \tt{STACK\_BOUND} optimal value was
@@
 following sequences of operations:
-\begin{algorithmic}[1]
+{\small
+\begin{algorithmic}
   \State $\tt{S} \gets \tt{stack.top()}$
@@
 \end{algorithmic}
+}
 If we use a stack of monoid, we can construct $\tt{S}^\tt{x}$ directly into
@@
 indexes. Here is the relevant part of the code:
-\begin{listing}{1}
+
+{\small
+\begin{verbatim}
   monoid data[MAX_GENUS-1], *stack[MAX_GENUS], *current;
@@
   }
-\end{listing}
+\end{verbatim}
+}
 
@@
 33 & 24896206 & 1.63843 \\
-34 & 40761087 & 1.63724 \\
-35 & 66687201 & 1.63605 \\
+
 \hline
@@
 \hline
+34 & 40761087 & 1.63724 \\
+35 & 66687201 & 1.63605 \\
 36 & 109032500 & 1.63498 \\
@@
 \footnotesize
-\bibliographystyle{ieeetr}
-\bibliography{biblio}
+
+\begin{thebibliography}{10}
+
+\bibitem{BrasAmoros2008}{M.~Bras-Amor{\'o}s, \emph{Fibonacci-like behavior of the number of numerical semigroups of a given genus}, Semigroup Forum \textbf{76}, no.~2 (2008), 379--384}
+
+ \bibitem{Delgado}{M.~Delgado, homepage, \url{http://cmup.fc.up.pt/cmup/mdelgado/numbers/}}
+
+\bibitem{NumericalSgps}{M.~Delgado, P.A.~Garcia-Sanchez, J.~Maris, \emph{NumericalSgps} -- GAP package, ver.0.971, 2001 \url{http://cmup.fc.up.pt/cmup/mdelgado/numericalsgps/}}
+
+\bibitem{GAP}{GAP -- Groups, Algorithms, and Programming, Version 4.6.3, \url{http://www.gap-system.org}}
+
+\bibitem{BookNS}{P.A.~Garsia-Sanchez, J.C.~Rosales, \emph{Numerical semigroups}, Springer, New York, 2009, Developments in Mathematics, vol.~20, x+181}
+
+\bibitem{GCCcilk}{B.V.~Iyer, R.~Geva, P.~Halpern, \emph{Cilk\texttrademark{}+ in GCC}
+,GNU Tools Cauldron, 2012, \url{http://gcc.gnu.org/wiki/cauldron2012?action=AttachFile&do=get&target=Cilkplus_GCC.pdf}}
+
+\bibitem{BookDFP}{J.L.~Ram{\'{\i}}rez Alfons{\'{\i}}n, \emph{The Diophantine Frobenius Problem}, Oxford University press, 2005, Oxford Lecture Series in Mathematics and its Applications, vol.~30, xvi+243}
+
+\bibitem{Rosales}{J.C.~Rosales, \emph{Fundamental gaps of numerical semigroups generated by two elements}, Linear Algebre Appl. \textbf{405} (2005), 200--208}
+
+\bibitem{OEIS}{N.J.A.~Sloane, \emph{The On-Line Encyclopedia of Integer Sequences}, \url{http://oeis.org}}
+
+\bibitem{IntelSSE}{Software.intel.com, \emph{Intel Instruction Set Architechture Extensions}, Intel\textregistered{} Developer Zone, 2013, \url{https://www.cilkplus.org/}}
+
+\bibitem{CilkIntel}{Software.intel.com, \emph{Intel\textregistered{} Cilk\texttrademark{} Homepage}, \url{ttps://www.cilkplus.org/}}
+
+\bibitem{CilkRefman}{Software.intel.com, \emph{Intel\textregistered{} Cilk\texttrademark{} Plus Reference Guide},\url{https://software.intel.com/en-us/node/522579}}
+
+\bibitem{AVX2}{Wikipedia, \emph{Advanced Vector Extension}, \url{http://en.wikipedia.org/wiki/Advanced_Vector_Extensions}}
+
+\bibitem{WikipediaDuff}{Wikipedia, \emph{Duff's device}, \url{http://en.wikipedia.org/wiki/Duff's_device}}
+
+\bibitem{WikipediaSIMD}{Wikipedia, \emph{SIMD}, \url{http://en.wikipedia.org/wiki/SIMD}}
+
+\bibitem{WikipediaSSE}{Wikipedia, \emph{Streaming SIMD Extensions}, \url{http://en.wikipedia.org/wiki/Streaming_SIMD_Extensions}}
+
+\bibitem{Zhai}{A. Zhai, \emph{Fibonacci-like growth of numerical semigroups of a given genus}, Semigroup Forum\textbf{86}, no.3 (2013), 634--662}
+\end{thebibliography}
 

### 2014-09-16 Papier/article.tex
Intro + abstract.

@@
 \begin{abstract}
-In this paper we describe an algorithm visiting all the numerical semigroups up to a given genus using a new representation of numerical semigroups. 
+  In this paper we describe an algorithm visiting all the numerical semigroups
+  up to a given genus using a new representation of numerical semigroups. The
+  interest of this algorithm is that it fits particularly well the
+  architecture of modern computer allowing very large optimizations.
 \end{abstract}
@@
 
-\Jean{Here we describe a new algorithm for the exploration of the tree of numerical semigroup $\mathcal{T}$ and achieve the computation of $n_g$ for $g\leq 67$. 
-The cornerstone of our method is a new combinatorial representation of numerical semigroup that is well suited for exploration of the tree~$\mathcal{T}$ and allow code optimization essentially based on use of vectorial instructions and parallelization.}
+Here we describe a new algorithm for the exploration of the tree of numerical
+semigroup $\mathcal{T}$ and achieve the computation of $n_g$ for $g\leq 67$.
+The cornerstone of our method is a new combinatorial representation of
+numerical semigroup that is well suited for exploration of the
+tree~$\mathcal{T}$ and allow \Florent{large} code optimization essentially based on use of
+vectorial instructions and parallelization. \Florent{The goal of the paper is twofold:
+first to present our new encoding of the semigroup and the associated
+algorithms, and second to present the optimization technique which allows for
+those kinds of algorithms to get speedups by factors of hundreds and even
+thousand. We claim that these technique are fairly general for those kinds of
+algorithms. As a support for the claim, we applied it to an algorithm of
+N.~Borie enumerating integer vector modulo permutation groups~\cite{Borie} and
+got a speedup by a factor larger than 2000 using 8 cores !}
 
@@
 
+\bibitem{Borie}{N.~Borie, \emph{Generation modulo the action of a permutation
+      group}, Proceeding of The 25th International Conference on Formal Power
+    Series and Algebraic Combinatorics (FPSAC '13), DMTCS Proceedings \textbf{0}(01) (2013),
+    767--778}
+
  \bibitem{Delgado}{M.~Delgado, homepage, \url{http://cmup.fc.up.pt/cmup/mdelgado/numbers/}}

### 2014-09-22 Papier/article.tex
Check of the paper + Wilf + Timings + Cilk/Timings + Abstract modif

@@
   interest of this algorithm is that it fits particularly well the
-  architecture of modern computer allowing very large optimizations.
+  architecture of modern computer allowing very large optimizations : we obtain the number of numerical semigroups of genus $g\leq  67$ and we confirm the Wilf conjecture for $g\leq 60$.
+
 \end{abstract}
@@
 
-\section{Introduction}
+\section*{Introduction}
 
@@
 \label{E:NSG}
-S_E=\{0,3,6,7,9,10\}\cup[12,+\infty[
+S_E=\{0,3,6,7,9,10\}\cup\{x\in\NN, x\geq 12\}
 \end{equation}
@@
 numerical semigroup that is well suited for exploration of the
-tree~$\mathcal{T}$ and allow \Florent{large} code optimization essentially based on use of
-vectorial instructions and parallelization. \Florent{The goal of the paper is twofold:
+tree~$\mathcal{T}$ and allow large code optimization essentially based on use of
+vectorial instructions and parallelization.
+The goal of the paper is twofold:
 first to present our new encoding of the semigroup and the associated
@@
 N.~Borie enumerating integer vector modulo permutation groups~\cite{Borie} and
-got a speedup by a factor larger than 2000 using 8 cores !}
+got a speedup by a factor larger than 2000 using 8 cores !
 
@@
 In Section~\ref{S:Tree} we describe the tree of numerical semigroups and give bounds for some parameters attached to a numerical semigroup.
-\Jean{The decription of our new representation of numerical semigroup is done in the third section.}
-\Delete{In Section~\ref{S:Tree} we describe a new representation of numerical semigrouos that is well suited to the construction of the tree.}
+The decription of our new representation of numerical semigroup is done in the third section.
 In Section~\ref{S:Algo} we describe an algorithm based on the representation given in Section~\ref{S:DecNumber} and give its complexity. 
 Section~\ref{S:Opti} is more technical, and is devoted to the optimisation of the algorithm introduced in Section~\ref{S:Algo}.
+Il the last section we emphase the results obtained using our algorithm.
 
@@
 
-By definition a numerical semigroup is an infinite object. 
-We need a finite description of such a semigroup. 
+By definition a numerical semigroup is infinite object and we need a finite description of such a object. 
 That is the role of generating sets.
@@
 \begin{lem}[Lemma 2.3, Chap. I, \cite{BookNS}]
-Let $S$ be a numerical semigroup. The set $\Irr(S)$ is the minimal generating set of $S$ relatively to the inclusion ordering.
+For a numerical semigroup $S$, the set $\Irr(S)$ is the minimal generating set of $S$ relatively to the inclusion ordering.
 \end{lem}
 
-%\begin{proof}
-%Assume for a contradiction, that there exists an integer  $x$ in $S$ than cannot be decomposed as a sum of irreducible elements. 
-%We may assume that $x$ is  minimal with this property. 
-%As  $x$ cannot be irreducible, there exist $y$ and $z$ in~$S\setminus\{0\}$ satisfying $x=y+z$. 
-%Since we have $y<x$ and $z<x$, the integers $y$ and $z$ can be expressed as a sum of irreducible elements of $S$ and so $x=y+z$ is a sum of irreducible elements, in contradiction to hypothesis. 
-%
-%As irreducible elements of $S$ cannot be decomposed as the sum of two non-zeros integers in $S$, they must occur in each generating set of $S$.
-%\end{proof}
-
-%Let $S$ be a numericla semigroup. As $m(S)$ lies in $S$ the irreducible elements are include in
-%\[
-%\App(S)=\{x\in S \, |\, x-m(S)\not\in S\}
-%\]
-%The set $\App(S)$ is the so-called Ap\'ery set relativey to the element $m(S)$.
-%
-%\begin{lem}[Lemma 2.4, Chap. I, \cite{BookNS}]
-%Let $S$ be a numerical semigroup. The Apéry set $\App(S)$ is finite of cardinality $m(S)$.
-%\end{lem}
-%
-%The following result is a direct consequence of the previous Lemma and the Inclusion of $\Irr(S)$ in $m(S)$.
-%
-%\begin{prop}
-%\label{P:Res}
-%A numerical semigroup has at most $m(S)$ irreducible elements.
-%\end{prop}
-
-The differents parameters we have defined on a numerical semigroup, statisfy the following relations.
+The different parameters we have defined on a numerical semigroup, satisfy the following relations.
 
@@
 \label{P:Res}
-Let $S$ be a numerical semigroup. We have
+For every  numerical semigroup  $S$,  we have
 
@@
 
-$iii)$ Let $x$ be an element of $S\cap[0,c(S)-1]$. 
-The integer $y=c(S)-1-x$ lies in $[0,c(S)-1]$. 
+$iii)$ Let $x$ be an element of $S\cap\{0,...,c(S)-1\}$. 
+The integer $y=c(S)-1-x$ lies in $\{0,...,c(S)-1\}$. 
 Moreover $y$ is not in $S$, for otherwise  we write $c(S)-1=y+x$ where $x,y$ are elements of $S$ implying $c(S)-1\in S$. 
@@
 \begin{array}{rcl}
-\psi:[0,c(S)-1]&\to&[0,c(S)-1]\\
+\psi:\{0,...,c(S)-1\}&\to&\{0,...,c(S)-1\}\\
 x&\mapsto&c(S)-1-x
@@
 \]
-that send $S\cap[0,c(S)-1]$ into $[0,c(S)-1]\setminus S$ which is  of cardinality $g(S)$. 
-Let us denote by $k$ the cardinality  of $S\cap[0,c(S)-1]$. 
+that send $S\{0,...,c(S)-1\}$ into $\{0,...,c(S)-1\}\setminus S$ which is  of cardinality $g(S)$. 
+Let us denote by $k$ the cardinality  of $S\cap\{0,...,c(S)-1\}$. 
 Since $\psi$ is injective we must have $k\leq g(S)$. 
@@
 \[
-[0,c(S)[=S\cap[0,c(S)[\sqcup[0,c(S)[\setminus S
+\{0,...,c(S)-1\}=S\cap\{0,...,c(S)-1\}\sqcup\{0,...,c(S)-1\}\setminus S
 \]
@@
 We now explain the construction of the tree of numerical semigroups.
-Let~$S$ be a numerical semigroup. The set $T=S\cup\{c(S)-1\}$ is also a numerical  semigroup and its genus is $g(S)-1$. 
-As  $[c(S)-1,+\infty[$ is included in $T$ we have $c(T)\leq c(S)-1$. 
-Therefore every semigroup $S$ of genus $g$ can be obtained from a semigroup $T$ of genus $g-1$ by removing an element of~$T$ greater than or equal to $c(T)$.
+Let~$S$ be a numerical semigroup. The set $S'=S\cup\{c(S)-1\}$ is also a numerical  semigroup and its genus is $g(S)-1$. 
+As  each integer greater than $c(S)-1$ is included in $S'$ we have $c(S')\leq c(S)-1$. 
+Therefore every semigroup $S$ of genus $g$ can be obtained from a semigroup $S'$ of genus $g-1$ by removing an element of~$S'$ greater than or equal to $c(S')$.
 
@@
 
-Proposition~\ref{P:Sx} implies that every semigroup $T$ of genus $g$ can be obtained from a semigroup $S$ by removing a generator $x$ of $S$ that is greater than $c(S)$.
-Hence relations~$T=S^x=S\setminus\{x\}$ hold.
+Proposition~\ref{P:Sx} implies that every semigroup $S$ of genus $g$ can be obtained from a semigroup $S$ by removing a generator $x$ of $S$ that is greater than $c(S)$.
+Hence the relation~$S'=S^x$ holds.
 
@@
 With this construction, a semigroup $S$ has depth $g$ in $\mathcal{T}$ if and only if its genus is~$g$, see Figure~\ref{F:Tree}.
-We denote by $\mathcal{T}_{\leq g}$ the subtree of $\mathcal{T}$ resticted to all semigroup of genus $\leq g$.
+We denote by $\mathcal{T}_{g}$ the subtree of $\mathcal{T}$ resticted to all semigroup of genus $\leq g$.
+
+\vspace{-3em}
 
  \begin{figure}[hf!]
- % \[
- % \pstree[,nodesep=3pt]{\sgnode{1}}{
- % 	\pstree[nodesep=3pt]{\sgnode{2,3} \tlput1}{
- % 		\pstree[treesep=4pt,nodesep=2pt]{\sgnode{3,4,5} \tlput2}{
- % 			\pstree[treesep=3pt,nodesep=3pt]{\sgnode{4,5,6,7} \tlput3}{
- % 				\pstree{\sgnode{5,6,7,8,9} \tlput4}{}
- % 				\pstree{\sgnode{\gr 4,6,7,9} \tlput5}{}
- % 				\pstree{\sgnode{\gr 4,\gr 5,7} \trput6}{}
- % 				\pstree{\sgnode{\gr 4,\gr 5,\gr 6} \trput7}{}
- % 			}
- % 			\pstree[treesep=3pt,nodesep=3pt]{\sgnode{\gr 3,5,7} \tlput4}{
- % 				\pstree{\sgnode{\gr 3,7,8} \tlput 5}{}
- % 				\pstree{\sgnode{\gr 3,\gr 5} \trput 7}{}
- % 			}
- % 			\pstree[treesep=10pt]{\sgnode{\gr 3,\gr 4} \trput5}{\Tn \Tn \Tn}
- % 		}
- % 		\pstree[nodesep=3pt]{~\hspace{-5em}\sgnode{\gr2,5} \trput3}{
- % 			\pstree[nodesep=3pt]{\sgnode{\gr2,7} \trput5}{
- % 				\pstree[nodesep=3pt]{\sgnode{\gr2,9} \trput7}{}
- % 			}
- % 		}
- % 	}
- % }
- % \vspace{-5em}
- % \]
-
-\[
-\begin{tikzpicture}[level distance=2cm, inner sep=2mm,
-  level/.style={sibling distance=1.55cm},
-%  edge from parent/.style={draw}
-  ]
+\[
+\begin{tikzpicture}[level distance=2cm, inner sep=2mm,level/.style={sibling distance=1.55cm}]
     \node {$\sgnode{1}$}
@@
   \]
-\caption{The first four layers of the tree $\mathcal{T}$ of numerical semigroups, corresponding to $\mathcal{T}_{\leq 4}$.
-A generator of a semigroup is in gray if it is not in $[c(S),+\infty[$.
+\caption{The first four layers of the tree $\mathcal{T}$ of numerical semigroups, corresponding to $\mathcal{T}_4$.
+A generator of a semigroup is in gray if is not greater that $c(S)$.
 An edge between a semigroup $S$ and its son $S'$ is labelled by  $x$ is $S'$ is obtained from $S$ by removing~$x$, that is $S'=S^x$.}
@@
 As the reader can check, the main difficulty to characterize the son of a semigroup is to determine its irreducible elements.
- In \cite{NumericalSgps}, the semigroup are given by their Ap\'ery set and then the difficulty is to describe the Ap\'ery set of $S_x$  from this of $S$. 
+ In \cite{NumericalSgps}, the semigroup are given by their Ap\'ery set and then the difficulty is to describe the Ap\'ery set of $S^x$  from this of $S$. 
  This approach is elegant but not sufficiently \emph{basic} for our optimisations.
@@
 
-The aim of this section is to describe a new representation of numerical semigroups.
+The aim of this section is to describe a new representation of numerical semigroups,
 which is well suited to an efficient exploration of the tree~$\mathcal{T}$ of numerical semigroups.
@@
  Hence~$D_S(x)$ describes how~$x$ can be decomposed as sums of two elements of $S$.
- This is we decide to call $d_S(x)$ the $S$-decomposition number of $x$.
+ This is why we decide to call $d_S(x)$ the $S$-decomposition number of $x$.
 
@@
 \label{P:Dandd}
-Let $S$ be a numerical semigroup and $x\in\NN\setminus\{0\}$.  We have:
+For a numerical semigroup $S$ and $x\in\NN\setminus\{0\}$, we have:
 
@@
 \]
-We construct three sets as follow:
+We construct two sets as follow:
 \begin{align*}
@@
 \end{align*}
-We have the relation $D_{S}(y)=D_{S^x}(y)\sqcup E\sqcup F$ and so $d_S(y)$ is equal to $d_{S^x}(y)+\card(E)+\card(F)$.
+We have the relation $D_{S}(y)=D_{S^x}(y)\sqcup E\sqcup F$ and so $d_S(y)$ is equal to $d_{S^x}(y)+\card(E)+\card(F)$
 and $D_{S^x}(y)$ is a subset of $D_S(y)$. We start by determining $E$. 
@@
 If $y-x$ not belong to $S$, then $E$ and $F$ are empty implying $d_{S}(y)=d_{S^x}(y)$.
-As the condition $y-x\in S$ is equivalent to $y\geq x$ together with $d_S(y-x)>0$ by Proposition~\ref{P:Dandd} the result holds.
+As the condition $y-x\in S$ is equivalent to $y\geq x$ together with $d_S(y-x)>0$, by Proposition~\ref{P:Dandd} the result holds.
 \end{proof}
@@
 \label{S:Algo}
-We can easily explore the tree of numerical semigroups up to a genus $G$ using
+We can easily explore the tree of numerical semigroups up to a genus~$G$ using
 a depth first search algorithm (see Algorithm~\ref{A:ExpRec}). This approach does not seem to
-have been used before.  In particular, M. Bras-Amor\'os and M. Delgado use
+have been used before.  In particular, M. Bras-Amor\'os and M.~Delgado use
 instead a breadth first search exploration.  The main advantage in our
@@
 explore the tree deeper, we must restart from the root.
+
 \begin{algorithm}
-\caption{Recursive Depth first search exploration of the tree of numerical semigroups}\label{A:ExpRec}
+\small
+\caption{\small Recursive Depth first search exploration of the tree $\mathcal{T}_g$.}\label{A:ExpRec}
 \begin{algorithmic}[1]
@@
 
-Equivalently, we can use an iterative version which uses a stack (see Algorithm~\ref{A:Exp}).
+Equivalently, we can use an iterative version which uses a stack:
 \begin{algorithm}
-\caption{Iterative Depth first search exploration of the tree of numerical semigroups}\label{A:Exp}
+\caption{\small Iterative Depth first search exploration of the tree $\mathcal{T}_g$}\label{A:Exp}
 \small
@@
 \begin{proof}
-By Proposition~\ref{P:Res} $iii)$ we have $c(S)\leq 2g(S)$ and so the $S$-decomposition number of $c(S)$ occurs in $\delta_S$.
+By Proposition~\ref{P:Res} $iii)$ we have  the relation $c(S)\leq 2g(S)$ and so the $S$-decomposition number of $c(S)$ occurs in $\delta_S$.
 Since $c(S)$ is equal to $\max(\NN\setminus S)$, Proposition~\ref{P:delta} implies 
 \[
-c(S)=1+\max\{i\in[0,...,3G],\ d_S(i)=0\}.
+c(S)=1+\max\{i\in\{0,...,3G\},\ d_S(i)=0\}.
 \]
@@
 \[
-g(S)=\card\{i\in[0,..,3G],\ d_S(i)=0\}.
+g(S)=\card\{i\in\{0,...,3G\},\ d_S(i)=0\}.
 \]
@@
 \[
-m(S)=\min\{i\in[0,...,3G],\ d_S(i)>0\}.
-\]
-Since, by Proposition~\ref{P:Res} $i)$,  all irreducible elements are smaller than $c(S)+m(S)-1$, which is itself smaller than $3G$ the relations $ii)$ and $iii)$ of Proposition~\ref{P:Res} give
-\[
-\Irr(S)=\{i\in[0,...,3G],\ d_S(i)=1\}.
+m(S)=\min\{i\in\{0,...,3G\},\ d_S(i)>0\}.
+\]
+Since, by Proposition~\ref{P:Res} $i)$,  all irreducible elements are smaller than $c(S)+m(S)-1$, which is itself smaller than $3G$ from the relations $ii)$ and $iii)$ of Proposition~\ref{P:Res}.
+So Proposition~\ref{P:Dandd} gives
+\[
+\Irr(S)=\{i \in\{0,...,3G\},\ d_S(i)=1\}.
 \]
@@
 
-For example the following Algorithm initializes a representation of the semigroup $\NN$ ready for an exploration up to genus $G$.
+For example the following Algorithm initializes a representation of the semigroup $\NN$ ready for an exploration of the tree $\mathcal{T}_G$, \ie, the tree of numerical semigroup upto genus $G$.
 
 \begin{algorithm}
-\caption{Returns the root of $\mathcal{T}$ for an exploration up to genus \tt{G}}\label{A:Root}
+\caption{\small Returns the root of $\mathcal{T}_G$}\label{A:Root}
 \begin{algorithmic}
@@
 
+\newpage
+
 \begin{algorithm}
-\caption{Returns the son $S_x$ of $S$ with $x\in\Irr(S)\cap[c(S),c(S)+m(S)[$.}\label{A:Son}
+\caption{\small Returns the son $S^x$ of $S$ with $x\in\Irr(S)\cap[c(S),c(S)+m(S)[$.}\label{A:Son}
 \begin{algorithmic}[1]
@@
 \Function{Son}{\tt{S},\tt{x},\tt{G}}
-\State \tt{T.c} $\gets$ $\tt{x}+1$ \Comment{\tt{T}  stands for $\tt{S}_x$}
-\State \tt{T.g} $\gets$ $\tt{S.g}+1$
+\State \tt{$\tt{S}^\tt{x}$.c} $\gets$ $\tt{x}+1$
+\State \tt{$\tt{S}^\tt{x}$.g} $\gets$ $\tt{S.g}+1$
 \If{$\tt{x}>\tt{S.m}$}
-	\State \tt{T.M} $\gets$ \tt{S.m}
+	\State \tt{$\tt{S}^\tt{x}$.m} $\gets$ \tt{S.m}
 \Else
-	\State \tt{T.M} $\gets$ $\tt{S.m}+1$
+	\State \tt{$\tt{S}^\tt{x}$.m} $\gets$ $\tt{S.m}+1$
 \EndIf
-\State\tt{T.d} $\gets$ \tt{S.d}
+\State\tt{$\tt{S}^\tt{x}$.d} $\gets$ \tt{S.d}
 \For{\tt{y} from \tt{x} to $3\,\tt{G}$}
 	\If{$\tt{S.d}[\tt{y}-\tt{x}]>0$}
-		\State $\tt{T.d}[\tt{y}]\gets\tt{T.d}[\tt{y}]-1$
+		\State $\tt{$\tt{S}^\tt{x}$.d}[\tt{y}]\gets\tt{$\tt{S}^\tt{x}$.d}[\tt{y}]-1$
 	\EndIf
 \EndFor
-\State{\textbf{return} \tt{T}}
+\State{\textbf{return} $\tt{S}^\tt{x}$}
 \EndFunction
@@
 \label{P:Son}
-Running on $(S,x,G)$ with $g(S)\leq G$, $x\in \Irr(S)$ and~$x\leq c(S)$, Algorithm~\ref{A:Son} returns the numerical semigroup $S_x$ in time $O(\log(G)\times G)$.
+Running on $(S,x,G)$ with $g(S)\leq G$, $x\in \Irr(S)$ and~$x\leq c(S)$, Algorithm~\ref{A:Son} returns the semigroup $S^x$ in time $O(\log(G)\times G)$.
 \end{prop}
@@
 \begin{proof}
+Let us check the correctness of the algorithm.
 By construction $S^x$ is the semigroup $S\setminus\{x\}$.
@@
 \begin{algorithm}
-\caption{Returns an array containing the value of $n_g$ for $g\leq G$}\label{A:Count}
+\caption{\small Returns an array containing the value of $n_g$ for $g\leq G$}\label{A:Count}
 \begin{algorithmic}[1]
@@
 \begin{prop}
-Running on $G\in\NN$, Algorithm \textsc{Count} returns the array~$[n_0,...,n_G]$ in time
+Running on $G\in\NN$, Algorithm~\ref{A:Count} returns the values of $n_g$ for $g\leq G$  in time
 \[
@@
 
-For the time complexity, let us remark that Algorithm~\textsc{Son} is called for every semigroup of the tree $\mathcal{T}^{\leq G}$ (the restriction of $\mathcal{T}$ to semigroup of genus $\leq G$). 
+For the time complexity, let us remark that Algorithm~\textsc{Son} is called for every semigroup of the tree $\mathcal{T}_G$ (the tree of semigroups of genus $\leq G$). 
 Since there are exactly $N=\sum_{g=0}^G n_g$ such semigroups, the time complexity of \textsc{Son} established in Proposition~\ref{P:Son} guarantees that the running time of \textsc{Count} is in~$O(\log(G)\times G\times N)$, as stated.
@@
 \]
-A semigroup is represented by a quadruple $(c(S),g(S),m(S),\delta_S)$. 
+A semigroup $S$ is represented by a quadruple $(c(S),g(S),m(S),\delta_S)$. 
 By relations $ii)$ and $iii)$ of Proposition~\ref{P:Res}, we have $c\leq 2g(S)$ and $m\leq g(S)+1$. 
@@
 \begin{itemize}
+
 \item Vectorization (\MMX, \SSE instructions sets) and careful memory alignment. 
@@
 
-Assume for example that we want to construct the tree~$\mathcal{T}^{\leq 100}$
+Assume for example that we want to construct the tree~$\mathcal{T}_{100}$
 of all numerical semigroup of genus smaller than $100$.  In this case, the
@@
 memory byte $\texttt{T.d[y]}$.  Using these two cursors, Lines \tt{10} to
-\tt{14} of Algorithm~\textsc{Son} can be rewritten as follows
-
-{\small
+\tt{14} of Algorithm~\textsc{Son} can be rewritten as follows :
+
+\begin{figure}
+\small
 \begin{algorithmic}
@@
 \State \textbf{increase} \tt{src},\tt{dst},\tt{i} \textbf{by} $1$
-
 \EndWhile
-\end{algorithmic}}
+\end{algorithmic}
+\end{figure}
 
@@
  16 consecutive entries of the decomposition number
-of the parent (called \tt{src} monoid) in the variable \tt{block}.
+of the parent (called \tt{src} semigroup) in the variable \tt{block}.
 Those entries will be used to compute
@@
 As we previously said, to gain more speed this core loop is actually unrolled
-using some kind of Duff device~\cite{WikipediaDuff} as
-
-{\small
-\begin{algorithmic}
-\State \textbf{Jump to case } $\tt{start}+1$
-\State \textbf{case 1:} \tt{block = ...; dst[1] -= ((block != zero) \& one)}
-\State \textbf{case 2:} \tt{block = ...; dst[2] -= ((block != zero) \& one)}
-\State \textbf{[...]}
-\State \textbf{case NBLOCK-1:} \tt{...}
-\end{algorithmic}}
+using some kind of Duff device~\cite{WikipediaDuff}.
+
+%{\small
+%\begin{algorithmic}
+%\State \textbf{Jump to case } $\tt{start}+1$
+%\State \textbf{case 1:} \tt{block = ...; dst[1] -= ((block != zero) \& one)}
+%\State \textbf{case 2:} \tt{block = ...; dst[2] -= ((block != zero) \& one)}
+%\State \textbf{[...]}
+%\State \textbf{case NBLOCK-1:} \tt{...}
+%\end{algorithmic}}
 
@@
 those kind of problems. For our computation, we used the free version which is
-integrated in the latest version of the GNU C compiler~\cite{GCCcilk}.
+integrated in the latest version of the GNU~C compiler~\cite{GCCcilk}.
 
@@
 \item\tt{cilk\_spawn}: used on a procedure call, indicates that the call can
-  safely operate in parallel with the remaining code if the current
+  safely operate in parallel with the remaining code of the current
   function. Note that the scheduler is not obligated to run this procedure in
@@
 {\small
-
 \begin{algorithm}
-\caption{\Cilk version of Algorithm~\ref{A:ExpRec}}
+\caption{\small \Cilk version of Algorithm~\ref{A:ExpRec}}
 \begin{algorithmic}
@@
 \end{algorithmic}
+
 \end{algorithm}
@@
 explored in parallel.
-
 The thing are actually only a little bit more complicated. First we have to
@@
 \begin{verbatim}
-void explore(const monoid &m) {
+void explore(const Semigroup &S) {
   unsigned long int nbr = 0;
-  if (m.genus < MAX_GENUS - STACK_BOUND) {
-    auto it = generator_iter<CHILDREN>(m);
-    while (it.move_next()) {
-      auto child = remove_generator(m, it.get_gen()).
+  if (S.g < MAX_GENUS - STACK_BOUND) {
+    auto it = generator_iter<CHILDREN>(S); 
+    while (it.move_next()) { //iterate along the children of S
+      auto child = remove_generator(S, it.get_gen()).
       cilk_spawn explore(child);
@@
     }
-    cilk_results[m.genus] += nbr;
-  }
-  else explore_stack(m, cilk_results.get_array());
+    cilk_results[S.g] += nbr;
+  }
+  else explore_stack(S, cilk_results.get_array());
 }
@@
 
-\subsubsection*{Avoiding dynamic allocation} Compared to the \textsc{Son}
-algorithm, dynamic allocation cost orders of magnitude more. Therefore, during
-the derecursived stack algorithm, we only allocate (on the system stack rather
-than on the heap) the stack of monoid. No further allocation are done.
-
 \subsubsection*{Avoiding all unnecessary copy} We also used a trick to avoid
@@
 following sequences of operations:
+
 {\small
@@
 }
+
 If we use a stack of monoid, we can construct $\tt{S}^\tt{x}$ directly into
@@
 \begin{verbatim}
-  monoid data[MAX_GENUS-1], *stack[MAX_GENUS], *current;
-  monoid **stack_pointer = stack + 1;
+  Semigroup data[MAX_GENUS-1], *stack[MAX_GENUS], *current;
+  Semigroup **stack_pointer = stack + 1;
   for (ind_t i=1; i<MAX_GENUS; i++) stack[i] = &(data[i-1]);
@@
 
+\subsubsection*{Avoiding dynamic allocation} Compared to the \textsc{Son}
+algorithm, dynamic allocation cost orders of magnitude more. Therefore, during
+the derecursived stack algorithm, we only allocate (on the system stack rather
+than on the heap) the stack of monoid. No further allocation are done.
+
 \subsubsection*{Pointer arithmetic and indexes}
@@
 
-\subsection{Timings}
-
-Figure~\ref{F:Time} summarizes the time complexity of various exploration algorithms.
-
-The version $\small depth-expl-\delta-mmx$ of our algorithm compute the value of $n_g$ for $g\leq 50$ in $196$ minutes on the \tt{i5-3570K} CPU while the parallel version running on the $4$ cores of the same CPU end the work in $50$ minutes.
-
-Using two \tt{i7} based computers and our  parallel algorithm  we computed in two days the values of $n_g$ for $g\leq 60$, confirming the values given by M.Bras-Amor\'os and M.Delgado :
-
-\begin{minipage}{0.5\textwidth}
-\begin{tabular}{|r|r|r|}
+\section{Results}
+
+Running the \Cilk version of our optimized algorithm  we have explored the tree of numerical semigroup upto genus $67$
+The computation are done on $64$ cores \REM{Ajouter descriptif de la machine + temps mis}.
+The values of $n_g$ for $g\leq 67$ are : 
+\begin{center}
+\begin{tabular}{|r|r||r|r||r|r|}
 \hline
-g & $n_g$ & $n_g/n_{g-1}$\\
+g & $n_g$ & g & $n_g$ & g & $n_g$ \\
 \hline
-0 & 1 &  \\
-1 & 1 & 1.0 \\
-2 & 2 & 2.0 \\
-3 & 4 & 2.0 \\
-4 & 7 & 1.75 \\
-5 & 12 & 1.71428 \\
-6 & 23 & 1.91666 \\
-7 & 39 & 1.69565 \\
-8 & 67 & 1.71794 \\
-9 & 118 & 1.76119 \\
-10 & 204 & 1.72881 \\
-11 & 343 & 1.68137 \\
-12 & 592 & 1.72594 \\
-13 & 1001 & 1.69087 \\
-14 & 1693 & 1.69130 \\
-15 & 2857 & 1.68753 \\
-16 & 4806 & 1.68218 \\
-17 & 8045 & 1.67394 \\
-18 & 13467 & 1.67395 \\
-19 & 22464 & 1.66807 \\
-20 & 37396 & 1.66470 \\
-21 & 62194 & 1.66311 \\
-22 & 103246 & 1.66006 \\
-23 & 170963 & 1.65588 \\
-24 & 282828 & 1.65432 \\
-25 & 467224 & 1.65197 \\
-26 & 770832 & 1.64981 \\
-27 & 1270267 & 1.64791 \\
-28 & 2091030 & 1.64613 \\
-29 & 3437839 & 1.64408 \\
-30 & 5646773 & 1.64253 \\
-31 & 9266788 & 1.64107 \\
-32 & 15195070 & 1.63973 \\
-33 & 24896206 & 1.63843 \\
-
+0 & 1 & 23 & 170963 & 46 & 14463633648\\
+1 & 1 & 24 & 282828 & 47 & 23527845502\\
+2 & 2 & 25 & 467224 & 48 & 38260496374\\
+3 & 4 & 26 & 770832 & 49 & 62200036752\\
+4 & 7 & 27 & 1270267 & 50 & 101090300128\\
+5 & 12 & 28 & 2091030 & 51 & 164253200784\\
+6 & 23 & 29 & 3437839 & 52 & 266815155103\\
+7 & 39 & 30 & 5646773 & 53 & 433317458741\\
+8 & 67 & 31 & 9266788 & 54 & 703569992121\\
+9 & 118 & 32 & 15195070 & 55 & 1142140736859\\
+10 & 204 & 33 & 24896206 & 56 & 1853737832107\\
+11 & 343 & 34 & 40761087 & 57 & 3008140981820\\
+12 & 592 & 35 & 66687201 & 58 & 4880606790010\\
+13 & 1001 & 36 & 109032500 & 59 & 7917344087695\\
+14 & 1693 & 37 & 178158289 & 60 & 12841603251351\\
+15 & 2857 & 38 & 290939807 & 61 & 20825558002053\\
+16 & 4806 & 39 & 474851445 & 62 & 33768763536686\\
+17 & 8045 & 40 & 774614284 & 63 & 54749244915730\\
+18 & 13467 & 41 & 1262992840 & 64 & 88754191073328\\
+19 & 22464 & 42 & 2058356522 & 65 & 143863484925550\\
+20 & 37396 & 43 & 3353191846 & 66 & 233166577125714\\
+21 & 62194 & 44 & 5460401576 & 67 & 377866907506273\\
+22 & 103246 & 45 & 8888486816 & \\
 \hline
 \end{tabular}
-\end{minipage}
-\begin{minipage}{0.5\textwidth}
-\begin{tabular}{|r|r|r|}
+\end{center}
+
+In \cite{Zhai}, A.Zhai establishes that the limit of the quotient   $\frac{n_g}{n_{g-1}}$, when $g$ go to infinity, is the golden ratio $\phi\approx 1.618$.
+As the reader can check the convergence is very slow : the quotient $n_{67}/n_{66}$ is $\approx 1.62$.
+
+\subsection{Wilf's conjecture}
+
+In the paper \cite{Wilf} of 1978, H.S. Wilf conjectured that all numerical semigroup $S$ satisfy the relation
+
+\[
+\card(\Irr(S))\geq \frac{c(S)}{c(S)-g(S)}
+\]
+
+Since the work of M.~Bras-Amor{\'o}s, see \cite{BrasAmoros2008}, we yet know that all numerical semigroup satisfy of genus $g\leq 50$ satisfy the Wilf's conjecture.
+With our exploration algorithm we have proved that there is no counterexample to the Wilf's conjecture of genus $g\leq 60$.
+
+\subsection{Timings}
+
+In this section we summarize the timing improvement through the different optimization of our algorithm.
+All the computations are done on a computer with an \tt{i5-3570K} CPU equipped with \texttt{8GB} of memory.
+
+The following table show the time needed by the algorithm for computing the values of $n_g$ for $g\leq G$ with $30\leq G \leq 40$.
+All the algorithm are executed on only one thread.
+Algorithm \texttt{depth} is based on a baser breadth exploration of the tree while Algorithm \texttt{depth} use a depth exploration.
+Algorithm \texttt{depth+$\delta$} is a refinement of \texttt{depth} based on the the $S$-decomposition function.
+Algorithm \texttt{$\delta$+sse} in an optimization of \texttt{depth+$\delta$} using the \SIMD extension \SSE.
+Times are in second.
+
+\begin{figure}[h!]
+\begin{tabular}{|c|c|c|c|c|c|c|c|c|c|c|c|}
 \hline
-g & $n_g$ & $n_g/n_{g-1}$\\
+\texttt{Algorithm}					&  30 &  31 &  32 &  33 &  34 &   35 &  36 &  37 &  38 &  39 &  40 \\
 \hline
-34 & 40761087 & 1.63724 \\
-35 & 66687201 & 1.63605 \\
-36 & 109032500 & 1.63498 \\
-37 & 178158289 & 1.63399 \\
-38 & 290939807 & 1.63304 \\
-39 & 474851445 & 1.63212 \\
-40 & 774614284 & 1.63127 \\
-41 & 1262992840 & 1.63047 \\
-42 & 2058356522 & 1.62974 \\
-43 & 3353191846 & 1.62906 \\
-44 & 5460401576 & 1.62841 \\
-45 & 8888486816 & 1.62780 \\
-46 & 14463633648 & 1.62723 \\
-47 & 23527845502 & 1.62668 \\
-48 & 38260496374 & 1.62617 \\
-49 & 62200036752 & 1.62569 \\
-50 & 101090300128 & 1.62524 \\
-51 & 164253200784 & 1.62481 \\
-52 & 266815155103 & 1.62441 \\
-53 & 433317458741 & 1.62403 \\
-54 & 703569992121 & 1.62368 \\
-55 & 1142140736859 & 1.62335 \\
-56 & 1853737832107 & 1.62303 \\
-57 & 3008140981820 & 1.62274 \\
-58 & 4880606790010 & 1.62246 \\
-59 & 7917344087695 & 1.62220 \\
-60 & 12841603251351 & 1.62195 \\
-61 & 20825558002053 & 1.62173 \\
-62 & 33768763536686 & 1.62151 \\
-63 & 54749244915730 & 1.62130 \\
-64 & 88754191073328 & 1.62110 \\
-65 & 143863484925550 & 1.62092 \\
-66 & 233166577125714 & 1.62075 \\
-67 & 377866907506273 & 1.62059 \\\hline
+\texttt{breadth}		& 5.0 & 8.3 &  14 &  23 &  38 & 1251 &     &     &     &     &     \\
+\texttt{depth}			& 3.4 & 5.8 & 9.2 &  16 &  27 &   45 &  75 & 125 & 204 & 346 & 557 \\
+\texttt{depth+$\delta$}	& 0.3 & 0.6 & 1.0 & 1.7 & 2.7 &  4.2 & 7.4 &  12 &  20 &  32 & 74  \\
+\texttt{$\delta$+sse}	& 0.1 & 0.2 & 0.3 & 0.4 & 0.8 &  1.2 & 2.0 & 3.1 & 5.1 & 9.0 & 14 \\
+\hline
 \end{tabular}
-\end{minipage}
-\vspace{2em}
-
-In \cite{Zhai}, A.Zhai establishes that the limit of the quotient   $\frac{n_g}{n_{g-1}}$, when $g$ go to $+\infty$, is the golden ratio $\phi\approx 1.618$. 
-As the reader can see the convergence is very slow.
-
-\begin{figure}[b]
-\begin{tikzpicture}[scale=1.3]
-        \begin{semilogyaxis}[xlabel=\small genus,ylabel=time,legend pos=north west,legend cell align=left]
-        \addplot [color=blue,mark=x] coordinates {
-            (20,0.02)
-            (21,0.04)
-            (22,0.07)
-            (23,0.12)
-            (24,0.2)
-            (25,0.35)
-            (26,0.61)
-            (27,1.04)
-            (28,1.76)
-            (29,2.96)
-            (30,4.96)
-            (31,8.26)
-            (32,13.77)
-            (33,22.95)
-            (34,38.22)
-            (35,1250.88)
-        };
-       \addplot [color=black,mark=x] coordinates {
-			(20,0.02)	
-            (21,0.04)
-            (22,0.06)
-            (23,0.09)
-            (24,0.16)
-            (25,0.26)
-            (26,0.45)
-            (27,0.72)
-            (28,1.25)
-            (29,2.1)
-            (30,3.39)
-            (31,5.85)
-            (32,9.25)
-            (33,16.34)
-            (34,27.23)
-            (35,45.39)
-            (36,75.58)
-            (37,125.59)
-            (38,203.56)
-            (39,346.02)
-            (40,557.51)
-        };
-         \addplot [color=red,mark=x] coordinates {
-			(20,0.02)	
-            (21,0.02)
-            (22,0.04)
-            (23,0.06)
-            (24,0.1)
-            (25,0.18)
-            (26,0.29)
-            (27,0.48)
-            (28,0.81)
-            (29,1.34)
-            (30,2.25)
-            (31,3.73)
-            (32,6.2)
-            (33,10.35)
-            (34,17.11)
-            (35,28.35)
-            (36,46.99)
-            (37,77.89)
-            (38,128.53)
-            (39,212.1)
-            (40,349.98)
-        };
-        \addplot [color=magenta,mark=x] coordinates {
-			(20,0.01)	
-            (21,0.01)
-            (22,0.01)
-            (23,0.01)
-            (24,0.02)
-            (25,0.03)
-            (26,0.05)
-            (27,0.08)
-            (28,0.12)
-            (29,0.21)
-            (30,0.35)
-            (31,0.57)
-            (32,0.97)
-            (33,1.65)
-            (34,2.69)
-            (35,4.18)
-            (36,7.36)
-            (37,11.66)
-            (38,20.35)
-            (39,32.55)
-            (40,74.46)
-        };
-	%\legend style={anchor=north,at={(20,1000)}}  
-        \legend{\tiny breadth-expl,\tiny depth-expl,\tiny depth-expl-$\delta$,\tiny depth-expl-$\delta$-mmx}
-        \end{semilogyaxis}
-    \end{tikzpicture}
-    \label{F:Time:new}
-    \caption{Comparison of time executions of different exploration algorithms of the tree $\mathcal{T}^{\leq g}$ on an Intel \tt{i5-3570K} CPU with $8GB$ of memory. 
-The scale is logarithmic in time. 
-The first version is based on a breadth-first search exploration and the peak at genus~$35$ is due to the consumption of all the memory and use of swap. 
-The second version uses a depth first search exploration.
-The third is based on the second one with use of decomposition numbers, it corresponds to the algorithm given in Section~\ref{S:Algo}. 
-The last one is an optimisation of the third one with \MMX.}
 \end{figure}
 
-\begin{figure}[b]
-\begin{tikzpicture}[scale=1.3]
-        \begin{semilogyaxis}[xlabel=\small genus,ylabel=time,legend pos=north west,legend cell align=left]
-        \addplot [color=blue,mark=x] coordinates {
-(20,0.045767)
-(21,0.067697)
-(22,0.112697)
-(23,0.194646)
-(24,0.335412)
-(25,0.581481)
-(26,1.01088)
-(27,1.72613)
-(28,2.9324)
-(29,4.91856)
-(30,8.39383)
-(31,14.0308)
-(32,23.7138)
-(33,39.6461)
-(34,66.1154)
-(35,109.406)
-(36,182.331)
-(37,302.765)
-        };
-       \addplot [color=black,mark=x] coordinates {
-(20,0.043697)
-(21,0.065487)
-(22,0.113301)
-(23,0.186903)
-(24,0.31748)
-(25,0.53251)
-(26,0.899982)
-(27,1.5205)
-(28,2.55423)
-(29,4.28809)
-(30,7.30236)
-(31,12.1284)
-(32,20.2132)
-(33,33.707)
-(34,56.5415)
-(35,93.7614)
-(36,157.09)
-(37,259.997)
-(38,427.497)
-(39,709.564)
-(40,1192.81)
-        };
-         \addplot [color=red,mark=x] coordinates {
-(20,0.027708)
-(21,0.044416)
-(22,0.074789)
-(23,0.122006)
-(24,0.203155)
-(25,0.344143)
-(26,0.594073)
-(27,1.07134)
-(28,1.622)
-(29,2.82573)
-(30,4.53882)
-(31,7.60975)
-(32,12.5442)
-(33,21.0466)
-(34,34.7638)
-(35,57.7621)
-(36,95.6878)
-(37,158.95)
-(38,264.495)
-(39,436.146)
-(40,719.347)
-        };
-        \addplot [color=magenta,mark=x] coordinates {
-(20,0.001562)
-(21,0.001957)
-(22,0.002711)
-(23,0.003811)
-(24,0.005605)
-(25,0.00861)
-(26,0.013127)
-(27,0.020991)
-(28,0.036053)
-(29,0.05879)
-(30,0.095882)
-(31,0.158853)
-(32,0.259237)
-(33,0.441074)
-(34,0.750349)
-(35,1.2067)
-(36,1.9608)
-(37,3.19713)
-(38,5.33436)
-(39,9.2136)
-(40,14.5226)
-        };
-	%\legend style={anchor=north,at={(20,1000)}}  
-        \legend{\tiny breadth-expl,\tiny depth-expl,\tiny depth-expl-$\delta$}
-        \end{semilogyaxis}
-    \end{tikzpicture}
-    \label{F:Time}
-    \caption{Comparison of time executions of different exploration algorithms of the tree $\mathcal{T}^{\leq g}$ on an Intel \tt{i7-3840QM} CPU with $32GB$ of memory. 
-The scale is logarithmic in time. 
-The first version is based on a breadth-first search exploration and the peak at genus~$35$ is due to the consumption of all the memory and use of swap. 
-The second version uses a depth first search exploration.
-The third is based on the second one with use of decomposition numbers, it corresponds to the algorithm given in Section~\ref{S:Algo}. 
-The last one is an optimisation of the third one with \SSE but not using \CilkP.}
+The computation of $n_g$ for $g\leq 33$ with algorithm \texttt{breadth} is very long because all the \texttt{8GB} of memory are consumed and the system must use swap memory to finish the computation.
+This algorithm was not launched
+
+On the following table we illustrate the impact of parallelization.
+The test are also be performed on the \tt{i5-3570K} CPU.
+This CPU has two physics cores that are viewed as four virtual cores by the system.
+Therefore we have decided to test the parallelization of our algorithm using \Cilk upto 4 threads.
+The time of the one threaded algorithm  must be compared with the \texttt{$delta$+sse} version of the previous table : it illustrate the additional cost induced by the use of \Cilk technology.
+
+  \begin{figure}[h!]
+\begin{tabular}{|c|c|c|c|c|c|}
+\hline
+Threads & 30 & 35 & 40 & 45 & 50 \\
+\hline
+1 &  0.11 & 1.26 & 14.9 & 182 & 2201 \\ 
+2 &  0.06 & 0.65 & 7.50 &  92 & 1110\\
+3 &  0.05 & 0.44 & 5.14 &  63 & 747 \\
+4 &  0.04 & 0.34 & 4.02 &  48 & 489 \\
+\hline
+\end{tabular}
 \end{figure}
+
+It should be mentioned that the hyper threading of the CPU has not been deactivated : the CPU clock is higher using one thread than  two and so on.
+
+\textbf{Acknowledgment.} We would like to thank Shalom Eliahou for his presentation of the tree of numerical semigroups and for the great full discussions on the subject.
+
+%\begin{figure}[b]
+%\begin{tikzpicture}[scale=1.3]
+%\begin{semilogyaxis}[xlabel=\small genus,ylabel=time,legend pos=north west,legend cell align=left]
+%\addplot[color=blue,mark=x] coordinates {(20,0.02)(21,0.04) (22,0.07)(23,0.12)(24,0.2)(25,0.35)(26,0.61)(27,1.04) (28,1.76)(29,2.96)(30,4.96)(31,8.26)(32,13.77)(33,22.95)(34,38.22)(35,1250.88)};
+%\addplot[color=black,mark=x] coordinates {(20,0.02)(21,0.04) (22,0.06)(23,0.09)(24,0.16)(25,0.26)(26,0.45)(27,0.72)(28,1.25)(29,2.1)(30,3.39)(31,5.85)(32,9.25)(33,16.34)(34,27.23)(35,45.39)(36,75.58)(37,125.59)(38,203.56)(39,346.02)(40,557.51)};
+%\addplot[color=red,mark=x] coordinates {(20,0.02)(21,0.02)(22,0.04)(23,0.06)(24,0.1)(25,0.18)(26,0.29)(27,0.48)(28,0.81)(29,1.34)(30,2.25)(31,3.73)(32,6.2)(33,10.35)(34,17.11)(35,28.35)(36,46.99)(37,77.89)(38,128.53)(39,212.1)(40,349.98)};
+%\addplot[color=magenta,mark=x] coordinates {(20,0.01)(21,0.01)(22,0.01)(23,0.01)(24,0.02)(25,0.03)(26,0.05)(27,0.08)(28,0.12)(29,0.21)(30,0.35)(31,0.57)(32,0.97)(33,1.65)(34,2.69)(35,4.18(36,7.36)(37,11.66)(38,20.35)(39,32.55)(40,74.46) };
+%          \addplot [color=green,mark=x] coordinates {
+%			(20,0.003)	
+%            (21,0.004)
+%            (22,0.006)
+%            (23,0.009)
+%            (24,0.011)
+%            (25,0.014)
+%            (26,0.018)
+%            (27,0.029)
+%            (28,0.046)
+%            (29,0.069)
+%            (30,0.110)
+%            (31,0.165)
+%            (32,0.261)
+%            (33,0.417)
+%            (34,0.755)
+%            (35,1.192)
+%            (36,1.956)
+%            (37,3.134)
+%            (38,5.089)
+%            (39,8.973)
+%            (40,14.336)
+%        };
+%	%\legend style={anchor=north,at={(20,1000)}}  
+%        \legend{\tiny breadth-expl,\tiny depth-expl,\tiny depth-expl-$\delta$,\tiny depth-expl-$\delta$-mmx}
+%        \end{semilogyaxis}
+%    \end{tikzpicture}
+%    \label{F:Time:new}
+%    \caption{Comparison of time executions of different exploration algorithms of the tree $\mathcal{T}^{\leq g}$ on an Intel \tt{i5-3570K} CPU with $8GB$ of memory. 
+%The scale is logarithmic in time. 
+%The first version is based on a breadth-first search exploration and the peak at genus~$35$ is due to the consumption of all the memory and use of swap. 
+%The second version uses a depth first search exploration.
+%The third is based on the second one with use of decomposition numbers, it corresponds to the algorithm given in Section~\ref{S:Algo}. 
+%The last one is an optimisation of the third one with \MMX.}
+%\end{figure}
+
+%%\begin{figure}[b]
+%%\begin{tikzpicture}[scale=1.3]
+%%        \begin{semilogyaxis}[xlabel=\small genus,ylabel=time,legend pos=north west,legend cell align=left]
+%%        \addplot [color=blue,mark=x] coordinates {
+%%(20,0.045767)
+%%(21,0.067697)
+%%(22,0.112697)
+%%(23,0.194646)
+%%(24,0.335412)
+%%(25,0.581481)
+%%(26,1.01088)
+%%(27,1.72613)
+%%(28,2.9324)
+%%(29,4.91856)
+%%(30,8.39383)
+%%(31,14.0308)
+%%(32,23.7138)
+%%(33,39.6461)
+%%(34,66.1154)
+%%(35,109.406)
+%%(36,182.331)
+%%(37,302.765)
+%%        };
+%%       \addplot [color=black,mark=x] coordinates {
+%%(20,0.043697)
+%%(21,0.065487)
+%%(22,0.113301)
+%%(23,0.186903)
+%%(24,0.31748)
+%%(25,0.53251)
+%%(26,0.899982)
+%%(27,1.5205)
+%%(28,2.55423)
+%%(29,4.28809)
+%%(30,7.30236)
+%%(31,12.1284)
+%%(32,20.2132)
+%%(33,33.707)
+%%(34,56.5415)
+%%(35,93.7614)
+%%(36,157.09)
+%%(37,259.997)
+%%(38,427.497)
+%%(39,709.564)
+%%(40,1192.81)
+%%        };
+%%         \addplot [color=red,mark=x] coordinates {
+%%(20,0.027708)
+%%(21,0.044416)
+%%(22,0.074789)
+%%(23,0.122006)
+%%(24,0.203155)
+%%(25,0.344143)
+%%(26,0.594073)
+%%(27,1.07134)
+%%(28,1.622)
+%%(29,2.82573)
+%%(30,4.53882)
+%%(31,7.60975)
+%%(32,12.5442)
+%%(33,21.0466)
+%%(34,34.7638)
+%%(35,57.7621)
+%%(36,95.6878)
+%%(37,158.95)
+%%(38,264.495)
+%%(39,436.146)
+%%(40,719.347)
+%%        };
+%%        \addplot [color=magenta,mark=x] coordinates {
+%%(20,0.001562)
+%%(21,0.001957)
+%%(22,0.002711)
+%%(23,0.003811)
+%%(24,0.005605)
+%%(25,0.00861)
+%%(26,0.013127)
+%%(27,0.020991)
+%%(28,0.036053)
+%%(29,0.05879)
+%%(30,0.095882)
+%%(31,0.158853)
+%%(32,0.259237)
+%%(33,0.441074)
+%%(34,0.750349)
+%%(35,1.2067)
+%%(36,1.9608)
+%%(37,3.19713)
+%%(38,5.33436)
+%%(39,9.2136)
+%%(40,14.5226)
+%%        };
+%%	%\legend style={anchor=north,at={(20,1000)}}  
+%%        \legend{\tiny breadth-expl,\tiny depth-expl,\tiny depth-expl-$\delta$}
+%%        \end{semilogyaxis}
+%%    \end{tikzpicture}
+%%    \label{F:Time}
+%%    \caption{Comparison of time executions of different exploration algorithms of the tree $\mathcal{T}^{\leq g}$ on an Intel \tt{i7-3840QM} CPU with $32GB$ of memory. 
+%%The scale is logarithmic in time. 
+%%The first version is based on a breadth-first search exploration and the peak at genus~$35$ is due to the consumption of all the memory and use of swap. 
+%%The second version uses a depth first search exploration.
+%%The third is based on the second one with use of decomposition numbers, it corresponds to the algorithm given in Section~\ref{S:Algo}. 
+%%The last one is an optimisation of the third one with \SSE but not using \CilkP.}
+%%\end{figure}
 
@@
 
+\bibitem{Wilf}{H. S. Wilf, \emph{A circle-of-lights algorithm for the money-changing problem}, Amer. Math. Monthly \textbf{85} (1978), 562--565}
+
 \bibitem{Zhai}{A. Zhai, \emph{Fibonacci-like growth of numerical semigroups of a given genus}, Semigroup Forum\textbf{86}, no.3 (2013), 634--662}

### 2014-09-22 Papier/article.tex
Fix a mistake on Acknowledgment

@@
 
-\textbf{Acknowledgment.} We would like to thank Shalom Eliahou for his presentation of the tree of numerical semigroups and for the great full discussions on the subject.
+\textbf{Acknowledgment.} We would like to thank Shalom Eliahou for his presentation of the tree of numerical semigroups and for the interesting discussions on the subject.
 

### 2014-09-30 Papier/article.tex
Florent table

@@
 first to present our new encoding of the semigroup and the associated
-algorithms, and second to present the optimization technique which allows for
+algorithms, and second to present the optimization techniques which allows for
 those kinds of algorithms to get speedups by factors of hundreds and even
-thousand. We claim that these technique are fairly general for those kinds of
+thousands. We claim that these technique are fairly general for those kinds of
 algorithms. As a support for the claim, we applied it to an algorithm of
@@
 
-Running the \Cilk version of our optimized algorithm  we have explored the tree of numerical semigroup upto genus $67$
-The computation are done on $64$ cores \REM{Ajouter descriptif de la machine + temps mis}.
-The values of $n_g$ for $g\leq 67$ are : 
+Running the \Cilk version of our optimized algorithm we have explored the tree
+of numerical semigroup upto genus $67$. The computation were done one a shared
+$64$ core AMD Opteron\texttrademark{} Processor 6276. As other heavy
+computation were running on the machine, we only used $32$ cores. The
+computation took $18$~days. The values of $n_g$ for $g\leq 67$ are :
 \begin{center}
@@
 The computation of $n_g$ for $g\leq 33$ with algorithm \texttt{breadth} is very long because all the \texttt{8GB} of memory are consumed and the system must use swap memory to finish the computation.
-This algorithm was not launched
-
+This algorithm was not launched.
+
+\Jean{
 On the following table we illustrate the impact of parallelization.
@@
 \end{figure}
-
-It should be mentioned that the hyper threading of the CPU has not been deactivated : the CPU clock is higher using one thread than  two and so on.
-
-\textbf{Acknowledgment.} We would like to thank Shalom Eliahou for his presentation of the tree of numerical semigroups and for the interesting discussions on the subject.
+}
+
+On the following table we illustrate the impact of parallelization. The test
+were performed on a machine holding two \tt{Intel\texttrademark{}
+  Xeon\texttrademark{} X5650} CPU running at $2.67~GHz$. Each of those CPU
+have 6 physical cores that are able to run 12 threads. However, due to Hyper
+threading, when more than 12 thread are running, the computation engine are
+shared between two threads. Therefore, the speedup should by much less when
+adding more core. We can see that this is indeed happening under the
+horizontal bar in the following table.
+  \begin{figure}[h!]
+\begin{tabular}{|c|c|c|c|c|c|}
+\hline
+Threads & 30 & 35 & 40 & 45 & 50 \\
+\hline
+1 &  0.33 & 2.50 & 26.1 & 307 & 3709 \\
+2 &  0.17 & 1.25 & 13.05 & 153.5 & 1110\\
+3 &  0.11 & 0.71 & 8.57 & 102.6 & 747 \\
+4 &  0.10 & 0.55 & 6.45 &  77.21 & 489 \\
+8 &  0.10 & 0.36 & 3.55 &  40.22 & 489 \\
+12 &  0.10 & 0.36 & 2.61 & 27.34 & 489 \\\hline
+16 &  0.10 & 0.37 & 2.50 & 25.96 & 489 \\
+24 &  0.12 & 0.51 & 3.18 & 25.27 & 489 \\
+\hline
+\end{tabular}
+\end{figure}
+
+\textbf{Acknowledgment.} We would like to thank Shalom Eliahou for his
+presentation of the tree of numerical semigroups and for the interesting
+discussions on the subject. We also would like to thanks Nathan Cohen who
+suggested the two level of indirection stack optimization.
 

### 2014-10-02 Papier/article.tex
More values to table
Completed table
MSC + keyword + adresses Mix of i5 + Xeon multi threads

@@
 \usepackage{xspace}
+\usepackage{pdfsync}
+
 \theoremstyle{plain}
@@
 \newtheorem{fact}[prop]{Fact}
-
-\title{Exploring the tree of numerical semigroups}
-\author{Jean Fromentin and  Florent Hivert}
 
@@
 \def\labelitemi{--}
+
+\title{Exploring the tree of numerical semigroups}
+\author{Jean Fromentin and  Florent Hivert}
+\date{2 October 2014}
+\keywords{semigroups, tree, algorithm, \texttt{SSE}, 
+\texttt{Cilk}, optimization}
+\subjclass[2010]{05A15,68R05,68W10}
+
 \begin{document}
@@
 \State Stack \tt{stack} \Comment{the empty stack}
-\State $\tt{S} \gets\left<1\right>$
-\While{stack is not empty} 
+\State \tt{stack.push($\mathbb{N}$)}
+\While{\tt{stack} is not empty} 
 	\State $\tt{S} \gets \tt{stack.top()}$
@@
 			\If{$\tt{x}\in \Irr(\tt{S})$}
-				\State $\tt{S.push}(\tt{S}^\tt{x})$
+				\State $\tt{stack.push}(\tt{S}^\tt{x})$
 			\EndIf
@@
 \State Stack \tt{stack} \Comment{the empty stack}
-\State $\tt{S} \gets \textsc{Root}(\tt{G})$
-\While{stack is not empty} 
+\State \tt{stack.push(\textsc{Root}(\tt{G}))}
+\While{\tt{stack} is not empty} 
 	\State $\tt{S} \gets \tt{stack.top()}$
@@
 			\If{$\tt{S.d[x]\,=\,1}$}			
-				\State $\tt{S.push}(\textsc{Son}(\tt{S},\tt{x},\tt{G}))$
+			      \State 
+$\tt{stack.push}(\textsc{Son}(\tt{S},\tt{x},\tt{G}))$
 			\EndIf
@@
 			\If{$\tt{x}\in \Irr(\tt{S})$}
-				\State\textbf{CilkSpawn} \Call{ExploreRec}{$\tt{S}^\tt{x}, \tt{G}$}
+				\State\textbf{cilk\_spawn} 
+\Call{ExploreRec}{$\tt{S}^\tt{x}, \tt{G}$}
 			\EndIf
@@
 
-In this section we summarize the timing improvement through the different optimization of our algorithm.
-All the computations are done on a computer with an \tt{i5-3570K} CPU equipped with \texttt{8GB} of memory.
-
-The following table show the time needed by the algorithm for computing the values of $n_g$ for $g\leq G$ with $30\leq G \leq 40$.
-All the algorithm are executed on only one thread.
-Algorithm \texttt{depth} is based on a baser breadth exploration of the tree while Algorithm \texttt{depth} use a depth exploration.
-Algorithm \texttt{depth+$\delta$} is a refinement of \texttt{depth} based on the the $S$-decomposition function.
-Algorithm \texttt{$\delta$+sse} in an optimization of \texttt{depth+$\delta$} using the \SIMD extension \SSE.
-Times are in second.
+In this section we summarize the timing improvement through the different 
+optimization of our algorithm.
+
+The following table show the time needed by the algorithm for computing 
+the values of $n_g$ 
+for $g\leq G$ with $30\leq G \leq 40$. 
+on a machine equipped with  an \tt{Intel\texttrademark{} i5-3570K} CPU 
+and \texttt{8GB} of memory.
+All algorithms are executed on only one thread. Algorithm \texttt{depth} is 
+based on a breadth exploration of the tree while Algorithm \texttt{depth} 
+use a depth exploration. Algorithm \texttt{depth+$\delta$} is a refinement of 
+\texttt{depth} based on the the $S$-decomposition function. Algorithm 
+\texttt{$\delta$+sse} in an optimization of \texttt{depth+$\delta$} using the 
+\SIMD extension \SSE. Times are in second.
 
@@
 
-The computation of $n_g$ for $g\leq 33$ with algorithm \texttt{breadth} is very long because all the \texttt{8GB} of memory are consumed and the system must use swap memory to finish the computation.
-This algorithm was not launched.
-
-\Jean{
-On the following table we illustrate the impact of parallelization.
-The test are also be performed on the \tt{i5-3570K} CPU.
-This CPU has two physics cores that are viewed as four virtual cores by the system.
-Therefore we have decided to test the parallelization of our algorithm using \Cilk upto 4 threads.
-The time of the one threaded algorithm  must be compared with the \texttt{$delta$+sse} version of the previous table : it illustrate the additional cost induced by the use of \Cilk technology.
+The computation of $n_g$ for $g\leq 33$ with algorithm \texttt{breadth} is very 
+long because all the \texttt{8GB} of memory are consumed and the system 
+must use swap memory to finish the computation. This algorithm was not launched 
+for genus~$g\geq 36$.
+
+The following table illustrates the impact of parallelization on the same 
+machine based on the \tt{Intel\texttrademark{} i5-3570K} CPU 
+running at \texttt{3.4GHz}. This CPU has two physical cores that are able 
+to run four threads thanks to the \texttt{Hyper-Threading} 
+technology~\cite{WikipediaHT}. However, when more 
+than two threads are running, the computation engine are shared between two 
+threads.
 
@@
 \end{figure}
-}
-
-On the following table we illustrate the impact of parallelization. The test
-were performed on a machine holding two \tt{Intel\texttrademark{}
-  Xeon\texttrademark{} X5650} CPU running at $2.67~GHz$. Each of those CPU
-have 6 physical cores that are able to run 12 threads. However, due to Hyper
-threading, when more than 12 thread are running, the computation engine are
-shared between two threads. Therefore, the speedup should by much less when
-adding more core. We can see that this is indeed happening under the
-horizontal bar in the following table.
+
+The time of the one threaded algorithm  must be compared with the 
+\texttt{$\delta$+sse} version of the previous table : it illustrates the 
+additional cost induced by the use of \Cilk technology.
+
+It should be notice that the \texttt{TurboBoost} technology \cite{WikipediaTB} 
+is present on the CPU. Therefore the clock of the CPU is a bit higher when the 
+number of thread 
+is smaller. Therefore the gain of using more threads is a bit over than one 
+suggested by the table.
+
+Finally we test our algorithm on a machine holding two 
+\tt{Intel\texttrademark{} Xeon\texttrademark{} X5650} CPU running 
+at \texttt{2.67GHz}.
+Each of those CPU have 6 physical cores that are able to run 12 threads : the 
+machine has also 12 physical cores.  
+For more than $12$ threads, the speedup should be much less when
+adding more core. The following table resume the time needed by the algorithm 
+to explore $\mathcal{T}_{50}$ on diffenent number of threads (the double 
+vertical line delimites the use of \texttt{Hyper Threadind}):
 \begin{figure}[h!]
-\begin{tabular}{|c|c|c|c|c|c|}
+\begin{tabular}{|c|c|c|c|c|c||c|c|c|}
 \hline
-Threads & 30 & 35 & 40 & 45 & 50 \\
+ Threads & 1 & 2 & 4 & 8 & 12 & 16 & 20 & 24\\
  \hline
-1 &  0.33 & 2.50 & 26.1 & 307 & 3709 \\
-2 &  0.17 & 1.25 & 13.05 & 153.5 & 1110\\
-3 &  0.11 & 0.71 & 8.57 & 102.6 & 747 \\
-4 &  0.10 & 0.55 & 6.45 &  77.21 & 489 \\
-8 &  0.10 & 0.36 & 3.55 &  40.22 & 489 \\
-12 &  0.10 & 0.36 & 2.61 & 27.34 & 489 \\\hline
-16 &  0.10 & 0.37 & 2.50 & 25.96 & 489 \\
-24 &  0.12 & 0.51 & 3.18 & 25.27 & 489 \\
+ Time (s) & 3709 & 1865 & 932.4 & 486.8 & 325.7 & 311.2 & 302.3 & 290.2\\
  \hline
@@
 \end{figure}
+% 
+% We can see that this is indeed happening under the
+% 
+% horizontal bar in the following table.
+%   \begin{figure}[h!]
+% \begin{tabular}{|c|c|c|c|c|c|}
+% \hline
+% Threads & 30 & 35 & 40 & 45 & 50 \\
+% \hline
+% 1 &  0.33 & 2.50 & 26.1 & 307 & 3709 \\
+% 2 &  0.17 & 1.25 & 13.05 & 153.5 & 1865\\
+% 3 &  0.11 & 0.71 & 8.57 & 102.6 & 1243 \\
+% 4 &  0.10 & 0.55 & 6.45 &  77.21 & 932.4 \\
+% 8 &  0.10 & 0.36 & 3.55 &  40.22 & 486.8 \\
+% 12 &  0.10 & 0.36 & 2.61 & 27.34 & 325.7 \\\hline
+% 16 &  0.10 & 0.37 & 2.50 & 25.96 & 311.2 \\
+% 20 &  0.10 & 0.37 & 2.69 & 26.04 & 302.3 \\
+% 24 &  0.12 & 0.51 & 3.18 & 25.27 & 290.2 \\
+% \hline
+% \end{tabular}
+% \end{figure}
 
@@
 
+\bibitem{WikipediaHT}{Wikipedia, \emph{Hyper-Threading}, 
+\url{http://en.wikipedia.org/wiki/Hyper-Threading}}
+
 \bibitem{WikipediaSIMD}{Wikipedia, \emph{SIMD}, \url{http://en.wikipedia.org/wiki/SIMD}}
@@
 \bibitem{WikipediaSSE}{Wikipedia, \emph{Streaming SIMD Extensions}, \url{http://en.wikipedia.org/wiki/Streaming_SIMD_Extensions}}
+
+\bibitem{WikipediaTB}{Wikipedia, \emph{Turbo Boost}, 
+\url{http://en.wikipedia.org/wiki/Intel_Turbo_Boost}}
 
@@
 
+\vspace{2em}
+
+\noindent \textbf{Jean Fromentin}
+
+Univ. Lille Nord de France, F-59000 Lille, France
+
+ULCO, LMPA J.~Liouville, B.P. 699, F-62228 Calais, France
+
+CNRS, FR 2956, France
+
+\url{fromentin@lmpa.univ-littoral.fr}
+
+\vspace{2em}
+
+\noindent \textbf{Florent Hivert}
+
+XXX
+
+\url{florent.hivert@lri.fr}
 \end{document}

### 2014-10-09 Papier/article.tex
Florent's rereading

@@
 first to present our new encoding of the semigroup and the associated
-algorithms, and second to present the optimization techniques which allows for
-those kinds of algorithms to get speedups by factors of hundreds and even
+algorithms, and second to present the optimization techniques which allows, for
+those kinds of algorithms, to get speedups by factors of hundreds and even
 thousands. We claim that these technique are fairly general for those kinds of
@@
 For more than $12$ threads, the speedup should be much less when
-adding more core. The following table resume the time needed by the algorithm 
-to explore $\mathcal{T}_{50}$ on diffenent number of threads (the double 
-vertical line delimites the use of \texttt{Hyper Threadind}):
+adding more core. The following table resumes the time needed by the algorithm 
+to explore $\mathcal{T}_{50}$ on different number of threads (the double 
+vertical line delimites the use of \texttt{Hyper Threadind}). For reference,
+we put in the column called ``C++'' the computation time of the same
+program compiled without Cilk.
 \begin{figure}[h!]
-\begin{tabular}{|c|c|c|c|c|c||c|c|c|}
+\begin{tabular}{|c|c|c|c|c|c|c||c|c|c|}
 \hline
- Threads & 1 & 2 & 4 & 8 & 12 & 16 & 20 & 24\\
+ Threads & C++ & 1 & 2 & 4 & 8 & 12 & 16 & 20 & 24\\
  \hline
- Time (s) & 3709 & 1865 & 932.4 & 486.8 & 325.7 & 311.2 & 302.3 & 290.2\\
+ Time (s) &      & 3709 & 1865 & 932.4 & 486.8 & 325.7 & 311.2 & 302.3 & 290.2\\
  \hline
@@
 \end{figure}
+
 % 
@@
 
-XXX
+Bureau 33, Laboratoire de Recherche en Informatique (UMR CNRS 8623)
+
+Bâtiment 650, Universit\'e Paris Sud 11, 91405 Orsay CEDEX, France
 

### 2014-10-13 Papier/article.tex
Reference Timing
Read 1 of the paper

@@
-\documentclass[reqno,11pt]{amsart}
+\documentclass[reqno]{amsart}
 \usepackage{amsmath, amssymb, amsthm}
@@
 \newtheorem{fact}[prop]{Fact}
+
+\setlength{\oddsidemargin}{0pt} 
+\setlength{\evensidemargin}{0pt}
+\setlength{\textwidth}{16cm}
+\setlength{\textheight}{24.2cm}
+\setlength{\topmargin}{0pt}
+\setlength{\voffset}{-1.04cm}
+\setlength{\headsep}{0.5cm}
+\setlength{\headheight}{0pt}
 
@@
 
+\maketitle
+
 \begin{abstract}
@@
 
-\maketitle
+\vspace{1em}
 
@@
 is a numerical semigroup.
-The \emph{genus} of a numerical semigroup $S$, denoted by~$g(S)$, is the cardinality of $\NN\setminus S$.
+The \emph{genus} of a numerical semigroup $S$, denoted by~$g(S)$, is the cardinality of~$\NN\setminus S$.
  For example the genus of $S_E$ is $6$,  the cardinality of $\{1,2,4,5,8,11\}$
@@
 N.~Borie enumerating integer vector modulo permutation groups~\cite{Borie} and
-got a speedup by a factor larger than 2000 using 8 cores !
+got a speedup by a factor larger than 2000 using 8 cores.
 
@@
 Section~\ref{S:Opti} is more technical, and is devoted to the optimisation of the algorithm introduced in Section~\ref{S:Algo}.
-Il the last section we emphase the results obtained using our algorithm.
+In the last section we emphase the results obtained using our algorithm.
 
@@
 
-By definition a numerical semigroup is infinite object and we need a finite description of such a object. 
+By definition a numerical semigroup is an infinite object and we need a finite description of such a object. 
 That is the role of generating sets.
@@
 \]
-that send $S\{0,...,c(S)-1\}$ into $\{0,...,c(S)-1\}\setminus S$ which is  of cardinality $g(S)$. 
+that send $S\{0,...,c(S)-1\}$ into $\{0,...,c(S)-1\}\setminus S$ which is  of cardinality~$g(S)$. 
 Let us denote by $k$ the cardinality  of $S\cap\{0,...,c(S)-1\}$. 
@@
 A consequence of Proposition~ \ref{P:Res}~$i)$ is that $\Irr(S)$ is finite and its cardinality is at most $c(S)+m(S)-1$.
-More precisely, using Ap\'ery set, we can prove that  the cardinality of $\Irr(S)$ is at most $m(S)$. See Section 2 of Chapter I of~\cite{BookNS} for more details.
+More precisely, using Ap\'ery set, we can prove that  the cardinality of $\Irr(S)$ is at most $m(S)$. See Section~2 of Chapter I of~\cite{BookNS} for more details.
 
@@
 If $x$ is not irreducible in  $S$, then there exist $a$ and $b$ in $S\setminus{0}$ such that $x=a+b$. 
-Since $a\not=0$ and $b\not=0$ hold, the integers  $a$ and $b$ belong to $S\setminus\{x\}$. 
+As $a\not=0$ and $b\not=0$ hold, the integers  $a$ and $b$ belong to $S\setminus\{x\}$. 
 Since $x=a+b$ and $x\not\in S^x$, it follows that  $S^x$ is not stable under addition.
@@
 With this construction, a semigroup $S$ has depth $g$ in $\mathcal{T}$ if and only if its genus is~$g$, see Figure~\ref{F:Tree}.
-We denote by $\mathcal{T}_{g}$ the subtree of $\mathcal{T}$ resticted to all semigroup of genus $\leq g$.
-
-\vspace{-3em}
+We denote by $\mathcal{T}_{g}$ the subtree of~$\mathcal{T}$ resticted to all semigroup of genus $\leq g$.
 
@@
   \]
-\caption{The first four layers of the tree $\mathcal{T}$ of numerical semigroups, corresponding to $\mathcal{T}_4$.
-A generator of a semigroup is in gray if is not greater that $c(S)$.
+\caption{The first four layers of the tree $\mathcal{T}$ of numerical semigroups, corresponding to~$\mathcal{T}_4$.
+A generator of a semigroup is in gray if is not greater than $c(S)$.
 An edge between a semigroup $S$ and its son $S'$ is labelled by  $x$ is $S'$ is obtained from $S$ by removing~$x$, that is $S'=S^x$.}
@@
 Reconsider the semigroup $S_E$ given at \eqref{E:NSG}.
-The integer $14$ admits three decompositions as sums of two elements of $S$, namely $14=0+14$, $14=3+11$ and $14=7+7$. 
-Thus the set $D_{S_E}(14)$ is equal to $\{0,3,7\}$ and $d_{S_E}$ equals $3$.
+The integer $14$ admits three decompositions as sums of two elements of $S$, namely $14=0+14$ and $14=7+7$. 
+Thus the set $D_{S_E}(14)$ is equal to $\{0,7\}$ and $d_{S_E}$ equals $2$.
 \end{exam}
@@
 \begin{proof}
-As the set $D_S(x)$ is included in $\{0,...,\left\lfloor\frac x2\right\rfloor\}$, the relation $d_S(x)\leq 1+\left\lfloor \frac{x}2\right\rfloor$ holds.
-For $S=\NN$ we have the equality for $D_S(x)$ and so for $d_S(x)$.
+As the set $D_S(x)$ is included in $\left\{0,...,\left\lfloor\frac x2\right\rfloor\right\}$, the relation $d_S(x)\leq 1+\left\lfloor \frac{x}2\right\rfloor$ holds.
+For $S=\NN$ we have the equality for the set~$D_S(x)$ and so for the integer $d_S(x)$.
 \end{proof}
@@
 The relation~$2\times 0\leq x$ and $0\in S$ imply that $D_S(x)$ contains $0$, and so $d_S(x)>0$ holds.
-Conversely, the relation  $d_S(x)>0$  implies that $D_S(x)$ is non-empty. 
+Conversely, the relation  $d_S(x)>0$  implies that~$D_S(x)$ is non-empty. 
 Let $y$ be an element of~$D_S(x)$. 
@@
 Assume $x$ is irreducible in $S$.
-There cannot exist $y$ and $z$ in~$(S\setminus\{0\})^2$ such that $x=y+z$. 
+There cannot exist $y$ and $z$ in~$(S\setminus\{0\})^2$ such that~$x=y+z$. 
 The only possible decomposition of $x$ as a sum of two elements of $S$ is $x=0+x$. 
@@
 Conversely, let $x$ such that $d_S(x)$ is $1$. 
-By  $i)$ the  integer $x$ must be in $S$. 
+By~$i)$ the  integer $x$ must be in $S$. 
 As $x=0+x$ is always a decomposition of $x$ as a sum of two elements in $S$, we obtain $D_S(x)=\{0\}$.
-If there exist $y$ and $z$ in $S$ such that $y\leq z$ and $x=y+z$ hold then $y$ lies in $D_S(x)$. 
+If there exist $y$ and $z$ in $S$ such that $y\leq z$ and $x=y+z$ hold then $y$ lies in~$D_S(x)$. 
 This implies $y=0$ and $z=x$. 
@@
 Let $S$ be a numerical semigroup and $x$ be an irreducible element of $S$. 
-Then for all $y\in \NN\setminus\{0\}$ we have
+Then for all~$y\in \NN\setminus\{0\}$ we have
 \[
@@
 \]
-We construct two sets as follow:
+We construct two sets as follow :
 \begin{align*}
@@
 a depth first search algorithm (see Algorithm~\ref{A:ExpRec}). This approach does not seem to
-have been used before.  In particular, M. Bras-Amor\'os and M.~Delgado use
+have been used before.  In particular, M.~Bras-Amor\'os and M.~Delgado use
 instead a breadth first search exploration.  The main advantage in our
@@
 
-Equivalently, we can use an iterative version which uses a stack:
+\noindent Equivalently, we can use an iterative version which uses a stack:
+
 \begin{algorithm}
@@
 \begin{proof}
-By Proposition~\ref{P:Res} $iii)$ we have  the relation $c(S)\leq 2g(S)$ and so the $S$-decomposition number of $c(S)$ occurs in $\delta_S$.
-Since $c(S)$ is equal to $\max(\NN\setminus S)$, Proposition~\ref{P:delta} implies 
+By Proposition~\ref{P:Res} $iii)$ we have  the relation $c(S)\leq 2g(S)$ and so the $S$-decomposition number of~$c(S)$ occurs in $\delta_S$.
+Since $c(S)$ is equal to $1+\max(\NN\setminus S)$ for $S\not=\NN$, Proposition~\ref{P:delta} implies 
 \[
@@
 
-For example the following Algorithm initializes a representation of the semigroup $\NN$ ready for an exploration of the tree $\mathcal{T}_G$, \ie, the tree of numerical semigroup upto genus $G$.
+For example the following Algorithm initializes a representation of the semigroup $\NN$ ready for an exploration of the tree $\mathcal{T}_G$ (the tree of numerical semigroup of genus at most $G$.)
 
@@
 \Function{Root}{\tt{G}}
-\State \tt{R.c} $\gets$ 1 \Comment{\tt{R} stands for $\NN$}
+\State \tt{R.c} $\gets$ 0 \Comment{\tt{R} stands for $\NN$}
 \State \tt{R.g} $\gets$ 0
 \State \tt{R.m} $\gets$ 1 
-\For{\tt{x} from $1$ to $3\,\tt{G}$}
+\For{\tt{x} from $0$ to $3\,\tt{G}$}
 	\State\tt{R.d[x]}$\gets1+\left\lfloor\frac{\tt{x}}2\right\rfloor$
@@
 We can now describe an algorithm that returns the representation of the semigroup $S^x$ from that of the semigroup $S$ where $x$ is an irreducible element of $S$ greater than $c(S)$.
-
-\newpage
 
@@
 Let us check the correctness of the algorithm.
-By construction $S^x$ is the semigroup $S\setminus\{x\}$.
-Thus the genus $S^x$ is $g(S)+1$, see Line~\tt{1}.
-Every integer of $I=[x+1,+\infty[$ lies in $S$ since $x$ is greater than $c(S)$, so the interval $I$ is included in $S^x$.
-As $x$ does not belong to $S$, the conductor of $S^x$ is $x+1$, see Line \tt{2}.
+By construction $S^x$ is the semigroup~$S\setminus\{x\}$.
+Thus the genus of $S^x$ is $g(S)+1$, see Line~\tt{3}.
+Every integer of $I=[x+1,+\infty[$ lies in $S$ since $x$ is greater than~$c(S)$, so the interval $I$ is included in $S^x$.
+As $x$ does not belong to $S$, the conductor of $S^x$ is $x+1$, see Line~\tt{2}.
 For the multiplicity of $S^x$ we have two cases. 
@@
 Assume now $x=m(S)$. 
-The relation $x(S)\geq c(S)$ and the characterisation of~$m(S)$ implies $x=m(S)=c(S)$.
+The relation $x(S)\geq c(S)$ and the characterization of~$m(S)$ implies $x=m(S)=c(S)$.
 Thus $S^x$ contains $m(S)+1$ which is $m(S^x)$.
-The initialisation of $m(S^x)$ is done by Lines \tt{4} to \tt{8}. 
-The correctness of the computation of $\delta_{S^x}$ (see Proposition~\ref{P:delta}) done from Line \tt{9} to Line \tt{15} is a direct consequence of~Proposition~\ref{P:Dandd}.
+The initialization of~$m(S^x)$ is done by Lines \tt{4} to \tt{8}. 
+The correctness of the computation of $\delta_{S^x}$ (see Proposition~\ref{P:delta} for the definition of $\delta_{S^x}$) done from Line \tt{9} to Line \tt{15} is a direct consequence of~Proposition~\ref{P:Dandd}.
 
@@
 
-\begin{algorithm}
+\begin{algorithm}[hb!]
 \caption{\small Returns an array containing the value of $n_g$ for $g\leq G$}\label{A:Count}
@@
 \Function{Count}{\tt{G}}
-\State \tt{n} $\gets [0,...,0]$ \Comment $\tt{n[g]}$ stands for $n_g$ and is initialised to $0$
+\State \tt{n} $\gets [0,...,0]$ \Comment $\tt{n[g]}$ stands for $n_g$ and is initialized to $0$
 \State Stack \tt{stack} \Comment{the empty stack}
@@
 
-For the time complexity, let us remark that Algorithm~\textsc{Son} is called for every semigroup of the tree $\mathcal{T}_G$ (the tree of semigroups of genus $\leq G$). 
+For the time complexity, let us remark that Algorithm~\textsc{Son} is called for every semigroup of the tree~$\mathcal{T}_G$ (the tree of semigroups of genus $\leq G$). 
 Since there are exactly $N=\sum_{g=0}^G n_g$ such semigroups, the time complexity of \textsc{Son} established in Proposition~\ref{P:Son} guarantees that the running time of \textsc{Count} is in~$O(\log(G)\times G\times N)$, as stated.
@@
 The first one is that reading the stack from the bottom to the top, the genus of  semigroup increases.  
-The second one is that, for all $g\in[0,G]$, every semigroup of genus $g$ in the stack has the same father.
-As the number of sons of a  semigroup $S$ is the number of $S$-irreducible elements in $[c(S),c(S)+m(S)-1]$,  a semigroup $S$ has at most $m(S)$ sons. 
+The second one is that, for all $g\in[0,G]$, every semigroup of genus~$g$ in the stack has the same father.
+As the number of sons of a  semigroup $S$ is the number of $S$-irreducible elements in the set $\{c(S),...,c(S)+m(S)-1\}$,  a semigroup $S$ has at most $m(S)$ sons. 
 By~Proposition~\ref{P:Res} $ii)$, this implies that a semigroup of genus $g$ as at most $g+1$ sons.
@@
 \[
-S=\sum_{g=0}^Gg=\frac{G(G+1)}2
+M=\sum_{g=0}^Gg=\frac{G(G+1)}2
 \]
@@
 Each entry of $\delta_S$ is the $S$-decomposition number of an integer smaller than $3G$ and hence
-requires $\log(G)$ bytes of memory space.
+requires $O(\log(G))$ bytes of memory space.
 Therefore the space complexity of $\delta_S$ is in~$O(\log(G)\times G)$, which implies that the space complexity of the \textsc{Count} algorithm is
 \[
-O(\log(G)\times G\times S)= O(\log(G)\times G^3).
+O(\log(G)\times G\times M)= O(\log(G)\times G^3).
 \]
@@
   kind of Duff's device.
-\item Carefull choice of data type (\verb|uint_fast8_t| for decomposition
+\item Careful choice of data type (\verb|uint_fast8_t| for decomposition
   number, vs \verb|uint_fast64_t| for all indexes).
@@
 memory byte $\texttt{T.d[y]}$.  Using these two cursors, Lines \tt{10} to
-\tt{14} of Algorithm~\textsc{Son} can be rewritten as follows :
-
-\begin{figure}
+\tt{14} of Algorithm~\ref{A:Son} can be rewritten as follows :
+
+\vspace{1em}
 \small
@@
 \end{algorithmic}
-\end{figure}
+\vspace{1em}
 
@@
 is 16~bytes aligned if it is a multiple of 16. \SSE memory access are much
-faster for aligned access.
+faster for aligned memory.
 
@@
 following \CPP code
+\vspace{1em}
 {
@@
 \end{alltt}}
+\vspace{1em}
+
 Here \tt{dec\_blocks} is a type for arrays of 16~bytes blocks whose size
@@
 the father which is to be removed in the children):
-{
-\small
+
+{\small
 \begin{alltt}
@@
 }
+
 The instruction \tt{load\_unaligned\_epi8} (specific to \SSE technology) loads
@@
 Those entries will be used to compute
-the entries $16i, \dots 16i+15$ of the children monoid. Since the removed
+the entries $16i, \dots 16i+15$ of the children semigroup. Since the removed
 generator~$x$ is not necessarily a multiple of $16$, the data are not aligned
@@
 (resp. \tt{one}) constants are initialized as 16 bytes equal to $0$
-(resp. $1$). The comparison \tt{(block != zero)} therfore returns a block
+(resp. $1$). The comparison \tt{(block != zero)} therefore returns a block
 which contains $0$ in the bytes corresponding the $0$ entries of block and
@@
 \begin{center}
-$\tt{dst} \gets \tt{dst} -
-\textbf{ if } \tt{block} \neq 0 \textbf{ then } 1 \textbf{ else } 0$
+$\tt{dst} \gets \tt{dst} - \textbf{ if } \tt{block} \neq 0 \textbf{ then } 1 \textbf{ else } 0$
 \end{center}
 
-which is equivalent to Lines \tt{10} to \tt{14} of Algorithm~\textsc{Son}.
+\noindent which is equivalent to Lines \tt{10} to \tt{14} of Algorithm~\ref{A:Son}.
 \medskip
@@
 particularly the scheduler, to decide during execution how to actually divide
-the work between processors.
+the work between cores.
 
@@
 to the following \CilkP code:
+\vspace{1em}
+
 {\small
@@
 }
+\vspace{1em}
+
 Note that in our version, we found that \tt{STACK\_BOUND} optimal value was
-around 10-12 for genus in the range $45\dots 67$ so that \tt{explore\_stack}
+around 10-12 for genus in the range~$45\dots 67$ so that \tt{explore\_stack}
 is used more than $99\%$ of the time. The \CilkP recursive function does
@@
 extremely fast \textsc{Son} algorithm. Indeed, its cost is comparable to the cost
-of copying a monoid. It is therefore crucial for performance to avoid any
+of copying a semigroup. It is therefore crucial for performance to avoid any
 extra cost. We list here various places where unnecessary cost can be avoided.
@@
 following sequences of operations:
+
+\vspace{1em}
 
@@
 
-If we use a stack of monoid, we can construct $\tt{S}^\tt{x}$ directly into
+\vspace{1em}
+
+If we use a stack of semigroup, we can construct $\tt{S}^\tt{x}$ directly into
 the stack memory but we have to copy the top of the stack to \tt{S}. Since the
@@
 being negligible. The trick is to use a level of indirection, replacing the
-stack by an array of monoid and an array of indexes pointing to the array of
-monoid. We keep the invariant that the array of indexes is a permutation of
-the array of monoid. Now instead of copying \tt{S} out of the stack, we keep
+stack by an array of semigroup and an array of indexes pointing to the array of
+semigroup. We keep the invariant that the array of indexes is a permutation of
+the array of semigroup. Now instead of copying \tt{S} out of the stack, we keep
 it on the stack, pushing the children in the second position by exchanging the
 indexes. Here is the relevant part of the code:
+
+\vspace{1em}
 
@@
 
+\vspace{1em}
+
 \subsubsection*{Avoiding dynamic allocation} Compared to the \textsc{Son}
@@
 the derecursived stack algorithm, we only allocate (on the system stack rather
-than on the heap) the stack of monoid. No further allocation are done.
+than on the heap) the stack of semigroup. No further allocation are done.
 
@@
 computation were running on the machine, we only used $32$ cores. The
-computation took $18$~days. The values of $n_g$ for $g\leq 67$ are :
+computation took $18$~days. The values of~$n_g$ for $g\leq 67$ are :
+
 \begin{center}
@@
 \]
-
-Since the work of M.~Bras-Amor{\'o}s, see \cite{BrasAmoros2008}, we yet know that all numerical semigroup satisfy of genus $g\leq 50$ satisfy the Wilf's conjecture.
+Since the work of M.~Bras-Amor{\'o}s, see \cite{BrasAmoros2008}, we yet know that all numerical semigroups of genus $g\leq 50$ satisfy the Wilf's conjecture.
 With our exploration algorithm we have proved that there is no counterexample to the Wilf's conjecture of genus $g\leq 60$.
@@
 the values of $n_g$ 
-for $g\leq G$ with $30\leq G \leq 40$. 
-on a machine equipped with  an \tt{Intel\texttrademark{} i5-3570K} CPU 
+for $g\leq G$ with $30\leq G \leq 40$
+on a machine equipped with  an \tt{Intel\texttrademark{} i5-3570K} CPU running at \texttt{3.4GHz}
 and \texttt{8GB} of memory.
-All algorithms are executed on only one thread. Algorithm \texttt{depth} is 
+All algorithms are executed on only one core. Algorithm \texttt{breadth} is 
 based on a breadth exploration of the tree while Algorithm \texttt{depth} 
@@
 
-\begin{figure}[h!]
+\begin{center}
 \begin{tabular}{|c|c|c|c|c|c|c|c|c|c|c|c|}
@@
 \end{tabular}
-\end{figure}
+\end{center}
 
@@
 The following table illustrates the impact of parallelization on the same 
-machine based on the \tt{Intel\texttrademark{} i5-3570K} CPU 
-running at \texttt{3.4GHz}. This CPU has two physical cores that are able 
+machine based on the \tt{Intel\texttrademark{} i5-3570K} CPU. This CPU has two physical cores that are able 
 to run four threads thanks to the \texttt{Hyper-Threading} 
@@
 vertical line delimites the use of \texttt{Hyper Threadind}). For reference,
-we put in the column called ``C++'' the computation time of the same
-program compiled without Cilk.
+we put in the column called \texttt{C++} the computation time of the same
+program compiled without \Cilk.
 \begin{figure}[h!]
@@
 \hline
- Threads & C++ & 1 & 2 & 4 & 8 & 12 & 16 & 20 & 24\\
+ Threads & \texttt{C++} & 1 & 2 & 4 & 8 & 12 & 16 & 20 & 24\\
  \hline
- Time (s) &      & 3709 & 1865 & 932.4 & 486.8 & 325.7 & 311.2 & 302.3 & 290.2\\
+ Time (s) & 3588.08 & 3709 & 1865 & 932.4 & 486.8 & 325.7 & 311.2 & 302.3 & 290.2\\
  \hline
@@
 % \end{figure}
+\vspace{2em}
 
@@
 suggested the two level of indirection stack optimization.
+
+\vspace{2em}
 

### 2014-11-03 Papier/article.tex
Hal version

@@
 \author{Jean Fromentin and  Florent Hivert}
-\date{2 October 2014}
+\date{3 November 2014}
 \keywords{semigroups, tree, algorithm, \texttt{SSE}, 
@@
 \begin{abstract}
-  In this paper we describe an algorithm visiting all the numerical semigroups
-  up to a given genus using a new representation of numerical semigroups. The
+  In this paper we describe an algorithm visiting all numerical semigroups
+  up to a given genus using a new representation. The
   interest of this algorithm is that it fits particularly well the
-  architecture of modern computer allowing very large optimizations : we obtain the number of numerical semigroups of genus $g\leq  67$ and we confirm the Wilf conjecture for $g\leq 60$.
+  architecture of modern computers allowing very large optimizations: we obtain the number of numerical semigroups of genus $g\leq  67$ and we confirm the Wilf conjecture for $g\leq 60$.
 
@@
 
-A \emph{numerical semigroup} $S$ is a subset of $\NN$ containing $0$, close under addition and of finite complement in $\NN$.  
+A \emph{numerical semigroup} $S$ is a subset of $\NN$ containing $0$, closed under addition and of finite complement in $\NN$.  
 For example the set 
@@
 The \emph{genus} of a numerical semigroup $S$, denoted by~$g(S)$, is the cardinality of~$\NN\setminus S$.
- For example the genus of $S_E$ is $6$,  the cardinality of $\{1,2,4,5,8,11\}$
+ For example the genus of $S_E$ is $6$,  the cardinality of $\{1,2,4,5,8,11\}$.
 
 For a given positive integer $g$, the number of numerical semigroups of genus $g$ is finite and is denoted by $n_g$. 
-In  J.A. Sloane's \emph{on-line encyclopedia of integer}~\cite{OEIS} we find the values of $n_g$ for $g\leq 52$. 
-These values have been obtain by M. Bras-Amor\'os (view \cite{BrasAmoros2008} for more details for $g\leq 50)$. 
-On his home page~\cite{Delgado}, M. Delgado  gives the value of $n_{55}$ without specifying the values of $n_{53}$ and~$n_{54}$.
+In  J.A. Sloane's \emph{on-line encyclopedia of integer sequences}~\cite{OEIS} we find the values of $n_g$ for $g\leq 52$. 
+These values have been obtained by M. Bras-Amor\'os (\cite{BrasAmoros2008} for more details for $g\leq 50)$. 
+On his home page~\cite{Delgado}, M. Delgado  gives the value of $n_{55}$. 
+The values of $n_{53}$ and $n_{54}$ seem to be so far available nowhere.
 
 M.~Bras-Amor\'os used a depth first search exploration of the tree of numerical semigroups $\mathcal{T}$ up to a given genus.
-This tree was introduced by J.C.~Rosales and al. in \cite{Rosales} and it is the subject of the Section~\ref{S:Tree}.
+This tree was introduced by J.C.~Rosales and al. in \cite{Rosales} and it is the subject of Section~\ref{S:Tree}.
 Starting with all the numerical semigroups of genus $49$ she obtained the number of numerical semigroups of genus $50$ in $18$ days on a pentium D runing at $3$GHz. 
-In the package~\tt{NumericalSgs} \cite{NumericalSgps} of \tt{GAP} \cite{GAP}, M. Delgado together with P.A.~Garcia-Sanchez and J. Morais used the same method of exploration.
+In the package~\tt{NumericalSgs} \cite{NumericalSgps} of \tt{GAP} \cite{GAP}, M.~Delgado together with P.A.~Garcia-Sanchez and J. Morais used the same method of exploration.
 
 Here we describe a new algorithm for the exploration of the tree of numerical
-semigroup $\mathcal{T}$ and achieve the computation of $n_g$ for $g\leq 67$.
+semigroups $\mathcal{T}$ and achieve the computation of $n_g$ for $g\leq 67$.
 The cornerstone of our method is a new combinatorial representation of
-numerical semigroup that is well suited for exploration of the
-tree~$\mathcal{T}$ and allow large code optimization essentially based on use of
+numerical semigroups that is well suited and allows large code optimization essentially based on the use of
 vectorial instructions and parallelization.
 The goal of the paper is twofold:
-first to present our new encoding of the semigroup and the associated
-algorithms, and second to present the optimization techniques which allows, for
+first to present our new encoding of numerical semigroups and the associated
+algorithms, and second to present the optimization techniques which allow, for
 those kinds of algorithms, to get speedups by factors of hundreds and even
-thousands. We claim that these technique are fairly general for those kinds of
+thousands. We claim that these techniques are fairly general for those kinds of
 algorithms. As a support for the claim, we applied it to an algorithm of
@@
 In Section~\ref{S:Tree} we describe the tree of numerical semigroups and give bounds for some parameters attached to a numerical semigroup.
-The decription of our new representation of numerical semigroup is done in the third section.
+The description of our new representation of numerical semigroups is done in the second section.
 In Section~\ref{S:Algo} we describe an algorithm based on the representation given in Section~\ref{S:DecNumber} and give its complexity. 
-Section~\ref{S:Opti} is more technical, and is devoted to the optimisation of the algorithm introduced in Section~\ref{S:Algo}.
-In the last section we emphase the results obtained using our algorithm.
+Section~\ref{S:Opti} is more technical, and is devoted to the optimization of the algorithm introduced in Section~\ref{S:Algo}.
+In the last section we emphasize the results obtained using our algorithm.
 
@@
 We start this section with definitions and properties of numerical semigroups that will be used in the sequel.
-For a more complete introduction, the reader can usefully consults the book \emph{Numerical Semigroups} of J.C.~Rosales and P.A.~Garc\'ia-S\'anchez \cite{BookNS} or the book \emph{The Diophantine Frobenius Problem} of J.L.~Ram\'irez~Alfons\'in \cite{BookDFP}.
+For a more complete introduction, the reader can usefully consult the book \emph{Numerical Semigroups} of J.C.~Rosales and P.A.~Garc\'ia-S\'anchez \cite{BookNS} or the book \emph{The Diophantine Frobenius Problem} of J.L.~Ram\'irez~Alfons\'in \cite{BookDFP}.
 
@@
 
-By definition a numerical semigroup is an infinite object and we need a finite description of such a object. 
-That is the role of generating sets.
+By definition a numerical semigroup is an infinite object and we need a finite description of such an object. 
+That is provided by generating sets.
 
@@
 \label{E:GNSG}
-S_E=\{0,3,6,7,9,10\}\cup[12,+\infty[=\left<3,7\right>
+S_E=\{0,3,6,7,9,10\}\cup[12,+\infty[=\left<3,7\right>.
 \end{equation}
 
-A non-zero element $x$ of a numerical semigroup $S$ is said to be \emph{irreducible} if it cannot be expressed as a sum of two non-zeros elements of $S$.  
+A non-zero element $x$ of a numerical semigroup $S$ is said to be \emph{irreducible} if it cannot be expressed as a sum of two non-zero elements of $S$.  
 We denote by $\Irr(S)$ the set of all irreducible elements of $S$.
@@
 \begin{proof}
-$i)$ Let $x$ ba a positive integer satisfying $x \geq c(S)+m(S)$. 
+$i)$ Let $x$ be a positive integer satisfying $x \geq c(S)+m(S)$. 
 As the integers $x$ and $x-m(S)$ are greater than $c(S)$, they belong to $S$.
-Since $x$ equals $(x-m(S))+m(S)$ with $x-m(S)\not=0$ and $m(S)\not=0$, it could not be irreducible.
+Since $x$ equals $(x-m(S))+m(S)$ with $x-m(S)\not=0$ and $m(S)\not=0$, it cannot be irreducible.
 
@@
 The integer $y=c(S)-1-x$ lies in $\{0,...,c(S)-1\}$. 
-Moreover $y$ is not in $S$, for otherwise  we write $c(S)-1=y+x$ where $x,y$ are elements of $S$ implying $c(S)-1\in S$. 
+Moreover $y$ is not in $S$, for otherwise  we write $c(S)-1=y+x$ where $x,y$ are elements of $S$ implying $c(S)-1$ belongs to $S$. 
 In contradiction with the definition of conductor. 
@@
 \]
-that send $S\{0,...,c(S)-1\}$ into $\{0,...,c(S)-1\}\setminus S$ which is  of cardinality~$g(S)$. 
+that sends $S\{0,...,c(S)-1\}$ into $\{0,...,c(S)-1\}\setminus S$ which is  of cardinality~$g(S)$. 
 Let us denote by $k$ the cardinality  of $S\cap\{0,...,c(S)-1\}$. 
@@
 A consequence of Proposition~ \ref{P:Res}~$i)$ is that $\Irr(S)$ is finite and its cardinality is at most $c(S)+m(S)-1$.
-More precisely, using Ap\'ery set, we can prove that  the cardinality of $\Irr(S)$ is at most $m(S)$. See Section~2 of Chapter I of~\cite{BookNS} for more details.
+Moreover, the cardinality of $\Irr(S)$ is at most $m(S)$ since any two distinct elements of $\Irr(S)$ cannot be congruent modulo $m(S)$. See Section~2 of Chapter I of~\cite{BookNS} for more details.
 
@@
 
-Proposition~\ref{P:Sx} implies that every semigroup $S$ of genus $g$ can be obtained from a semigroup $S$ by removing a generator $x$ of $S$ that is greater than $c(S)$.
+Proposition~\ref{P:Sx} implies that every semigroup $S$ of genus $g$ can be obtained from a semigroup $S$ by removing a generator $x$ of $S$ that is greater than or equal to $c(S)$.
 Hence the relation~$S'=S^x$ holds.
@@
 The root of the tree is the unique semigroup of genus $0$, \ie, $\left<1\right>$ that is equal to $\NN$. 
-If~$S$ is a semigroup in the tree,  the sons of $S$ are exactly the semigroup~$S^x$ where $x$ belongs to $\Irr(S)\cap[c(s),+\infty[$. 
+If~$S$ is a semigroup in the tree,  the sons of $S$ are exactly the semigroups~$S^x$ where $x$ belongs to $\Irr(S)\cap[c(s),+\infty[$. 
 By convention, when depicting the tree, the numerical semigroup $S^x$ is in the left of $S^y$ if $x$ is smaller than~$y$. 
 With this construction, a semigroup $S$ has depth $g$ in $\mathcal{T}$ if and only if its genus is~$g$, see Figure~\ref{F:Tree}.
-We denote by $\mathcal{T}_{g}$ the subtree of~$\mathcal{T}$ resticted to all semigroup of genus $\leq g$.
+We denote by $\mathcal{T}_{g}$ the subtree of~$\mathcal{T}$ restricted to all semigroups of genus $\leq g$.
 
@@
 \caption{The first four layers of the tree $\mathcal{T}$ of numerical semigroups, corresponding to~$\mathcal{T}_4$.
-A generator of a semigroup is in gray if is not greater than $c(S)$.
-An edge between a semigroup $S$ and its son $S'$ is labelled by  $x$ is $S'$ is obtained from $S$ by removing~$x$, that is $S'=S^x$.}
+A generator of a semigroup is it in gray if is not greater than $c(S)$.
+An edge between a semigroup $S$ and its son $S'$ is labelled by  $x$ if $S'$ is obtained from $S$ by removing~$x$, that is if $S'=S^x$ holds.}
 \label{F:Tree}
@@
 
-As the reader can check, the main difficulty to characterize the son of a semigroup is to determine its irreducible elements.
- In \cite{NumericalSgps}, the semigroup are given by their Ap\'ery set and then the difficulty is to describe the Ap\'ery set of $S^x$  from this of $S$. 
- This approach is elegant but not sufficiently \emph{basic} for our optimisations.
+As the reader can check, the main difficulty to characterize the sons of a semigroup is to determine its irreducible elements.
+ In \cite{NumericalSgps}, the semigroups are given by their Ap\'ery sets and then the difficulty is to describe the Ap\'ery set of $S^x$  from that of $S$. 
+ This approach is elegant but not sufficiently \emph{basic} for our optimizations.
 
@@
  Hence~$D_S(x)$ describes how~$x$ can be decomposed as sums of two elements of $S$.
- This is why we decide to call $d_S(x)$ the $S$-decomposition number of $x$.
+ This justifies the name given to the function $d_S$.
 
@@
 Reconsider the semigroup $S_E$ given at \eqref{E:NSG}.
-The integer $14$ admits three decompositions as sums of two elements of $S$, namely $14=0+14$ and $14=7+7$. 
-Thus the set $D_{S_E}(14)$ is equal to $\{0,7\}$ and $d_{S_E}$ equals $2$.
+The integer $14$ admits two decompositions as sums of two elements of $S$, namely $14=0+14$ and $14=7+7$. 
+Thus the set $D_{S_E}(14)$ is equal to $\{0,7\}$ and $d_{S_E}=2$ holds.
 \end{exam}
@@
 Hence, the set $D_S(x)$ is equal to~$\{0\}$ and we have $d_S(x)=1$. 
-Conversely, let $x$ such that $d_S(x)$ is $1$. 
+Conversely, let $x$ be such that $d_S(x)$ is $1$. 
 By~$i)$ the  integer $x$ must be in $S$. 
@@
 
-We note  that $0$ is never irreducible despite the fact $d_S(0)$ is $1$ for all numerical semigroup $S$.
-
-We now explain how to compute the $S$-decomposition numbers function of a numerical semigroup  from this of its father.
+We note  that $0$ is never irreducible despite the fact $d_S(0)$ is $1$ for all numerical semigroups $S$.
+
+We now explain how to compute the $S$-decomposition numbers function of a numerical semigroup  from that of its father.
 
@@
 \{y-x\}&\text{if $y-x\in S^x$ and $y\leq 2x$}\\
-\emptyset&\text{otherwise}
+\emptyset&\text{otherwise}.
 \end{cases}
@@
 \{y-x\}&\text{if $y-x\in S$ and $y< 2x$}\\
-\emptyset&\text{otherwise}
+\emptyset&\text{otherwise}.
 \end{cases}
@@
 Then if $y-x\in S$ holds, we have $\card(E)+\card(F)=1$ and so $d_{S}(y)$ is equal to $d_{S^x}(y)+1$.
-If $y-x$ not belong to $S$, then $E$ and $F$ are empty implying $d_{S}(y)=d_{S^x}(y)$.
+If $y-x$ does not belong to $S$, then $E$ and $F$ are empty implying $d_{S}(y)=d_{S^x}(y)$.
 As the condition $y-x\in S$ is equivalent to $y\geq x$ together with $d_S(y-x)>0$, by Proposition~\ref{P:Dandd} the result holds.
@@
 In Algorithm~\ref{A:ExpRec} we do not specify how to compute $c(S)$, $g(S)$ and $m(S)$  from $S$ neither how to test if an integer is irreducible. 
-It also miss  the characterisation of $S^x$ from $S$.
+It also misses  the characterization of $S^x$ from $S$.
 These items depend heavily of the representation of $S$.
 Our choice is to use the $S$-decomposition numbers function. 
-The first task is to use a finite set of such numbers to characterise the whole semigroup.
+The first task is to use a finite set of such numbers to characterize the whole semigroup.
 
@@
 Let $G$ be an integer and $S$ be a numerical semigroup of genus $g\leq G$.
-Then $S$ is entirely described by $\delta_S=(d_S(0),...,d_S(3\,G))$. More precisely we can obtain $c(S), g(S), m(S)$ and $\Irr(S)$ from $\delta_S$.
+Then $S$ is entirely described by the vector $\delta_S=(d_S(0),...,d_S(3\,G)) \in \NN^{3G+1}$. More precisely we can obtain $c(S), g(S), m(S)$ and $\Irr(S)$ from $\delta_S$.
 \end{prop}
@@
 \]
-Since, by Proposition~\ref{P:Res} $i)$,  all irreducible elements are smaller than $c(S)+m(S)-1$, which is itself smaller than $3G$ from the relations $ii)$ and $iii)$ of Proposition~\ref{P:Res}.
-So Proposition~\ref{P:Dandd} gives
+By Proposition~\ref{P:Res},  all irreducible elements are smaller than $c(S)+m(S)-1$, which is itself smaller than~$3G$.
+Hence, Proposition~\ref{P:Dandd} gives
 \[
@@
 We represent a numerical semigroup $S$ of genus $g\leq G$ by $(c(S),g(S),c(S),\delta_S)$. 
-In an algorithmic context, if the variable \texttt{S} stands for a numerical semigroup  we use:
+In an algorithmic context, if the variable \texttt{S} stands for a numerical semigroups  we use:
 
@@
 Since the stack is filled  with a depth first search algorithm, it has two properties.
-The first one is that reading the stack from the bottom to the top, the genus of  semigroup increases.  
+The first one is that reading the stack from the bottom to the top, the genus increases.  
 The second one is that, for all $g\in[0,G]$, every semigroup of genus~$g$ in the stack has the same father.
@@
 By~Proposition~\ref{P:Res} $ii)$, this implies that a semigroup of genus $g$ as at most $g+1$ sons.
-Therefore the stack contains at most $g+1$ semigroup of genus $g+1$ for $g\leq G$. 
+Therefore the stack contains at most $g+1$ semigroups of genus $g+1$ for $g\leq G$. 
 So the size of the stack is bounded~by 
 \[
-M=\sum_{g=0}^Gg=\frac{G(G+1)}2
+M=\sum_{g=0}^Gg=\frac{G(G+1)}2.
 \]
@@
 
-Though there are asymptotically faster algorithms than the one presented here,
+Even though there are asymptotically faster algorithms than the one presented here,
 thank to careful optimizations, we were able to compute $n_g$ for much larger
-genus than before. This is due to the fact that our algorithm is particularly
+genuses than before. This is due to the fact that our algorithm is particularly
 well suited for the current processor architecture. In particular, it allows
-to use parallelism at various scale (parallel branch exploration,
+to use parallelism at various scales (parallel branch exploration,
 vectorization)...
 
-To get the greatest speed from modern processor, we used several optimization
+To get the greatest speed from modern processors, we used several optimization
 tricks, which we will elaborate in the following section:
@@
 
+The source code of our algorithm is available here \cite{code}.
 \subsection{Vectorization}
@@
 Assume for example that we want to construct the tree~$\mathcal{T}_{100}$
-of all numerical semigroup of genus smaller than $100$.  In this case, the
-representation of numerical semigroup given in Section~\ref{S:DecNumber} uses
+of all numerical semigroups of genus smaller than $100$.  In this case, the
+representation of numerical semigroups given in Section~\ref{S:DecNumber} uses
 decomposition numbers of integers smaller than~$300$.  By Lemma~\ref{L:RepN},
@@
 memory byte $\texttt{T.d[y]}$.  Using these two cursors, Lines \tt{10} to
-\tt{14} of Algorithm~\ref{A:Son} can be rewritten as follows :
+\tt{14} of Algorithm~\ref{A:Son} can be rewritten as follows:
 
@@
 \MMX, \SSE, etc.  The acronym \SIMD~\cite{WikipediaSIMD} stands for Single Operation Multiple Data.  
-We used \SSEV~\cite{WikipediaSSE, IntelSSE} technology as it allows for the largest speedup\footnote{A much more speedup can be certainly obtained using \texttt{AVX2} technology \cite{AVX2}.
-However, at this time, we can't access a performent computer with this set of instructions.}.
+We used \SSEV~\cite{WikipediaSSE, IntelSSE} technology as it allows for the largest speedup\footnote{A much greater  speedup can be certainly obtained using \texttt{AVX2} technology \cite{AVX2}.
+However, at this time, we can't access a performant computer with this set of instructions.}.
 This need to respect some constraints in the memory
@@
 $3G$ rounded up to a multiple of 16). The instruction \tt{dst[i] = src[i]}
-actually copy a full 16~bytes block.
+actually copies a full 16~bytes block.
 
@@
 Those entries will be used to compute
-the entries $16i, \dots 16i+15$ of the children semigroup. Since the removed
+the entries $16i, \dots 16i+15$ of the children semigroups. Since the removed
 generator~$x$ is not necessarily a multiple of $16$, the data are not aligned
@@
 
-%{\small
-%\begin{algorithmic}
-%\State \textbf{Jump to case } $\tt{start}+1$
-%\State \textbf{case 1:} \tt{block = ...; dst[1] -= ((block != zero) \& one)}
-%\State \textbf{case 2:} \tt{block = ...; dst[2] -= ((block != zero) \& one)}
-%\State \textbf{[...]}
-%\State \textbf{case NBLOCK-1:} \tt{...}
-%\end{algorithmic}}
-
-% The MMX technology permits to work on $8$ bytes in parallel while the SSE
-% works in $16$ bytes.  As the rest of the CPU works on $8$ bytes, the SSE
-% technology needs some constraint on memory access than cannot be fulfilled in
-% our algorithm.
-
-% This motivate our choice to use the MMX technology. 
-% More precisely, we use three commands: \tt{pcmpeqb}, \tt{pandn} and \tt{psubb}.
-%  These commands work on two arrays of $8$ bytes, called~\tt{s} and \tt{d} here. 
-%  In each case the array \tt{d} is modified. 
-%  We denote bytes by $\{\tt{0},\tt{1}\}$-words of length $8$.
-%  After a call to \tt{pcmpeqb(s,d)} the $i$th entry of $d$ contains \tt{11111111} if $\tt{s}[i]=\tt{d}[i]$ holds and \tt{00000000} otherwise. 
-%  The command \tt{pand(s,d)}  store in $\tt{d}[i]$  the value of the byte-logic operation \tt{s}[i] \tt{and} (\tt{not} \tt{d}[i]).
-%   When \tt{pand(s,d)} is called the new value of $\tt{d}[i]$ is $\tt{s}[i]-\tt{d}[i]$.  
-%   Glueing all the pieces together we obtain the following version of the \textbf{for} loop of Algorithm~\textsc{Son} :
-
-% \begin{algorithmic}[1]
-% \State $\tt{src}\gets \text{address}(\tt{S.d[0]})$
-% \State $\tt{dst}\gets \text{address}(\tt{T.d[x]})$
-% \State $\tt{i}\gets 0$
-% \While{$\tt{i}\leq \tt{3G}-\tt{x}$}
-% \State $\tt{t}\gets [\tt{00000000},...,\tt{00000000}]$ \Comment{$8$ bytes equal to \tt{00000000}}
-% \State \tt{pcmpeqb(\tt{src},t)}
-% \State \tt{pandn([00000001,...,00000001],t)}
-% \State \tt{psub(\tt{dst},t)}
-% \State $\tt{dst}\gets \tt{t}$ \Comment{copy the array $[\tt{t}[0],...,\tt{t}[7]]$ to $[\tt{d}[0],...,\tt{d}[7]]$}
-% \State \textbf{increase} \tt{src},\tt{dst},\tt{i} \textbf{by} $8$
-% \EndWhile
-% \end{algorithmic}
-
-% Let us explain how this works in more details. 
-% Let $i$ be an integer in $\{0,...,7\}$.  
-% After Line~\tt{5}, the value of $\tt{t}[i]$ is \tt{0000000}. 
-% After Line~\tt{6}, we have 
-% \[
-% \tt{t}[i]=\begin{cases}
-% \tt{11111111}&\text{if $\tt{src}[i]=\tt{00000000}$ (corresponding to $0$)}\\
-% \tt{00000000}&\text{otherwise}
-% \end{cases}
-% \]
-% Line \tt{7} performs a logical \tt{and} between $\tt{00000001}$ and $\tt{not t}[i]$. Hence the result byte is \tt{00000001} if the last byte of \tt{t}[i] is $0$ and \tt{00000000} otherwise: 
-% \[
-% \tt{t}[i]=\begin{cases}
-% \tt{00000000}&\text{if $\tt{src}[i]=\tt{00000000}$ (corresponding to $0$)}\\
-% \tt{00000001}&\text{otherwise}
-% \end{cases}
-% \]
-%  Line \tt{8} subtracts \tt{t}[i] (which is equal to $0$ or $1$) from $\tt{dst}[i]$ according to Proposition~\ref{P:Sx}. 
-% Finally we shift the cursor \tt{src} and \tt{dst} by eight cases.
-
 \subsection{Parallel tree exploration using \CilkP}
@@
 Our second optimization is to use parallelism on exploration of the tree.
-Today, CPU of personal computer have several cores (2, 4 or more).  The given
+Today, CPU of personal computers have several cores (2, 4 or more).  The given
 version of our exploration algorithm uses a single core and so a fraction only
@@
 can be explored in parallel by different cores of the computer. The tricky
-part is to ensure that each cores are busy, giving a new branch when a core is
+part is to ensure that all cores are busy, giving a new branch when a core is
 done with a former one. Fortunately there is a technology called
 \CilkP~\cite{CilkIntel, CilkRefman} which is particularly well suited for
-those kind of problems. For our computation, we used the free version which is
+those kinds of problems. For our computation, we used the free version which is
 integrated in the latest version of the GNU~C compiler~\cite{GCCcilk}.
@@
   safely operate in parallel with the remaining code of the current
-  function. Note that the scheduler is not obligated to run this procedure in
+  function. Note that the scheduler is not obliged to run this procedure in
   parallel; the keyword merely alerts the scheduler that it can do so.
@@
 As a consequence, to get a parallel version of the recursive
-Algorithm~\ref{A:ExpRec}, you only need to modify it as
+Algorithm~\ref{A:ExpRec}, one only needs to modify it as
 
@@
 explored in parallel.
-The thing are actually only a little bit more complicated. First we have to
+Things are actually only a little bit more complicated. First we have to
 gather the results of the exploration. If we simply write
@@
  is possible: Read1 / Read2 / Add1 / Add2 / Store1 / Store2. Then the two
-cores performs the same modification resulting in incrementing the variable
+cores perform the same modification resulting in incrementing the variable
 only once. This is called a data race and leads to nondeterministic wrong
-result. To cope with those kinds of synchronization problems, \CilkP provide the
+results. To cope with those kinds of synchronization problems, \CilkP provides the
 notion of \emph{reducer} which are variables local to each thread which are
@@
 
-Note that in our version, we found that \tt{STACK\_BOUND} optimal value was
-around 10-12 for genus in the range~$45\dots 67$ so that \tt{explore\_stack}
+Note that in our version, we found that the \tt{STACK\_BOUND} optimal value was
+around 10 to 12 for genus in the range~$45\dots 67$ so that \tt{explore\_stack}
 is used more than $99\%$ of the time. The \CilkP recursive function does
-actually very little work but ensure that the work is balanced between the
+actually very little work but ensures that the work is balanced between the
 different cores.
@@
 
-Using vectorization and loop enrolling as described previously leads to an
+Using vectorization and loop unrolling as described previously leads to an
 extremely fast \textsc{Son} algorithm. Indeed, its cost is comparable to the cost
@@
 \subsubsection*{Avoiding all unnecessary copy} We also used a trick to avoid
-copying from and to the top of the stack. Indeed, then main loop perform the
+copying from and to the top of the stack. Indeed, then main loop performs the
 following sequences of operations:
@@
 
-\vspace{1em}
+\vspace{0.5em}
 
 If we use a stack of semigroup, we can construct $\tt{S}^\tt{x}$ directly into
-the stack memory but we have to copy the top of the stack to \tt{S}. Since the
-average number of children is the golden ratio, this single copy is far from
+the stack memory but we have to copy the top of the stack to \tt{S}. 
+In \cite{Zhai}, A.Zhai establishes that the limit of the quotient   $\frac{n_g}{n_{g-1}}$, when $g$ go to infinity, is the golden ratio $\phi\approx 1.618$. Therefore this single copy is far from
 being negligible. The trick is to use a level of indirection, replacing the
-stack by an array of semigroup and an array of indexes pointing to the array of
-semigroup. We keep the invariant that the array of indexes is a permutation of
-the array of semigroup. Now instead of copying \tt{S} out of the stack, we keep
+stack by an array of semigroups $A$ and an array of indexes $I$ pointing to the array of
+semigroups. The array $I$ can be viewed as a permutation of
+the array~$A$. Now instead of copying \tt{S} out of the stack, we keep
 it on the stack, pushing the children in the second position by exchanging the
-indexes. Here is the relevant part of the code:
-
-\vspace{1em}
+indexes in $I$. Here is the relevant part of the code:
+
+\vspace{0.5em}
 
@@
 \subsubsection*{Avoiding dynamic allocation} Compared to the \textsc{Son}
-algorithm, dynamic allocation cost orders of magnitude more. Therefore, during
+algorithm, dynamic allocation costs orders of magnitude more. Therefore, during
 the derecursived stack algorithm, we only allocate (on the system stack rather
-than on the heap) the stack of semigroup. No further allocation are done.
+than on the heap) the stack of semigroups. No further allocation are done.
 
@@
 indexes (namely \verb|uint_fast64_t|) to avoid conversion and sign extension
-when computing addresses of indexed element. This single standard trick save
+when computing addresses of indexed elements. This single standard trick save
 10\% of speed.
@@
 Running the \Cilk version of our optimized algorithm we have explored the tree
-of numerical semigroup upto genus $67$. The computation were done one a shared
+of numerical semigroups up to genus $67$. The computation were done one a shared
 $64$ core AMD Opteron\texttrademark{} Processor 6276. As other heavy
@@
 \hline
-0 & 1 & 23 & 170963 & 46 & 14463633648\\
-1 & 1 & 24 & 282828 & 47 & 23527845502\\
-2 & 2 & 25 & 467224 & 48 & 38260496374\\
-3 & 4 & 26 & 770832 & 49 & 62200036752\\
-4 & 7 & 27 & 1270267 & 50 & 101090300128\\
-5 & 12 & 28 & 2091030 & 51 & 164253200784\\
-6 & 23 & 29 & 3437839 & 52 & 266815155103\\
-7 & 39 & 30 & 5646773 & 53 & 433317458741\\
-8 & 67 & 31 & 9266788 & 54 & 703569992121\\
-9 & 118 & 32 & 15195070 & 55 & 1142140736859\\
-10 & 204 & 33 & 24896206 & 56 & 1853737832107\\
-11 & 343 & 34 & 40761087 & 57 & 3008140981820\\
-12 & 592 & 35 & 66687201 & 58 & 4880606790010\\
-13 & 1001 & 36 & 109032500 & 59 & 7917344087695\\
-14 & 1693 & 37 & 178158289 & 60 & 12841603251351\\
-15 & 2857 & 38 & 290939807 & 61 & 20825558002053\\
-16 & 4806 & 39 & 474851445 & 62 & 33768763536686\\
-17 & 8045 & 40 & 774614284 & 63 & 54749244915730\\
-18 & 13467 & 41 & 1262992840 & 64 & 88754191073328\\
-19 & 22464 & 42 & 2058356522 & 65 & 143863484925550\\
-20 & 37396 & 43 & 3353191846 & 66 & 233166577125714\\
-21 & 62194 & 44 & 5460401576 & 67 & 377866907506273\\
-22 & 103246 & 45 & 8888486816 & \\
+0 & 1 & 23 & 170\,963 & 46 & 14\,463\,633\,648\\
+1 & 1 & 24 & 282\,828 & 47 & 23\,527\,845\,502\\
+2 & 2 & 25 & 467\,224 & 48 & 38\,260\,496\,374\\
+3 & 4 & 26 & 770\,832 & 49 & 62\,200\,036\,752\\
+4 & 7 & 27 & 1\,270\,267 & 50 & 101\,090\,300\,128\\
+5 & 12 & 28 & 2\,091\,030 & 51 & 164\,253\,200\,784\\
+6 & 23 & 29 & 3\,437\,839 & 52 & 266\,815\,155\,103\\
+7 & 39 & 30 & 5\,646\,773 & 53 & 433\,317\,458\,741\\
+8 & 67 & 31 & 9\,266\,788 & 54 & 703\,569\,992\,121\\
+9 & 118 & 32 & 15\,195\,070 & 55 & 1\,142\,140\,736\,859\\
+10 & 204 & 33 & 24\,896\,206 & 56 & 1\,853\,737\,832\,107\\
+11 & 343 & 34 & 40\,761\,087 & 57 & 3\,008\,140\,981\,820\\
+12 & 592 & 35 & 66\,687\,201 & 58 & 4\,880\,606\,790\,010\\
+13 & 1\,001 & 36 & 109\,032\,500 & 59 & 7\,917\,344\,087\,695\\
+14 & 1\,693 & 37 & 178\,158\,289 & 60 & 12\,841\,603\,251\,351\\
+15 & 2\,857 & 38 & 290\,939\,807 & 61 & 20\,825\,558\,002\,053\\
+16 & 4\,806 & 39 & 474\,851\,445 & 62 & 33\,768\,763\,536\,686\\
+17 & 8\,045 & 40 & 774\,614\,284 & 63 & 54\,749\,244\,915\,730\\
+18 & 13\,467 & 41 & 1\,262\,992\,840 & 64 & 88\,754\,191\,073\,328\\
+19 & 22\,464 & 42 & 2\,058\,356\,522 & 65 & 143\,863\,484\,925\,550\\
+20 & 37\,396 & 43 & 3\,353\,191\,846 & 66 & 233\,166\,577\,125\,714\\
+21 & 62\,194 & 44 & 5\,460\,401\,576 & 67 & 377\,866\,907\,506\,273\\
+22 & 103\,246 & 45 & 8\,888\,486\,816 & &\\
 \hline
@@
 
-In \cite{Zhai}, A.Zhai establishes that the limit of the quotient   $\frac{n_g}{n_{g-1}}$, when $g$ go to infinity, is the golden ratio $\phi\approx 1.618$.
-As the reader can check the convergence is very slow : the quotient $n_{67}/n_{66}$ is $\approx 1.62$.
+As the reader can check the convergence of sequence $\frac{n_g}{n_{g-1}}$ established by A.~Zhai in \cite{Zhai} is very slow: $\frac{n_{67}}{n_{66}}\approx 1.62$.
 
@@
 In the paper \cite{Wilf} of 1978, H.S. Wilf conjectured that all numerical semigroup $S$ satisfy the relation 
-
-\[
-\card(\Irr(S))\geq \frac{c(S)}{c(S)-g(S)}
-\]
-Since the work of M.~Bras-Amor{\'o}s, see \cite{BrasAmoros2008}, we yet know that all numerical semigroups of genus $g\leq 50$ satisfy the Wilf's conjecture.
-With our exploration algorithm we have proved that there is no counterexample to the Wilf's conjecture of genus $g\leq 60$.
+\[
+\card(\Irr(S))\geq \frac{c(S)}{c(S)-g(S)}.
+\]
+Since the work of M.~Bras-Amor{\'o}s, see \cite{BrasAmoros2008}, we yet know that all numerical semigroups of genus $g\leq 50$ satisfy Wilf's conjecture.
+With our exploration algorithm we have proved that there is no counterexample to Wilf's conjecture up to genus $60$.
 
@@
 
-In this section we summarize the timing improvement through the different 
-optimization of our algorithm.
-
-The following table show the time needed by the algorithm for computing 
+In this section we summarize the timing improvements through the different 
+optimizations of our algorithm.
+
+The following table shows the time needed by the algorithm for computing 
 the values of $n_g$ 
@@
 use a depth exploration. Algorithm \texttt{depth+$\delta$} is a refinement of 
-\texttt{depth} based on the the $S$-decomposition function. Algorithm 
+\texttt{depth} based on the $S$-decomposition function. Algorithm 
 \texttt{$\delta$+sse} in an optimization of \texttt{depth+$\delta$} using the 
@@
 technology~\cite{WikipediaHT}. However, when more 
-than two threads are running, the computation engine are shared between two 
+than two threads are running, the computation engines are shared between two 
 threads.
@@
 
+\vspace{-0.2em}
+
 The time of the one threaded algorithm  must be compared with the 
@@
 
-It should be notice that the \texttt{TurboBoost} technology \cite{WikipediaTB} 
+It should be noticed that the \texttt{TurboBoost} technology \cite{WikipediaTB} 
 is present on the CPU. Therefore the clock of the CPU is a bit higher when the 
-number of thread 
-is smaller. Therefore the gain of using more threads is a bit over than one 
-suggested by the table.
-
-Finally we test our algorithm on a machine holding two 
+number of threads 
+is smaller. Therefore the gain of using more threads is a bit over than the one 
+suggestested by the table.
+
+Finally we tested our algorithm on a machine holding two 
 \tt{Intel\texttrademark{} Xeon\texttrademark{} X5650} CPU running 
 at \texttt{2.67GHz}.
-Each of those CPU have 6 physical cores that are able to run 12 threads : the 
+Each of those CPU has 6 physical cores that are able to run 12 threads : the 
 machine has also 12 physical cores.  
 For more than $12$ threads, the speedup should be much less when
-adding more core. The following table resumes the time needed by the algorithm 
-to explore $\mathcal{T}_{50}$ on different number of threads (the double 
-vertical line delimites the use of \texttt{Hyper Threadind}). For reference,
+adding more cores. The following table resumes the time needed by the algorithm 
+to explore $\mathcal{T}_{50}$ on different numbers of threads (the double 
+vertical line delimite the use of \texttt{Hyper Threading}). For reference,
 we put in the column called \texttt{C++} the computation time of the same
@@
 
-% 
-% We can see that this is indeed happening under the
-% 
-% horizontal bar in the following table.
-%   \begin{figure}[h!]
-% \begin{tabular}{|c|c|c|c|c|c|}
-% \hline
-% Threads & 30 & 35 & 40 & 45 & 50 \\
-% \hline
-% 1 &  0.33 & 2.50 & 26.1 & 307 & 3709 \\
-% 2 &  0.17 & 1.25 & 13.05 & 153.5 & 1865\\
-% 3 &  0.11 & 0.71 & 8.57 & 102.6 & 1243 \\
-% 4 &  0.10 & 0.55 & 6.45 &  77.21 & 932.4 \\
-% 8 &  0.10 & 0.36 & 3.55 &  40.22 & 486.8 \\
-% 12 &  0.10 & 0.36 & 2.61 & 27.34 & 325.7 \\\hline
-% 16 &  0.10 & 0.37 & 2.50 & 25.96 & 311.2 \\
-% 20 &  0.10 & 0.37 & 2.69 & 26.04 & 302.3 \\
-% 24 &  0.12 & 0.51 & 3.18 & 25.27 & 290.2 \\
-% \hline
-% \end{tabular}
-% \end{figure}
+\textbf{Acknowledgment.} We would like to thank Shalom Eliahou for introducing us to the tree of numerical semigroups and for interesting
+discussions on the subject. We also would like to thanks Nathan Cohen who
+suggested the two levels of indirection stack optimization.
+
+\bibliographystyle{plain} 
+\bibliography{biblio}
 \vspace{2em}
 
-\textbf{Acknowledgment.} We would like to thank Shalom Eliahou for his
-presentation of the tree of numerical semigroups and for the interesting
-discussions on the subject. We also would like to thanks Nathan Cohen who
-suggested the two level of indirection stack optimization.
+\noindent \textbf{Jean Fromentin}
+
+ULCO, LMPA J.~Liouville, B.P. 699, F-62228 Calais, France
+
+CNRS, FR 2956, France
+
+\url{fromentin@lmpa.univ-littoral.fr}
 
@@
 
-%\begin{figure}[b]
-%\begin{tikzpicture}[scale=1.3]
-%\begin{semilogyaxis}[xlabel=\small genus,ylabel=time,legend pos=north west,legend cell align=left]
-%\addplot[color=blue,mark=x] coordinates {(20,0.02)(21,0.04) (22,0.07)(23,0.12)(24,0.2)(25,0.35)(26,0.61)(27,1.04) (28,1.76)(29,2.96)(30,4.96)(31,8.26)(32,13.77)(33,22.95)(34,38.22)(35,1250.88)};
-%\addplot[color=black,mark=x] coordinates {(20,0.02)(21,0.04) (22,0.06)(23,0.09)(24,0.16)(25,0.26)(26,0.45)(27,0.72)(28,1.25)(29,2.1)(30,3.39)(31,5.85)(32,9.25)(33,16.34)(34,27.23)(35,45.39)(36,75.58)(37,125.59)(38,203.56)(39,346.02)(40,557.51)};
-%\addplot[color=red,mark=x] coordinates {(20,0.02)(21,0.02)(22,0.04)(23,0.06)(24,0.1)(25,0.18)(26,0.29)(27,0.48)(28,0.81)(29,1.34)(30,2.25)(31,3.73)(32,6.2)(33,10.35)(34,17.11)(35,28.35)(36,46.99)(37,77.89)(38,128.53)(39,212.1)(40,349.98)};
-%\addplot[color=magenta,mark=x] coordinates {(20,0.01)(21,0.01)(22,0.01)(23,0.01)(24,0.02)(25,0.03)(26,0.05)(27,0.08)(28,0.12)(29,0.21)(30,0.35)(31,0.57)(32,0.97)(33,1.65)(34,2.69)(35,4.18(36,7.36)(37,11.66)(38,20.35)(39,32.55)(40,74.46) };
-%          \addplot [color=green,mark=x] coordinates {
-%			(20,0.003)	
-%            (21,0.004)
-%            (22,0.006)
-%            (23,0.009)
-%            (24,0.011)
-%            (25,0.014)
-%            (26,0.018)
-%            (27,0.029)
-%            (28,0.046)
-%            (29,0.069)
-%            (30,0.110)
-%            (31,0.165)
-%            (32,0.261)
-%            (33,0.417)
-%            (34,0.755)
-%            (35,1.192)
-%            (36,1.956)
-%            (37,3.134)
-%            (38,5.089)
-%            (39,8.973)
-%            (40,14.336)
-%        };
-%	%\legend style={anchor=north,at={(20,1000)}}  
-%        \legend{\tiny breadth-expl,\tiny depth-expl,\tiny depth-expl-$\delta$,\tiny depth-expl-$\delta$-mmx}
-%        \end{semilogyaxis}
-%    \end{tikzpicture}
-%    \label{F:Time:new}
-%    \caption{Comparison of time executions of different exploration algorithms of the tree $\mathcal{T}^{\leq g}$ on an Intel \tt{i5-3570K} CPU with $8GB$ of memory. 
-%The scale is logarithmic in time. 
-%The first version is based on a breadth-first search exploration and the peak at genus~$35$ is due to the consumption of all the memory and use of swap. 
-%The second version uses a depth first search exploration.
-%The third is based on the second one with use of decomposition numbers, it corresponds to the algorithm given in Section~\ref{S:Algo}. 
-%The last one is an optimisation of the third one with \MMX.}
-%\end{figure}
-
-%%\begin{figure}[b]
-%%\begin{tikzpicture}[scale=1.3]
-%%        \begin{semilogyaxis}[xlabel=\small genus,ylabel=time,legend pos=north west,legend cell align=left]
-%%        \addplot [color=blue,mark=x] coordinates {
-%%(20,0.045767)
-%%(21,0.067697)
-%%(22,0.112697)
-%%(23,0.194646)
-%%(24,0.335412)
-%%(25,0.581481)
-%%(26,1.01088)
-%%(27,1.72613)
-%%(28,2.9324)
-%%(29,4.91856)
-%%(30,8.39383)
-%%(31,14.0308)
-%%(32,23.7138)
-%%(33,39.6461)
-%%(34,66.1154)
-%%(35,109.406)
-%%(36,182.331)
-%%(37,302.765)
-%%        };
-%%       \addplot [color=black,mark=x] coordinates {
-%%(20,0.043697)
-%%(21,0.065487)
-%%(22,0.113301)
-%%(23,0.186903)
-%%(24,0.31748)
-%%(25,0.53251)
-%%(26,0.899982)
-%%(27,1.5205)
-%%(28,2.55423)
-%%(29,4.28809)
-%%(30,7.30236)
-%%(31,12.1284)
-%%(32,20.2132)
-%%(33,33.707)
-%%(34,56.5415)
-%%(35,93.7614)
-%%(36,157.09)
-%%(37,259.997)
-%%(38,427.497)
-%%(39,709.564)
-%%(40,1192.81)
-%%        };
-%%         \addplot [color=red,mark=x] coordinates {
-%%(20,0.027708)
-%%(21,0.044416)
-%%(22,0.074789)
-%%(23,0.122006)
-%%(24,0.203155)
-%%(25,0.344143)
-%%(26,0.594073)
-%%(27,1.07134)
-%%(28,1.622)
-%%(29,2.82573)
-%%(30,4.53882)
-%%(31,7.60975)
-%%(32,12.5442)
-%%(33,21.0466)
-%%(34,34.7638)
-%%(35,57.7621)
-%%(36,95.6878)
-%%(37,158.95)
-%%(38,264.495)
-%%(39,436.146)
-%%(40,719.347)
-%%        };
-%%        \addplot [color=magenta,mark=x] coordinates {
-%%(20,0.001562)
-%%(21,0.001957)
-%%(22,0.002711)
-%%(23,0.003811)
-%%(24,0.005605)
-%%(25,0.00861)
-%%(26,0.013127)
-%%(27,0.020991)
-%%(28,0.036053)
-%%(29,0.05879)
-%%(30,0.095882)
-%%(31,0.158853)
-%%(32,0.259237)
-%%(33,0.441074)
-%%(34,0.750349)
-%%(35,1.2067)
-%%(36,1.9608)
-%%(37,3.19713)
-%%(38,5.33436)
-%%(39,9.2136)
-%%(40,14.5226)
-%%        };
-%%	%\legend style={anchor=north,at={(20,1000)}}  
-%%        \legend{\tiny breadth-expl,\tiny depth-expl,\tiny depth-expl-$\delta$}
-%%        \end{semilogyaxis}
-%%    \end{tikzpicture}
-%%    \label{F:Time}
-%%    \caption{Comparison of time executions of different exploration algorithms of the tree $\mathcal{T}^{\leq g}$ on an Intel \tt{i7-3840QM} CPU with $32GB$ of memory. 
-%%The scale is logarithmic in time. 
-%%The first version is based on a breadth-first search exploration and the peak at genus~$35$ is due to the consumption of all the memory and use of swap. 
-%%The second version uses a depth first search exploration.
-%%The third is based on the second one with use of decomposition numbers, it corresponds to the algorithm given in Section~\ref{S:Algo}. 
-%%The last one is an optimisation of the third one with \SSE but not using \CilkP.}
-%%\end{figure}
-
-\footnotesize
-
-\begin{thebibliography}{10}
-
-\bibitem{BrasAmoros2008}{M.~Bras-Amor{\'o}s, \emph{Fibonacci-like behavior of the number of numerical semigroups of a given genus}, Semigroup Forum \textbf{76}, no.~2 (2008), 379--384}
-
-\bibitem{Borie}{N.~Borie, \emph{Generation modulo the action of a permutation
-      group}, Proceeding of The 25th International Conference on Formal Power
-    Series and Algebraic Combinatorics (FPSAC '13), DMTCS Proceedings \textbf{0}(01) (2013),
-    767--778}
-
- \bibitem{Delgado}{M.~Delgado, homepage, \url{http://cmup.fc.up.pt/cmup/mdelgado/numbers/}}
-
-\bibitem{NumericalSgps}{M.~Delgado, P.A.~Garcia-Sanchez, J.~Maris, \emph{NumericalSgps} -- GAP package, ver.0.971, 2001 \url{http://cmup.fc.up.pt/cmup/mdelgado/numericalsgps/}}
-
-\bibitem{GAP}{GAP -- Groups, Algorithms, and Programming, Version 4.6.3, \url{http://www.gap-system.org}}
-
-\bibitem{BookNS}{P.A.~Garsia-Sanchez, J.C.~Rosales, \emph{Numerical semigroups}, Springer, New York, 2009, Developments in Mathematics, vol.~20, x+181}
-
-\bibitem{GCCcilk}{B.V.~Iyer, R.~Geva, P.~Halpern, \emph{Cilk\texttrademark{}+ in GCC}
-,GNU Tools Cauldron, 2012, \url{http://gcc.gnu.org/wiki/cauldron2012?action=AttachFile&do=get&target=Cilkplus_GCC.pdf}}
-
-\bibitem{BookDFP}{J.L.~Ram{\'{\i}}rez Alfons{\'{\i}}n, \emph{The Diophantine Frobenius Problem}, Oxford University press, 2005, Oxford Lecture Series in Mathematics and its Applications, vol.~30, xvi+243}
-
-\bibitem{Rosales}{J.C.~Rosales, \emph{Fundamental gaps of numerical semigroups generated by two elements}, Linear Algebre Appl. \textbf{405} (2005), 200--208}
-
-\bibitem{OEIS}{N.J.A.~Sloane, \emph{The On-Line Encyclopedia of Integer Sequences}, \url{http://oeis.org}}
-
-\bibitem{IntelSSE}{Software.intel.com, \emph{Intel Instruction Set Architechture Extensions}, Intel\textregistered{} Developer Zone, 2013, \url{https://www.cilkplus.org/}}
-
-\bibitem{CilkIntel}{Software.intel.com, \emph{Intel\textregistered{} Cilk\texttrademark{} Homepage}, \url{ttps://www.cilkplus.org/}}
-
-\bibitem{CilkRefman}{Software.intel.com, \emph{Intel\textregistered{} Cilk\texttrademark{} Plus Reference Guide},\url{https://software.intel.com/en-us/node/522579}}
-
-\bibitem{AVX2}{Wikipedia, \emph{Advanced Vector Extension}, \url{http://en.wikipedia.org/wiki/Advanced_Vector_Extensions}}
-
-\bibitem{WikipediaDuff}{Wikipedia, \emph{Duff's device}, \url{http://en.wikipedia.org/wiki/Duff's_device}}
-
-\bibitem{WikipediaHT}{Wikipedia, \emph{Hyper-Threading}, 
-\url{http://en.wikipedia.org/wiki/Hyper-Threading}}
-
-\bibitem{WikipediaSIMD}{Wikipedia, \emph{SIMD}, \url{http://en.wikipedia.org/wiki/SIMD}}
-
-\bibitem{WikipediaSSE}{Wikipedia, \emph{Streaming SIMD Extensions}, \url{http://en.wikipedia.org/wiki/Streaming_SIMD_Extensions}}
-
-\bibitem{WikipediaTB}{Wikipedia, \emph{Turbo Boost}, 
-\url{http://en.wikipedia.org/wiki/Intel_Turbo_Boost}}
-
-\bibitem{Wilf}{H. S. Wilf, \emph{A circle-of-lights algorithm for the money-changing problem}, Amer. Math. Monthly \textbf{85} (1978), 562--565}
-
-\bibitem{Zhai}{A. Zhai, \emph{Fibonacci-like growth of numerical semigroups of a given genus}, Semigroup Forum\textbf{86}, no.3 (2013), 634--662}
-\end{thebibliography}
-
-\vspace{2em}
-
-\noindent \textbf{Jean Fromentin}
-
-Univ. Lille Nord de France, F-59000 Lille, France
-
-ULCO, LMPA J.~Liouville, B.P. 699, F-62228 Calais, France
-
-CNRS, FR 2956, France
-
-\url{fromentin@lmpa.univ-littoral.fr}
-
-\vspace{2em}
-
 \noindent \textbf{Florent Hivert}
@@
 
-Bâtiment 650, Universit\'e Paris Sud 11, 91405 Orsay CEDEX, France
+B\^atiment 650, Universit\'e Paris Sud 11, 91405 Orsay CEDEX, France
 

### 2015-04-10 Papier/article.tex
Revisions asked by referees

@@
+\synctex=1
 \documentclass[reqno]{amsart}
@@
 \usepackage{xspace}
-\usepackage{pdfsync}
+%\usepackage{pdfsync}
 
@@
 \author{Jean Fromentin and  Florent Hivert}
-\date{3 November 2014}
+\date{April 10,  2014}
 \keywords{semigroups, tree, algorithm, \texttt{SSE}, 
@@
   In this paper we describe an algorithm visiting all numerical semigroups
-  up to a given genus using a new representation. The
+  up to a given genus using a well suited representation. The
   interest of this algorithm is that it fits particularly well the
@@
 On his home page~\cite{Delgado}, M. Delgado  gives the value of $n_{55}$.
-The values of $n_{53}$ and $n_{54}$ seem to be so far available nowhere.
 
@@
 semigroups $\mathcal{T}$ and achieve the computation of $n_g$ for $g\leq 67$.
-The cornerstone of our method is a new combinatorial representation of
+The cornerstone of our method is a combinatorial representation of
 numerical semigroups that is well suited and allows large code optimization essentially based on the use of
@@
 The goal of the paper is twofold:
-first to present our new encoding of numerical semigroups and the associated
+first to present our encoding of numerical semigroups and the associated
 algorithms, and second to present the optimization techniques which allow, for
@@
 In Section~\ref{S:Tree} we describe the tree of numerical semigroups and give bounds for some parameters attached to a numerical semigroup.
-The description of our new representation of numerical semigroups is done in the second section.
+The description of our representation of numerical semigroups is done in the second section.
 In Section~\ref{S:Algo} we describe an algorithm based on the representation given in Section~\ref{S:DecNumber} and give its complexity. 
@@
 We start this section with definitions and properties of numerical semigroups that will be used in the sequel.
-For a more complete introduction, the reader can usefully consult the book \emph{Numerical Semigroups} of J.C.~Rosales and P.A.~Garc\'ia-S\'anchez \cite{BookNS} or the book \emph{The Diophantine Frobenius Problem} of J.L.~Ram\'irez~Alfons\'in \cite{BookDFP}.
+For a more complete introduction, the reader can usefully consult the book \emph{Numerical Semigroups} by J.C.~Rosales and P.A.~Garc\'ia-S\'anchez \cite{BookNS} or the book \emph{The Diophantine Frobenius Problem} by J.L.~Ram\'irez~Alfons\'in \cite{BookDFP}.
 
@@
 
-$iii)$ $c(S)=1+\max(\NN\setminus S)$, the \emph{conductor} of $S$ for $S$ different from $\NN$. By convention, the conductor of $\NN$ is $0$. 
+$iii)$ $f(S)=\max(\mathbb{Z}\setminus S)$, the \emph{Frobenius} of $S$;
+
+$iv)$ $c(S)=f(S)+1$, the \emph{conductor of $S$}. 
 \end{defi}
@@
 
-\begin{lem}[Lemma 2.3, Chap. I, \cite{BookNS}]
+\begin{lem}[Lemma 2.3 of~\cite{BookNS}]
 For a numerical semigroup $S$, the set $\Irr(S)$ is the minimal generating set of $S$ relatively to the inclusion ordering.
@@
 
-\begin{prop}
+\begin{prop}[Proposition~2.12 and Lemma~2.14 of~\cite{BookNS}]
 \label{P:Res}
@@
 
-\begin{proof}
-$i)$ Let $x$ be a positive integer satisfying $x \geq c(S)+m(S)$. 
-As the integers $x$ and $x-m(S)$ are greater than $c(S)$, they belong to $S$.
-Since $x$ equals $(x-m(S))+m(S)$ with $x-m(S)\not=0$ and $m(S)\not=0$, it cannot be irreducible.
-
-$ii)$ The set $\NN\setminus{S}$ of cardinality $g(S)$ contains $\{1,..,m(S)-1\}$ and so the relation $m(S)\leq g(S)+1$ holds.
-
-$iii)$ Let $x$ be an element of $S\cap\{0,...,c(S)-1\}$. 
-The integer $y=c(S)-1-x$ lies in $\{0,...,c(S)-1\}$. 
-Moreover $y$ is not in $S$, for otherwise  we write $c(S)-1=y+x$ where $x,y$ are elements of $S$ implying $c(S)-1$ belongs to $S$. 
-In contradiction with the definition of conductor. 
-Thus we define an involution 
-\[
-\begin{array}{rcl}
-\psi:\{0,...,c(S)-1\}&\to&\{0,...,c(S)-1\}\\
-x&\mapsto&c(S)-1-x
-\end{array}
-\]
-that sends $S\{0,...,c(S)-1\}$ into $\{0,...,c(S)-1\}\setminus S$ which is  of cardinality~$g(S)$. 
-Let us denote by $k$ the cardinality  of $S\cap\{0,...,c(S)-1\}$. 
-Since $\psi$ is injective we must have $k\leq g(S)$. 
-From the relation 
-\[
-\{0,...,c(S)-1\}=S\cap\{0,...,c(S)-1\}\sqcup\{0,...,c(S)-1\}\setminus S
-\]
- we obtain $c(S)=k+g(S)$ and so $c(S)\leq 2 g(S)$.
-\end{proof}
-
 A consequence of Proposition~ \ref{P:Res}~$i)$ is that $\Irr(S)$ is finite and its cardinality is at most $c(S)+m(S)-1$.
@@
 We now explain the construction of the tree of numerical semigroups.
-Let~$S$ be a numerical semigroup. The set $S'=S\cup\{c(S)-1\}$ is also a numerical  semigroup and its genus is $g(S)-1$. 
-As  each integer greater than $c(S)-1$ is included in $S'$ we have $c(S')\leq c(S)-1$. 
+Let~$S$ be a numerical semigroup. The set $S'=S\cup\{f(S)\}$ is also a numerical  semigroup and its genus is $g(S)-1$. 
+As  each integer greater than $f(S)$ is included in $S'$ we have $c(S')\leq f(S)$. 
 Therefore every semigroup $S$ of genus $g$ can be obtained from a semigroup $S'$ of genus $g-1$ by removing an element of~$S'$ greater than or equal to $c(S')$.
 
-\begin{prop}
+\begin{prop}[Proposition~7.28 of \cite{BookNS}]
 \label{P:Sx}
@@
 
-\begin{proof}
-If $x$ is not irreducible in  $S$, then there exist $a$ and $b$ in $S\setminus{0}$ such that $x=a+b$. 
-As $a\not=0$ and $b\not=0$ hold, the integers  $a$ and $b$ belong to $S\setminus\{x\}$. 
-Since $x=a+b$ and $x\not\in S^x$, it follows that  $S^x$ is not stable under addition.
-
-Conversely, assume that $x$ is irreducible in  $S$. 
-As $0$ is never irreducible, the set~$S^x$ contains~$0$. 
-Let $a$ and $b$ be two integers belonging to $S^x$. 
-The set~$S$ is stable under addition, hence $a+b$ lies in $S$. 
-As $S$ is equal to $S^x\cup\{x\}$, the integer $a+b$ also lies in $S^x$ except if it is equal to $x$. 
-The latter is impossible since $a$ and $b$ are different from $x$ and $x$ is irreducible.
-\end{proof}
-
 Proposition~\ref{P:Sx} implies that every semigroup $S$ of genus $g$ can be obtained from a semigroup $S$ by removing a generator $x$ of $S$ that is greater than or equal to $c(S)$.
@@
 The root of the tree is the unique semigroup of genus $0$, \ie, $\left<1\right>$ that is equal to $\NN$. 
-If~$S$ is a semigroup in the tree,  the sons of $S$ are exactly the semigroups~$S^x$ where $x$ belongs to $\Irr(S)\cap[c(s),+\infty[$. 
+If~$S$ is a semigroup in the tree,  the sons of $S$ are exactly the semigroups~$S^x$ where $x$ belongs to $\Irr(S)\cap[c(S),+\infty[$. 
 By convention, when depicting the tree, the numerical semigroup $S^x$ is in the left of $S^y$ if $x$ is smaller than~$y$. 
@@
 
-As the reader can check, the main difficulty to characterize the sons of a semigroup is to determine its irreducible elements.
- In \cite{NumericalSgps}, the semigroups are given by their Ap\'ery sets and then the difficulty is to describe the Ap\'ery set of $S^x$  from that of $S$. 
- This approach is elegant but not sufficiently \emph{basic} for our optimizations.
-
 \section{Decomposition number}
@@
 
-The aim of this section is to describe a new representation of numerical semigroups,
+The aim of this section is to describe a representation of numerical semigroups,
 which is well suited to an efficient exploration of the tree~$\mathcal{T}$ of numerical semigroups.
@@
 Assume that $y$ is an element of $D_S(x)$.
-By very definition of $D_S(x)$, the integer~$z=x-y$ also belongs to $S$. 
+By definition of $D_S(x)$, the integer~$z=x-y$ also belongs to $S$. 
 Then $x$ can be decomposed as $x=y+z$ with $y$ and $z$ in~$S$. 
@@
 
+A straightforward consequence of the definition of $S$-decomposition numbers is :
+
 \begin{prop}
@@
 
-\begin{proof}
-We start with $i)$. If $x$ is an element of $S$ then $x$ equals $0+x$. 
-The relation~$2\times 0\leq x$ and $0\in S$ imply that $D_S(x)$ contains $0$, and so $d_S(x)>0$ holds.
-Conversely, the relation  $d_S(x)>0$  implies that~$D_S(x)$ is non-empty. 
-Let $y$ be an element of~$D_S(x)$. 
-As $y$ and $x-y$ belong to  $S$, by definition, the integer $x=(x-y)+y$ is in $S$  (since $S$ is stable by addition).
-
-Let us show $ii)$. 
-Assume $x$ is irreducible in $S$.
-There cannot exist $y$ and $z$ in~$(S\setminus\{0\})^2$ such that~$x=y+z$. 
-The only possible decomposition of $x$ as a sum of two elements of $S$ is $x=0+x$. 
-Hence, the set $D_S(x)$ is equal to~$\{0\}$ and we have $d_S(x)=1$. 
-Conversely, let $x$ be such that $d_S(x)$ is $1$. 
-By~$i)$ the  integer $x$ must be in $S$. 
-As $x=0+x$ is always a decomposition of $x$ as a sum of two elements in $S$, we obtain $D_S(x)=\{0\}$.
-If there exist $y$ and $z$ in $S$ such that $y\leq z$ and $x=y+z$ hold then $y$ lies in~$D_S(x)$. 
-This implies $y=0$ and $z=x$. 
-Hence $x$ is irreducible in~$S$.
-\end{proof}
-
 We note  that $0$ is never irreducible despite the fact $d_S(0)$ is $1$ for all numerical semigroups $S$.
-
 We now explain how to compute the $S$-decomposition numbers function of a numerical semigroup  from that of its father.
@@
 \begin{proof}
-Let $y$ be in $\NN\setminus\{0\}$. We have 
-\[
-D_{S^x}(y)=\{z \in S^x\ |\ y-z\in S^x\ \text{and}\  2z\leq y\}.
-\]
-We construct two sets as follow :
-\begin{align*}
-E&=\{z\in S^x\ |\ y-z=x\ \text{and}\ 2z\leq y\}\\
-F&=\begin{cases}
-\{x\}&\text{if $y-x\in S$ and $2x\leq y$},\\
-\emptyset&\text{otherwise}.
-\end{cases}
-\end{align*}
-We have the relation $D_{S}(y)=D_{S^x}(y)\sqcup E\sqcup F$ and so $d_S(y)$ is equal to $d_{S^x}(y)+\card(E)+\card(F)$
-and $D_{S^x}(y)$ is a subset of $D_S(y)$. We start by determining $E$. 
-The relation $y-z=x$ implies $z=y-x$ and $2z\leq y$ becomes $y\leq 2x$. Thus we obtain
-\[
-E=\begin{cases}
-\{y-x\}&\text{if $y-x\in S^x$ and $y\leq 2x$}\\
-\emptyset&\text{otherwise}.
-\end{cases}
-\] 
-The relation $y-x\in S^x$ together with $y\leq2x$ is equivalent to $y-x\in S$ together with $y<2x$.
-Therefore we obtain
-\[
-E=\begin{cases}
-\{y-x\}&\text{if $y-x\in S$ and $y< 2x$}\\
-\emptyset&\text{otherwise}.
-\end{cases}
-\] 
-Assume $y-x$ lies in $S$.  If $y<2x$ holds, we have $E=\{y-x\}$ and $F=\emptyset$. Otherwise, we obtain $E=\emptyset$ and $F=\{x\}$. 
-Then if $y-x\in S$ holds, we have $\card(E)+\card(F)=1$ and so $d_{S}(y)$ is equal to $d_{S^x}(y)+1$.
-If $y-x$ does not belong to $S$, then $E$ and $F$ are empty implying $d_{S}(y)=d_{S^x}(y)$.
-As the condition $y-x\in S$ is equivalent to $y\geq x$ together with $d_S(y-x)>0$, by Proposition~\ref{P:Dandd} the result holds.
+A direct consequence of~$D_{S^x}(y)=D_S(y)\setminus\{y-x,x\}$.
 \end{proof}
@@
 instead a breadth first search exploration.  The main advantage in our
-approach is the small memory needs.  The cost to pay is that, if we want to
-explore the tree deeper, we must restart from the root.
+approach is the small memory needs.  
+%[MODIF]
+Indeed, in the case of breadth first search algorithm one needs to compute and store the list $L_g$ of all numerical semigroups of genus $g$ before 
+visiting the numerical semigroups of genus $g+1$, which is not required for a depth first search exploration.
+In this paper we are interested in the exploration of the list~$L_g$ for $g\in \NN$, not in storing it.
+This is the reason we use a depth first search algorithm for the exploration of the tree $\mathcal{T}_g$.
+The only limitation is then the duration of the exploration and not the amount of available memory.
+For example the list~$L_{54}$ needs several terabytes to be stored. 
+%[/MODIF]
 
@@
 \label{P:delta}
-Let $G$ be an integer and $S$ be a numerical semigroup of genus $g\leq G$.
+Let $G$ be an integer and $S$ be a numerical semigroup of genus $0<g\leq G$.
 Then $S$ is entirely described by the vector $\delta_S=(d_S(0),...,d_S(3\,G)) \in \NN^{3G+1}$. More precisely we can obtain $c(S), g(S), m(S)$ and $\Irr(S)$ from $\delta_S$.
@@
 By Proposition~\ref{P:Res} $iii)$ we have  the relation $c(S)\leq 2g(S)$ and so the $S$-decomposition number of~$c(S)$ occurs in $\delta_S$.
-Since $c(S)$ is equal to $1+\max(\NN\setminus S)$ for $S\not=\NN$, Proposition~\ref{P:delta} implies 
+Proposition~\ref{P:delta} implies 
 \[
@@
 \[
-\Irr(S)=\{i \in\{0,...,3G\},\ d_S(i)=1\}.
+\Irr(S)=\{i \in\{0,...,3G\},\ d_S(i)=1\}.\qedhere
 \]
 \end{proof}
+
+The previous representation of numerical semigroup (in terms of the vector $\delta_S$)  is similar but a little different from this used in \cite{Maria} and by people concerned with coding theory.
 
@@
 \EndIf
-\State\tt{$\tt{S}^\tt{x}$.d} $\gets$ \tt{S.d}
+\State\tt{$\tt{S}^\tt{x}$.d} $\gets$ \tt{S.d}\Comment{copy all the decomposition numbers}
 \For{\tt{y} from \tt{x} to $3\,\tt{G}$}
 	\If{$\tt{S.d}[\tt{y}-\tt{x}]>0$}
-		\State $\tt{$\tt{S}^\tt{x}$.d}[\tt{y}]\gets\tt{$\tt{S}^\tt{x}$.d}[\tt{y}]-1$
+		\State $\tt{$\tt{S}^\tt{x}$.d}[\tt{y}]\gets\tt{$\tt{S}$.d}[\tt{y}]-1$ \Comment{decrease the decomposition number by $1$}
 	\EndIf
@@
 \label{P:Son}
-Running on $(S,x,G)$ with $g(S)\leq G$, $x\in \Irr(S)$ and~$x\leq c(S)$, Algorithm~\ref{A:Son} returns the semigroup $S^x$ in time $O(\log(G)\times G)$.
+Running on $(S,x,G)$ with $g(S)\leq G$, $x\in \Irr(S)$ and~$x\geq c(S)$, Algorithm~\ref{A:Son} returns the semigroup $S^x$ in time $O(\log(G)\times G)$.
 \end{prop}
@@
 Every integer of $I=[x+1,+\infty[$ lies in $S$ since $x$ is greater than~$c(S)$, so the interval $I$ is included in $S^x$.
-As $x$ does not belong to $S$, the conductor of $S^x$ is $x+1$, see Line~\tt{2}.
+As $x$ does not belong to $S^x$, the conductor of $S^x$ is $x+1$, see Line~\tt{2}.
 For the multiplicity of $S^x$ we have two cases. 
@@
 As the number of sons of a  semigroup $S$ is the number of $S$-irreducible elements in the set $\{c(S),...,c(S)+m(S)-1\}$,  a semigroup $S$ has at most $m(S)$ sons. 
-By~Proposition~\ref{P:Res} $ii)$, this implies that a semigroup of genus $g$ as at most $g+1$ sons.
+By~Proposition~\ref{P:Res} $ii)$, this implies that a semigroup of genus $g$ has at most $g+1$ sons.
 Therefore the stack contains at most $g+1$ semigroups of genus $g+1$ for $g\leq G$. 
@@
 \[
-O(\log(G)\times G\times M)= O(\log(G)\times G^3).
+O(\log(G)\times G\times M)= O(\log(G)\times G^3).\qedhere
 \]
@@
 
-\item Vectorization (\MMX, \SSE instructions sets) and careful memory alignment. 
+\item Vectorization (\MMX, \SSE instructions sets) and careful memory alignment; 
 \item Shared memory multi-core computing using \CilkP for low level
@@
 \item Avoiding all unnecessary copy (the cost of the \textsc{Son} algorithm is
-  roughly the same as copying) !
+  roughly the same as copying);
 \item Aggressive loop unrolling: the main loop is unrolled by hand using some
-  kind of Duff's device.
+  kind of Duff's device;
 \item Careful choice of data type (\verb|uint_fast8_t| for decomposition
@@
 
-The source code of our algorithm is available here \cite{code}.
+The source code of our algorithm is available in \cite{code}.
 \subsection{Vectorization}
@@
 We used \SSEV~\cite{WikipediaSSE, IntelSSE} technology as it allows for the largest speedup\footnote{A much greater  speedup can be certainly obtained using \texttt{AVX2} technology \cite{AVX2}.
-However, at this time, we can't access a performant computer with this set of instructions.}.
+However, at this time, we cannot access a performant computer with this set of instructions.}.
 This need to respect some constraints in the memory
@@
 following \CPP code
-\vspace{1em}
+%\vspace{1em}
 {
@@
 \end{alltt}}
-\vspace{1em}
+%\vspace{1em}
 
@@
 shift = x & 0xF;     // offset of x inside the block
-...                  // some specific instructions to
-...                  // handle the first incomplete block.
+...                  // some specific instructions to handle the first incomplete block.
 for (long int i=start+1; i<NBLOCKS; i++) {
@@
 which contains $0$ in the bytes corresponding the $0$ entries of block and
-$255$ in the non zero one. This results is then bitwise anded with one so that
+$255$ in the non zero one. This result is then bitwise anded with one so that
 the instruction actually performs a 16~bytes parallel version of
@@
 Our second optimization is to use parallelism on exploration of the tree.
-Today, CPU of personal computers have several cores (2, 4 or more).  The given
+Today, CPUs of personal computers have several cores (2, 4 or more).  The given
 version of our exploration algorithm uses a single core and so a fraction only
@@
 
+\vspace{1em}
 {\small
@@
 }
-
-We just tells \CilkP that the subtrees rooted at various children can be
+\vspace{1em}
+
+We just tell \CilkP that the subtrees rooted at various children can be
 explored in parallel.
@@
 \subsubsection*{Avoiding all unnecessary copy} We also used a trick to avoid
-copying from and to the top of the stack. Indeed, then main loop performs the
+copying from and to the top of the stack. Indeed, the main loop performs the
 following sequences of operations:
@@
 the stack memory but we have to copy the top of the stack to \tt{S}. 
-In \cite{Zhai}, A.Zhai establishes that the limit of the quotient   $\frac{n_g}{n_{g-1}}$, when $g$ go to infinity, is the golden ratio $\phi\approx 1.618$. Therefore this single copy is far from
+In \cite{Zhai}, A.~Zhai establishes that the limit of the quotient   $\frac{n_g}{n_{g-1}}$, when $g$ go to infinity, is the golden ratio $\phi\approx 1.618$. Therefore this single copy is far from
 being negligible. The trick is to use a level of indirection, replacing the
@@
 the derecursived stack algorithm, we only allocate (on the system stack rather
-than on the heap) the stack of semigroups. No further allocation are done.
+than on the heap) the stack of semigroups. No further allocations are done.
 
@@
 
-Due to the way, the \CPP does its pointer arithmetic, even if the index in the
+Due to the way \CPP does its pointer arithmetic, even if the index in the
 array are less than $3G$ and therefore fits in $8$~bits, we use $64$~bits
@@
 Running the \Cilk version of our optimized algorithm we have explored the tree
-of numerical semigroups up to genus $67$. The computation were done one a shared
+of numerical semigroups up to genus $67$. The computations were done on a shared
 $64$ core AMD Opteron\texttrademark{} Processor 6276. As other heavy
-computation were running on the machine, we only used $32$ cores. The
-computation took $18$~days. The values of~$n_g$ for $g\leq 67$ are :
+calculations were running on the machine, we only used $32$ cores. The
+computations took $18$~days. The values of~$n_g$ for $g\leq 67$ are:
 
@@
 With our exploration algorithm we have proved that there is no counterexample to Wilf's conjecture up to genus $60$.
+%[MODIF]
+We have tested the Wilf's conjecture on a different machine than the one used to determine $n_{67}$.
+As its performance is lower we have only tested the conjecture for $g\leq 60$. 
+%[\MODIF]
 
@@
 based on a breadth exploration of the tree while Algorithm \texttt{depth} 
-use a depth exploration. Algorithm \texttt{depth+$\delta$} is a refinement of 
+use a depth exploration. 
+%[MODIF]
+These two algorithms are based on the same naive representation of numerical semigroups.
+The only difference concerns the tree exploration algorithm used.
+%[\MODIF]
+Algorithm \texttt{depth+$\delta$} is a refinement of 
 \texttt{depth} based on the $S$-decomposition function. Algorithm 
 \texttt{$\delta$+sse} in an optimization of \texttt{depth+$\delta$} using the 
-\SIMD extension \SSE. Times are in second.
+\SIMD extension \SSE. Times are in seconds.
+
+\vspace{1em}
 
@@
 
-The computation of $n_g$ for $g\leq 33$ with algorithm \texttt{breadth} is very 
+\vspace{1em}
+
+The computation of $n_g$ for $g\leq 35$ with algorithm \texttt{breadth} is very 
 long because all the \texttt{8GB} of memory are consumed and the system 
@@
 The following table illustrates the impact of parallelization on the same 
-machine based on the \tt{Intel\texttrademark{} i5-3570K} CPU. This CPU has two physical cores that are able 
-to run four threads thanks to the \texttt{Hyper-Threading} 
-technology~\cite{WikipediaHT}. However, when more 
-than two threads are running, the computation engines are shared between two 
-threads.
+machine based on the \tt{Intel\texttrademark{} i5-3570K} CPU which have $4$ physical cores that are able to run four threads.
 
@@
 
-\vspace{-0.2em}
-
+%\vspace{-1em}
 The time of the one threaded algorithm  must be compared with the 
@@
 at \texttt{2.67GHz}.
-Each of those CPU has 6 physical cores that are able to run 12 threads : the 
-machine has also 12 physical cores.  
-For more than $12$ threads, the speedup should be much less when
-adding more cores. The following table resumes the time needed by the algorithm 
+%[MODIF]
+Each of those CPU has 6 physical cores that are able to run 12 threads  thanks to the \texttt{Hyper-Threading} 
+technology~\cite{WikipediaHT}: the 
+machine has also 12 physical cores and  is able to run 24 threads.
+However, when more 
+than $12$ threads are running, the computation engines are shared between two 
+threads and the speedup should be much less when
+adding more cores. 
+%[/MODIF]
+The following table resumes the time needed by the algorithm 
 to explore $\mathcal{T}_{50}$ on different numbers of threads (the double 
@@
  \hline
- Time (s) & 3588.08 & 3709 & 1865 & 932.4 & 486.8 & 325.7 & 311.2 & 302.3 & 290.2\\
+ Time (s) & 3588 & 3709 & 1865 & 932.4 & 486.8 & 325.7 & 311.2 & 302.3 & 290.2\\
  \hline
@@
 \end{figure}
+
+%[MODIF]
+Further improvements on the computation of $n_g$  or on the verification of Wilf's conjecture will be published on our home pages and on \cite{code}.
+%[\MODIF]
+\vspace{2em}
 
@@
 \bibliography{biblio}
-\vspace{2em}
+\newpage
 

### 2015-04-10 Papier/article.tex
Added thanks to L.G.

@@
 
-\textbf{Acknowledgment.} We would like to thank Shalom Eliahou for introducing us to the tree of numerical semigroups and for interesting
-discussions on the subject. We also would like to thanks Nathan Cohen who
-suggested the two levels of indirection stack optimization.
+\textbf{Acknowledgment.} We would like to thank Shalom Eliahou for introducing
+us to the tree of numerical semigroups and for interesting discussions on the
+subject. We are grateful to Laura Grigori who suggested us that Cilk could be
+a good technology to tackle the tree-level parallelism for this kind of
+problems.  We would finally like to thanks Nathan Cohen who suggested the two
+levels of indirection stack optimization.
 

### 2015-04-10 Papier/modifs.tex
Response to referees + modifs

\documentclass{amsart}

\begin{document}

\begin{center}
\textbf{Response to Referee's comment}

 Jean Fromentin and Florent Hivert

 April 10, 2014
\end{center}

\vspace{2em}

We thank the two referees for their useful remarks and comments.
According to referee's suggestion, we corrected typos and minor errors.

The followings are a list of response to the other refree's comments

\vspace{2em}

\section{Report 1C}

\begin{itemize}
\item We didn't compute new values of the $n_g$'s but at the end of the last section we have added references for possible further improvement,
both on the $n_g$'s and on the verfication of Wilf's conjecture.

 \item Following referee's suggestion, we removed the sentence ``The values of $n_{53}$ and $n_{54}$...''.

 \item The referee thinks a ``h'' is missing in ``anded''.  The word ``anded'' is correct and is connected to the bit operator ``and'' in computer science.
 We have remplaced it by ``and-ed'' for clarity.

 \item We have kept the comparision of the depth and breadth first search algorithm, but we have added a comment precising that both algorithms are based on the same naive representation of numerical semigroups.

 \item Following referee's suggestion, we have updated the GAP reference in bibliography.
\end{itemize}

\section{Report 1M}

\begin{itemize}

\item As explained by the referee, the paragraph about the computation of the Ap\'ery set of the son is unfortunate. We have removed it.

\item The referee says that our representation is not completely new and has been used by people concerned with coding theory.
  Indeed the sequence $\{v_i\}_{i\in \mathbb{N}}$ is relatively similar to the one we use but seems to be different. 
  We have then removed the word ``new'' when we talk about our representation. 
  We have also added a sentence, with a reference to [M. Br\`as-Amorr\'os, Addition behavior of a numerical semigroup], after the descripition of our representation.

\item Following referee's suggestion, we defined the Frobenius number $f(S)=\max(\mathbb{Z}\setminus S)$ and replaced each instance of $c(S)-1$ by $f(S)$.

\item Following referee's suggestion, we have removed or simplified some proofs.

\item Following referee's suggestion, we have removed the last sentence of the first paragraph of section 3 and discuss the advantage of a depth first search exploration in terms of memory consumption.

\item About point $(12)$ of referee report we had an explanation for why the Wilf's conjecture is only tested for $g\leq 60$ and not $g\leq 67$.

\end{itemize}
\end{document}

### 2017-05-09 src/Cilk++/alarm.hpp
Fixed Makefile

@@ -1,4 +1,4 @@
-#define PERIOD 10
+#define PERIOD 60
 
 extern void progress_report(void);
 void start_alarm(void);

### 2017-05-12 src/Cilk++/treewalk.cpp
Finer progress report

@@ -52,17 +52,14 @@ results_type cilk_results[128];
 
 void walk_children(const monoid m)
 {
-  unsigned long int nbr = 0;
-
   if (m.genus < MAX_GENUS - STACK_BOUND)
     {
       auto it = generator_iter<CHILDREN>(m);
       while (it.move_next())
 	{
 	  cilk_spawn walk_children(remove_generator(m, it.get_gen()));
-	  nbr++;
+	  cilk_results[__cilkrts_get_worker_number()][m.genus] ++;
 	}
-      cilk_results[__cilkrts_get_worker_number()][m.genus] += nbr;
      }
   else
     walk_children_stack(m, cilk_results[__cilkrts_get_worker_number()]);
@@ -108,17 +105,14 @@ void walk_children_stack(monoid m, ind_t bound, results_type &res)
 
 void walk_children(const monoid &m, ind_t bound)
 {
-  unsigned long int nbr = 0;
-
   if (m.genus < bound - STACK_BOUND)
     {
       auto it = generator_iter<CHILDREN>(m);
       while (it.move_next())
 	{
 	  cilk_spawn walk_children(remove_generator(m, it.get_gen()), bound);
-	  nbr++;
+	  cilk_results[__cilkrts_get_worker_number()][m.genus] ++;
 	}
-      cilk_results[__cilkrts_get_worker_number()][m.genus] += nbr;
      }
   else
     walk_children_stack(m, bound, cilk_results[__cilkrts_get_worker_number()]);

### 2017-05-12 src/Cilk++/alarm.hpp
Configurable progress report period

@@ -1,4 +1,6 @@
+#ifndef PERIOD
 #define PERIOD 60
+#endif
 
 extern void progress_report(void);
 void start_alarm(void);

### 2017-05-14 src/Cilk++/treewalk.hpp
Fixed typo

@@ -22,5 +22,5 @@ extern cilk::reducer_list_append<monoid> cilk_list_results;
 void walk_children(const monoid m);
 void walk_children(const monoid &m, ind_t bound);
 void reset_sizes();
-unsigned long int get_size(results_type &restab);
+unsigned long int get_sizes(results_type &restab);
 void list_children(const monoid &m, ind_t bound);
